---
title: "Metadata Cleaning"
author: "Euan McDonnell"
date: "`r Sys.Date()`"
output: html_document
---



# Notes

# Code

## Pre-amble

```{r}

# Get here if required
if (!require("here", quietly = TRUE)) { install.packages("here") }

# Source metadata
here::i_am("01_cleanMetadata.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

## Main Code

### Read in metadata

```{r}

# Read in metadata
## Cross-sectional
metaCrossSectional_discoData  <- read_xlsx(here::here(baseMetaDir, "preSchool_crossSectional_tallFormat.xlsx"), sheet=1)
metaCrossSectional_validData  <- read_xlsx(here::here(baseMetaDir, "preSchool_crossSectional_tallFormat.xlsx"), sheet=2)

# Read in liveprool collection dates
livDates <- read.table(here::here(baseMetaDir, "liverpoolCollectionDates.tsv"), sep="\t", header=TRUE)
livDates_valid <- data.frame(read_xlsx(here::here(baseMetaDir, "metaDataValidation_withLiverpoolDates.xlsx"), sheet=2))
livDates_valid <- livDates_valid[livDates_valid$Data.Group == "Validation" &
                                 !is.na(livDates_valid$Date.of.collection) &
                                 livDates_valid$Condition == "Healthy",]
livDates_valid <- livDates_valid[,c(1, 10)]
colnames(livDates_valid) <- c("Sample_ID", "Date.of.sample")

```

### Read in additional metadata

```{r}

# Read in additional metadata
addMeta <- lapply(seq(2,8), function(sht) data.frame(read_xlsx(here::here(baseMetaDir, "preSchool_masterSpreadsheet.xlsx"), sheet=sht)))

# Tidy names and remove unique columns
for (i in seq_len(length(addMeta))) { 
  
  # Fix columnames
  colnames(addMeta[[i]]) <- gsub(" ",".",colnames(addMeta[[i]])) 

  # Remove mismatched columns
  addMeta[[i]] <- addMeta[[i]][,!grepl("Notes|Rack",colnames(addMeta[[i]]))]
  
}

# Sort dates
addMeta[[1]]$Date.of.sample <- lubridate::dmy(addMeta[[1]]$Date.of.sample)
livDates$Date.of.sample <- lubridate::dmy(livDates$Date.of.sample)
livDates_valid$Date.of.sample <- lubridate::dmy(livDates_valid$Date.of.sample)

# Format
addMeta <- do.call("rbind", addMeta)

# Set "NK to NA
for (colVal in colnames(addMeta)) { if ("NK" %in% addMeta[[colVal]]) { print(colVal); addMeta[addMeta[[colVal]] == "NK" & !is.na(addMeta[[colVal]]),][[colVal]] <- NA } }

```

### Tidy condition names

```{r}

# Change condition names to more correct/manageable
metaCrossSectional_discoData$Condition <- mgsub(metaCrossSectional_discoData$Condition,
                                                    c("Crohns", "Infectious Osteomyelitis"),
                                                    c("Crohn",  "InfOsteitis"))
metaCrossSectional_validData$Condition <- mgsub(metaCrossSectional_validData$Condition,
                                                c("Crohns", "Infectious Osteomyelitis"),
                                                c("Crohn",  "InfOsteitis"))

```

### Fix variables

```{r}

# Loop over columns and fix repeat symbols
for (colVal in colnames(addMeta)) { 

  # Fix values
  addMeta[[colVal]] <- FixRepeatSymbols(data=addMeta[[colVal]], repeatVals=c("\\", "/"))  
  
}

```

### Rename wonky columns

```{r}

# Rename spaces
colnames(metaCrossSectional_discoData) <- gsub("-| ","_",colnames(metaCrossSectional_discoData))
colnames(metaCrossSectional_discoData) <- gsub("[.]","_",colnames(metaCrossSectional_discoData))
colnames(metaCrossSectional_validData) <- gsub("-| ","_",colnames(metaCrossSectional_validData))
colnames(metaCrossSectional_validData) <- gsub("[.]","_",colnames(metaCrossSectional_validData))

```

### Make time in years

```{r}

# additional metadata
addMeta[grepl("DATE|[?]",addMeta$Date.of.sample),] <- NA
addMeta$Sex <- toupper(addMeta$Sex)

```

### Investigate IDs

#### Check for duplicate IDs

```{r}

# Check IDs
nSampleIDs <- unlist(c(table(metaCrossSectional_discoData$Sample_ID)))
nSampleIDs[nSampleIDs > 1] # Only "MISSING" duplicated, this is fine

# Exclude missing
metaCrossSectional_discoData <- metaCrossSectional_discoData[metaCrossSectional_discoData$Sample_ID != "MISSING",]

```

#### Rename wonky IDs

```{r}

# Rename
metaCrossSectional_discoData$Sample_ID <- gsub("___","_",mgsub(metaCrossSectional_discoData$Sample_ID, c(" ","-","#"),c("_","_","")))
metaCrossSectional_discoData$Sample_ID <- gsub("___","_",mgsub(metaCrossSectional_discoData$Sample_ID, c(" ","-","#"),c("_","_","")))
metaCrossSectional_validData$Sample_ID <- gsub("___","_",mgsub(metaCrossSectional_validData$Sample_ID, c(" ","-","#"),c("_","_","")))
metaCrossSectional_validData$Sample_ID <- gsub("___","_",mgsub(metaCrossSectional_validData$Sample_ID, c(" ","-","#"),c("_","_","")))

# Fix IDs
addMeta$Preschool.sample.ID <- gsub("[*]","",addMeta$Preschool.sample.ID)
addMeta$Preschool.sample.ID <- gsub("___","_",mgsub(addMeta$Preschool.sample.ID, c(" ","-","#"),c("_","_","")))
livDates$Sample_ID <- gsub("[*]","",livDates$Sample_ID)
livDates$Sample_ID <- gsub("___","_",mgsub(livDates$Sample_ID, c(" ","-","#"),c("_","_","")))

addMeta[order(addMeta$Preschool.sample.ID, decreasing=TRUE),]
metaCrossSectional_discoData[order(metaCrossSectional_discoData$Sample_ID, decreasing=TRUE),]

# Add condition as summarised diagnosis
addMeta["Condition"] <- "Oncology"
addMeta$Diagnosis <- gsub(" [Cc]ontrol","",addMeta$Diagnosis)
addMeta[grepl("Infectious osteomyelitis", addMeta$Diagnosis),]["Diagnosis"] <- "InfOsteitis"
addMeta[grepl("Crohn", addMeta$Diagnosis),]["Diagnosis"] <- "Crohn"
addMeta[which(addMeta$Diagnosis %in% c("CNO","Healthy","Hyperphosphatasia","InfOsteitis","Crohn")),]["Condition"] <- addMeta[which(addMeta$Diagnosis %in% c("CNO","Healthy","Hyperphosphatasia","InfOsteitis","Crohn")),]$Diagnosis
table(addMeta$Condition)

# Fix weeks 
addMeta[grepl("eeks",addMeta$Age),]["Age"] <- round(as.numeric(gsub(" [W,w]eeks","",addMeta[grepl("eeks",addMeta$Age),]$Age)) / 52.1775, 1)

```

### Standardise ethnicities

```{r}

# Rename
metaCrossSectional_discoData$Ethnicity <- gsub("american","American",metaCrossSectional_discoData$Ethnicity)
metaCrossSectional_discoData[metaCrossSectional_discoData$Ethnicity == "NK" & !is.na(metaCrossSectional_discoData$Ethnicity),]["Ethnicity"] <- NA
metaCrossSectional_validData$Ethnicity <- gsub("american","American",metaCrossSectional_validData$Ethnicity)
metaCrossSectional_validData[metaCrossSectional_validData$Ethnicity == "NK" & !is.na(metaCrossSectional_validData$Ethnicity),]["Ethnicity"] <- NA

```

#### Exclude conditions

```{r}

# If we want to remove a condition (JIA mostly)
if (length(excludeConditions) > 0 | !is.null(excludeConditions)) {

  # Drop for train/test data
  metaCrossSectional_discoData <- metaCrossSectional_discoData[!metaCrossSectional_discoData$Condition %in% excludeConditions,]
  
  # Drop for validation data
  metaCrossSectional_validData <- metaCrossSectional_validData[!metaCrossSectional_validData$Condition %in% excludeConditions,]
  
}

```

### Plotting factors

#### Age

```{r, fig.height=5, fig.width=12}

# Cross-sectional
## Distribution
metaCrossSectional_discoData$Age <- as.numeric(metaCrossSectional_discoData$Age)
hist(metaCrossSectional_discoData[metaCrossSectional_discoData$Age_Group == "Child",]$Age)
hist(metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",]$Age)

## Age/Sex split
p1 <- ggplot(metaCrossSectional_discoData[metaCrossSectional_discoData$Age_Group == "Child",], aes(y=Age, x=Sex, fill=Sex)) + 
  geom_boxplot(fill="white", width=0.25) +
  geom_beeswarm(alpha=0.65, shape=21) + 
  geom_hline(yintercept=0, linetype="dashed", color="darkgrey") +
  facet_grid(cols=vars(Condition), scale="free_x", space="free_x") +
  scale_fill_manual(values=c("darkorange","mediumpurple")) +
  xlab("Sex") +
  ggtitle("Cross-Sectional Study (Children)") +
  theme_classic() + theme(legend.title=element_blank())
p1

```

#### Site in relation to other features

### Convert between Plate Coordinate ID to metadata ID

#### Read in Coord-to-metadata ID conversions

```{r}

# Read in ID conversion table
idConvertCross <- read_xlsx(here::here(baseMetaDir, "preSchoolProject_crossSectional_plateInfo.xlsx"), sheet=2)
idConvertValid <- read_xlsx(here::here(baseMetaDir, "Validation_crosssectional_plates_randomised_08042025.xlsx"), sheet=2)

# Melt to tall format
idConvertCross <- reshape2::melt(idConvertCross, id.vars="ID")
idConvertValid <- reshape2::melt(idConvertValid, id.vars=c("PLATE","ROW"))
colnames(idConvertCross) <- c("PLATE_COORD","COLUMN","META_ID")
colnames(idConvertValid) <- c("PLATE", "ROW", "COLUMN", "META_ID")

# Get plate coordinate for validation data
idConvertValid["PLATE_COORD"] <- paste0(idConvertValid$PLATE, "_", idConvertValid$ROW, idConvertValid$COLUMN)
idConvertValid <- idConvertValid[,c("PLATE_COORD", "COLUMN", "META_ID", "PLATE", "ROW")]

# Drop NAs
idConvertCross  <- idConvertCross[!is.na(idConvertCross$META_ID),]
idConvertValid  <- idConvertValid[!is.na(idConvertValid$META_ID),]

# Make matching to quantification table
idConvertCross$PLATE_COORD <- gsub("P0","P",paste0(gsub("[-]","_",idConvertCross$PLATE_COORD), idConvertCross$COLUMN))
rownames(idConvertCross) <- idConvertCross$PLATE_COORD
rownames(idConvertValid) <- idConvertValid$PLATE_COORD
max(table(idConvertCross$PLATE_COORD))
max(table(idConvertValid$PLATE_COORD))

# Add specific Plate ID/Column
idConvertCross["PLATE"] <- gsub("_.*","",idConvertCross$PLATE_COORD)
idConvertCross["ROW"] <- gsub(".*_|[0-9]*","",idConvertCross$PLATE_COORD)

# Discard validation coordinates with no sample
idConvertValid <- idConvertValid[idConvertValid$META_ID != "no sample",]

# Extract QCs
idConvertCrossQCs <- idConvertCross[grepl("QC",idConvertCross$META_ID),]

```

#### Read in proteomics table

```{r}

# Read in pivot table
quantTable_crossSectional <- read.table(here::here(baseDiaDir, "spectronaut", "20240716_091619_crossSectionalOnly", 
                                                   "crossSectionalOnly_Report_Protein Quant (Pivot).tsv"), sep="\t", quote="\"", header=TRUE)
quantTable_validData <- read.table(here::here(baseValidDir, "DIAProteomics", "directDIA_spectronautProcessed", "spectronaut", "20250409_PreSchool_2_20250312_ValidationData", 
                                              "2022_0105d_PreSchool_2_20250312_ValidationData_Report_Protein Quant (Pivot).tsv"), sep="\t", quote="\"", header=TRUE)
rownames(quantTable_crossSectional) <- quantTable_crossSectional$PG.ProteinGroups
rownames(quantTable_validData) <- quantTable_validData$PG.ProteinGroups
dim(quantTable_crossSectional)
dim(quantTable_validData)

# Drop BLANK
quantTable_crossSectional <- quantTable_crossSectional[,!grepl("BLANK",colnames(quantTable_crossSectional))]
quantTable_validData <- quantTable_validData[,!grepl("BLANK", colnames(quantTable_validData))]

# Set column names to plate coord-only IDs
colnames(quantTable_crossSectional) <- gsub(".*300424_|.*100524_|.*290524_|_*[0-9]uL.*","",colnames(quantTable_crossSectional))
colnames(quantTable_validData) <- gsub(".*022.0105d_|_[0-9]ul.*|[.]d[.]PG.*","",colnames(quantTable_validData))

# Save all samples with serum QCs (has plate coords as IDs)
write.table(quantTable_crossSectional, 
            here::here(diaDir, "processed", "crossSectional_pivotTable_PGQuantity.plateCoords_withSerumQCs.tsv"), 
            sep="\t", quote=FALSE, row.names=FALSE)
write.table(quantTable_validData, 
            here::here(diaDir, "processed", "validationData_pivotTable_PGQuantity.plateCoords_withSerumQCs.tsv"), 
            sep="\t", quote=FALSE, row.names=FALSE)

# Check no ids of interest are missing and get those of interest
colnames(quantTable_crossSectional)[!colnames(quantTable_crossSectional) %in% rownames(idConvertCross)]
colnames(quantTable_validData)[!colnames(quantTable_validData) %in% rownames(idConvertValid)]
chosenCrossIDs <- colnames(quantTable_crossSectional)[which(colnames(quantTable_crossSectional) %in% rownames(idConvertCross))]
chosenValidIDs <- colnames(quantTable_validData)[which(colnames(quantTable_validData) %in% rownames(idConvertValid))]
if (length(chosenCrossIDs) != nrow(idConvertCross) | all(chosenCrossIDs %in% idConvertCross$META_ID)) { warning("Number of IDs do not match between conversion and quantification tables, or IDs do not match") }
if (length(chosenValidIDs) != nrow(idConvertValid) | all(chosenValidIDs %in% idConvertValid$META_ID)) { warning("Number of IDs do not match between conversion and quantification tables, or IDs do not match") }


```

#### Fix IDs

```{r}

# Edit metadata IDs
idConvertValid$META_ID <- mgsub(idConvertValid$META_ID,
                                c("-"," [#]","^AC"),
                                c("_","_","AC_healthycontrol_AC"))
metaCrossSectional_validData$Sample_ID <- gsub("-","_",metaCrossSectional_validData$Sample_ID)
metaCrossSectional_validData$Sample_ID <- gsub("thy_con","thycon",metaCrossSectional_validData$Sample_ID)
metaCrossSectional_validData$Sample_ID <- gsub("Healthy","healthy",metaCrossSectional_validData$Sample_ID)

# Fix mis-labelled IDs
metaCrossSectional_validData[metaCrossSectional_validData$Sample_ID == "PS_DR_434",]["Sample_ID"] <- "PS_DR_440"
metaCrossSectional_validData[metaCrossSectional_validData$Sample_ID == "AC_healthycontrol_missing",]["Sample_ID"] <- "AC_healthycontrol_AC35"

inConvertNotInMeta <- setdiff(idConvertValid$META_ID,
                              metaCrossSectional_validData$Sample_ID)
inMetaNotInConvert <- setdiff(metaCrossSectional_validData$Sample_ID,
                              idConvertValid$META_ID)

# Check missing valeus
if (length(inMetaNotInConvert) > 0) { stop("Missing values in metadata not in ID conversion table") }

# Read in CPR plate metadata
validPlateMeta <- read_xlsx(here::here(baseMetaDir, "Validation_crosssectional_plates_randomised_08042025.xlsx"), sheet=3)
validPlateMeta$sampleID <- mgsub(validPlateMeta$sampleID,
                                 c("-"," [#]","^AC"),
                                 c("_","_","AC_healthycontrol_AC"))

# Check missing values are hyperphosphatasia
missingSamples <- validPlateMeta[which(validPlateMeta$sampleID %in% inConvertNotInMeta),]

```

#### Get validation QC samples

```{r}

# Get QC samples from abundance data
quantTable_validDataQCs <- quantTable_validData[,grepl("PooledPlasma",colnames(quantTable_validData))]

# Make dataframe
idConvertValidQCs <- data.frame(PLATE_COORD=NA,
                                COLUMN=gsub(".*S1[.]|_[0-9]*|A","",colnames(quantTable_validDataQCs)),
                                META_ID=colnames(quantTable_validDataQCs),
                                ROW=gsub(".*S1[.]|_[0-9]*|[0-9]","",colnames(quantTable_validDataQCs)),
                                PLATE=NA)


```

#### Save QC data

```{r}

# Save ID conversion
write.table(rbind(data.frame(DATASET="CrossSectional",idConvertCrossQCs),
                  data.frame(DATASET="ValidData",idConvertValidQCs)),
            here::here(metaDir, "serumQCs_idConversionMetadata.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

# Save tables
## Serum QC
write.table(quantTable_crossSectional[,c("PG.ProteinGroups", idConvertCrossQCs$PLATE_COORD)],  
            here::here(diaDir, "processed", "QCs", "crossSectional_pivotTable_PGQuantity.plateCoords_justSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(quantTable_validData[,grepl("PG.ProteinGroups|PooledPlasma", colnames(quantTable_validData))],  
            here::here(diaDir, "processed", "QCs", "validationData_pivotTable_PGQuantity.plateCoords_justSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

## Instrument QC
write.table(quantTable_crossSectional[,c("PG.ProteinGroups", colnames(quantTable_crossSectional)[grepl("nstrument",colnames(quantTable_crossSectional))])],  
            here::here(diaDir, "processed", "QCs", "crossSectional_pivotTable_PGQuantity.justInstrumentQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Save unclean metadata

```{r}

# Save metadata
write.table(metaCrossSectional_discoData, 
            here::here(metaDir, "preSchool_crossSectional_allSamples_unClean.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

#### Add ID conversion to metadata

```{r}

# Fix IDs
## Cross-sectional
idConvertCross$META_ID <- mgsub(idConvertCross$META_ID, c("[-]|[#]"," ","Heal","ealthy_control"),c("_","","heal","ealthycontrol"))
metaCrossSectional_discoData$Sample_ID <- mgsub(metaCrossSectional_discoData$Sample_ID, c("Heal","AC_55","AC_38","ealthy_control"),c("heal","AC55","AC38","ealthycontrol"))

idConvertCross$META_ID <- gsub("Heal","heal",idConvertCross$META_ID)
metaCrossSectional_discoData$Sample_ID <- gsub("Heal","heal",metaCrossSectional_discoData$Sample_ID)
metaCrossSectional_discoData$Sample_ID <- gsub("AC_55","AC55",metaCrossSectional_discoData$Sample_ID)

# Combine edited names
idConvert <- idConvertCross

# Drop QCs
idConvertCross <- idConvertCross[!grepl("QC",idConvertCross$META_ID),]

# Drop "BLANK"
idConvertCross <- idConvertCross[idConvertCross$META_ID != "BLANK",]

# Combine edited names
idConvert_noQCs <- rbind(idConvertCross)

# Check overlaps
### In abundance not in metadata - secpected some missing here - JIA etc
if (length(setdiff(idConvertCross$META_ID, metaCrossSectional_discoData$Sample_ID)) > 0) { warning(paste0("Missing IDs in abundance data not in Cross-Sectional metadata :- ",
                                                                                                paste(setdiff(idConvertCross$META_ID, metaCrossSectional_discoData$Sample_ID), collapse=", "))) }

```

### Merge ID conversions to correct

```{r}

# Combine ID conversion with metadata
metaCrossSectional_discoData <- merge(metaCrossSectional_discoData, idConvertCross, by.x="Sample_ID", by.y="META_ID")
metaCrossSectional_validData <- merge(metaCrossSectional_validData, idConvertValid, by.x="Sample_ID", by.y="META_ID")

```

### Save ID Conversion

```{r}

# Subset to just samples
idConvertCross <- idConvertCross[which(idConvertCross$PLATE_COORD %in% colnames(quantTable_crossSectional)),]
idConvertValid <- idConvertValid[which(idConvertValid$PLATE_COORD %in% colnames(quantTable_validData)),]

# Save
write.table(rbind(idConvertCross,
                  idConvertValid),
            here::here(metaDir, "idConversionProcessed.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Save ID converted QC samples

```{r}

# Get full quant table with all samples and plate coordinate column names
colnames(quantTable_crossSectional) <- gsub(".*300424_|.*100524_|.*290524_|_*[0-9]uL.*","",colnames(quantTable_crossSectional))
colnames(quantTable_validData) <- gsub(".*300424_|.*100524_|.*290524_|_*[0-9]uL.*","",colnames(quantTable_validData))
colnames(quantTable_crossSectional)
colnames(quantTable_validData)

# Change column IDs to sample names
quantTable_crossSectionalQC <- cbind(PG.ProteinGroups=quantTable_crossSectional$PG.ProteinGroups, quantTable_crossSectional[,idConvertCrossQCs$PLATE_COORD])
colnames(quantTable_crossSectionalQC) <- c("PG.ProteinGroups", idConvertCrossQCs$META_ID)

# Save non-QC samples with sample ID names
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed", "QCs"))
write.table(quantTable_crossSectionalQC[,c("PG.ProteinGroups", idConvertCrossQCs$META_ID)],  
            here::here(diaDir, "processed", "QCs", "crossSectional_pivotTable_PGQuantity.sampleIDs_justSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(quantTable_validData[,grepl("PG.ProteinGroups|PooledPlasma", colnames(quantTable_validData))],  
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed", "QCs", "validData_pivotTable_PGQuantity.sampleIDs_justSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Convert IDs

```{r}

# Get full quant table with all samples and plate coordinate column names
colnames(quantTable_crossSectional) <- gsub(".*300424_|.*100524_|.*290524_|_*[0-9]uL.*","",colnames(quantTable_crossSectional))
colnames(quantTable_crossSectional)

# Update rownames for subsetting
rownames(metaCrossSectional_discoData) <- metaCrossSectional_discoData$PLATE_COORD
rownames(metaCrossSectional_validData) <- metaCrossSectional_validData$PLATE_COORD

# Change column IDs to sample names
quantTable_crossSectional <- cbind(PG.ProteinGroups=quantTable_crossSectional$PG.ProteinGroups, quantTable_crossSectional[,idConvertCross$PLATE_COORD])
quantTable_validData <- cbind(PG.ProteinGroups=quantTable_validData$PG.ProteinGroups, quantTable_validData[,idConvertValid$PLATE_COORD])
colnames(quantTable_crossSectional) <- c("PG.ProteinGroups", idConvertCross$META_ID)
colnames(quantTable_validData) <- c("PG.ProteinGroups", idConvertValid$META_ID)

```

### Match metadata and quantification

```{r}

# Get lists of common IDs
commonCrossIDs <- intersect(colnames(quantTable_crossSectional), metaCrossSectional_discoData$Sample_ID); length(commonCrossIDs)
commonValidIDs <- intersect(colnames(quantTable_validData), metaCrossSectional_validData$Sample_ID); length(commonValidIDs)

# Metadata
metaCrossSectional_discoData <- metaCrossSectional_discoData[which(metaCrossSectional_discoData$Sample_ID %in% commonCrossIDs),]
metaCrossSectional_validData <- metaCrossSectional_validData[which(metaCrossSectional_validData$Sample_ID %in% commonValidIDs),]

```

#### Add site data

DR = Dresden, Germany
VA = Vanderbilt University, Tennessee, US
Healthy controls listed as 'AC' - Liverpool, UK
The samples starting with J were all taken from the JSLE study

```{r}

# Add as column
metaCrossSectional_discoData["Site"] <- NA
metaCrossSectional_discoData[grepl("_VA_",metaCrossSectional_discoData$Sample_ID),]["Site"] <- "Vanderbilt"
metaCrossSectional_discoData[grepl("_DR_",metaCrossSectional_discoData$Sample_ID),]["Site"] <- "Dresden"
metaCrossSectional_discoData[grepl("^J|^C",metaCrossSectional_discoData$Sample_ID),]["Site"] <- "Liverpool"

table(metaCrossSectional_discoData$Site)
metaCrossSectional_discoData[is.na(metaCrossSectional_discoData$Site),]

metaCrossSectional_validData["Site"] <- NA
metaCrossSectional_validData[grepl("_VA_",metaCrossSectional_validData$Sample_ID),]["Site"] <- "Vanderbilt"
metaCrossSectional_validData[grepl("_DR_",metaCrossSectional_validData$Sample_ID),]["Site"] <- "Dresden"
metaCrossSectional_validData[grepl("^J|^C",metaCrossSectional_validData$Sample_ID),]["Site"] <- "Liverpool"

```

#### Plot site in relation to other features

##### Cross-sectional data

```{r, fig.width=20, fig.height=8}

# Enumerate smaples per site per condition
metaCrossSectional_child <- metaCrossSectional_discoData[metaCrossSectional_discoData$Age_Group == "Child",]
sitePerCondition <- data.frame(table(metaCrossSectional_child[,c("Site","Condition")]))
sitePerCondition["Prop"] <- NA
platePerCondition <- data.frame(table(metaCrossSectional_child[,c("PLATE","Condition")]))
platePerCondition["Prop"] <- NA
sitePerPlate <- data.frame(table(metaCrossSectional_child[,c("Site","PLATE")]))
sitePerPlate["Prop"] <- NA
for (site in unique(sitePerCondition$Site)) { 
  
  # Calculate proportion of condition samples in site
  sitePerCondition[sitePerCondition$Site == site,]["Prop"] <- sitePerCondition[sitePerCondition$Site == site,]$Freq / 
                                                              sum(sitePerCondition[sitePerCondition$Site == site,]$Freq)
  
  # Calculate proportion of plate by condition
  sitePerPlate[sitePerPlate$Site == site,]["Prop"] <- sitePerPlate[sitePerPlate$Site == site,]$Freq / 
                                                              sum(sitePerPlate[sitePerPlate$Site == site,]$Freq)
  
}
  
for (plate in unique(platePerCondition$PLATE)) { 

  # Calculate proportion of condition samples in plate
  platePerCondition[platePerCondition$PLATE == plate,]["Prop"] <- platePerCondition[platePerCondition$PLATE == plate,]$Freq / 
                                                                  sum(platePerCondition[platePerCondition$PLATE == plate,]$Freq)
  
}

# Plot proportion of samples in each condition wrt site data
SiteCond_plt <- ggplot(sitePerCondition, aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=condCols) +
  xlab("Site") + ylab("Proportion of Site Samples") +
  theme_classic() +
  theme(legend.position="top",
        legend.direction="horizontal")
SitePlate_plt <- ggplot(sitePerPlate, aes(x=Site, y=Prop, fill=PLATE)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=plateCols) +
  xlab("Site") + ylab("Proportion of Site Samples") +
  theme_classic() +
  theme(legend.position="top",
        legend.direction="horizontal")
PlateCond_plt <- ggplot(platePerCondition, aes(x=PLATE, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=condCols) +
  xlab("Plate") + ylab("Proportion of Site Samples") +
  theme_classic() +
  theme(legend.position="top",
        legend.direction="horizontal")
SiteCond_plt + SitePlate_plt + PlateCond_plt

```

##### validation data

```{r, fig.width=14, fig.height=5}

# Enumerate smaples per site per condition
metaCrossSectional_child <- metaCrossSectional_discoData[metaCrossSectional_validData$Age_Group == "Child",]
sitePerCondition <- data.frame(table(metaCrossSectional_validData[,c("Site","Condition")]))
sitePerCondition["Prop"] <- NA
platePerCondition <- data.frame(table(metaCrossSectional_validData[,c("PLATE","Condition")]))
platePerCondition["Prop"] <- NA
sitePerPlate <- data.frame(table(metaCrossSectional_validData[,c("Site","PLATE")]))
sitePerPlate["Prop"] <- NA
for (site in unique(sitePerCondition$Site)) { 
  
  # Calculate proportion of condition samples in site
  sitePerCondition[sitePerCondition$Site == site,]["Prop"] <- sitePerCondition[sitePerCondition$Site == site,]$Freq / 
                                                              sum(sitePerCondition[sitePerCondition$Site == site,]$Freq)
  
  # Calculate proportion of plate by condition
  sitePerPlate[sitePerPlate$Site == site,]["Prop"] <- sitePerPlate[sitePerPlate$Site == site,]$Freq / 
                                                              sum(sitePerPlate[sitePerPlate$Site == site,]$Freq)
  
}
  
for (plate in unique(platePerCondition$PLATE)) { 

  # Calculate proportion of condition samples in plate
  platePerCondition[platePerCondition$PLATE == plate,]["Prop"] <- platePerCondition[platePerCondition$PLATE == plate,]$Freq / 
                                                                  sum(platePerCondition[platePerCondition$PLATE == plate,]$Freq)
  
}

# Plot proportion of samples in each condition wrt site data
SiteCond_plt <- ggplot(sitePerCondition, aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=condCols) +
  xlab("Site") + ylab("Proportion of Site Samples") +
  theme_classic() + 
  theme(legend.position="top",
        legend.direction="horizontal")
SitePlate_plt <- ggplot(sitePerPlate, aes(x=Site, y=Prop, fill=PLATE)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=plateCols) +
  xlab("Site") + ylab("Proportion of Site Samples") +
  theme_classic() + 
  theme(legend.position="top",
        legend.direction="horizontal")
PlateCond_plt <- ggplot(platePerCondition, aes(x=PLATE, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity",
           color="black") + 
  scale_fill_manual(values=condCols) +
  xlab("Plate") + ylab("Proportion of Site Samples") +
  theme_classic() + 
  theme(legend.position="top",
        legend.direction="horizontal")
SiteCond_plt + SitePlate_plt + PlateCond_plt

```

### Add combined site/condition

```{r}

metaCrossSectional_discoData["CondSite"] <- paste(metaCrossSectional_discoData$Condition, metaCrossSectional_discoData$Site, sep="_")
metaCrossSectional_discoData["Condition"] <- gsub("Infectious ","Inf",metaCrossSectional_discoData$Condition)
metaCrossSectional_validData["CondSite"] <- paste(metaCrossSectional_validData$Condition, metaCrossSectional_validData$Site, sep="_")
metaCrossSectional_validData["Condition"] <- gsub("Infectious ","Inf",metaCrossSectional_validData$Condition)

```

### Combine additional and cross sectional metadata

```{r}

# Subset metadata
addMetaCross <- addMeta[which(addMeta$Preschool.sample.ID %in% metaCrossSectional_discoData$Sample_ID),]
addMetaValid <- addMeta[which(addMeta$Preschool.sample.ID %in% metaCrossSectional_validData$Sample_ID),]
addMetaCross <- merge(metaCrossSectional_discoData, addMetaCross, by.x="Sample_ID", by.y="Preschool.sample.ID")
addMetaValid <- merge(metaCrossSectional_validData, addMetaValid, by.x="Sample_ID", by.y="Preschool.sample.ID")

# Check
if (!all(addMetaCross$Sex.x %in% addMetaCross$Sex.y)) { stop("Sex scrambled between cross-sectional and additional metadata") }
if (!all(addMetaCross$Age.x %in% addMetaCross$Age.y)) { stop("Age scrambled between cross-sectional and additional metadata") }
if (!all(addMetaCross$Condition.x %in% addMetaCross$Condition.y)) { stop("Condition scrambled between cross-sectional and additional metadata") }
if (!all(addMetaCross$Ethnicity.x %in% addMetaCross$Ethnicity.y)) { stop("Ethnicity scrambled between cross-sectional and additional metadata") }
if (!all(addMetaValid$Sex.x %in% addMetaValid$Sex.y)) { stop("Sex scrambled between cross-sectional and additional metadata") }
if (!all(addMetaValid$Age.x %in% addMetaValid$Age.y)) { stop("Age scrambled between cross-sectional and additional metadata") }
if (!all(addMetaValid$Condition.x %in% addMetaValid$Condition.y)) { stop("Condition scrambled between cross-sectional and additional metadata") }
if (!all(addMetaValid$Ethnicity.x %in% addMetaValid$Ethnicity.y)) { stop("Ethnicity scrambled between cross-sectional and additional metadata") }

# Remove duplicate columns
addMetaCross <- addMetaCross[,!grepl("[.]y",colnames(addMetaCross))]
addMetaValid <- addMetaValid[,!grepl("[.]y",colnames(addMetaValid))]
colnames(addMetaCross)[grepl("[.]x",colnames(addMetaCross))] <- c("Sex","Age","Ethnicity","Condition")
colnames(addMetaValid)[grepl("[.]x",colnames(addMetaValid))] <- c("Sex","Age","Ethnicity","Condition")

```

```{r}

# match
livDates <- livDates[which(livDates$Sample_ID %in% metaCrossSectional_discoData$Sample_ID),]
livDates_valid <- livDates_valid[which(livDates_valid$Sample_ID %in% metaCrossSectional_validData$Sample_ID),]

# Add liverpool dates
rownames(metaCrossSectional_discoData) <- metaCrossSectional_discoData$Sample_ID
rownames(metaCrossSectional_validData) <- metaCrossSectional_validData$Sample_ID
metaCrossSectional_discoData["Date.of.sample"] <- NA
metaCrossSectional_validData["Date.of.sample"] <- NA
metaCrossSectional_discoData[livDates$Sample_ID,]["Date.of.sample"] <- livDates$Date.of.sample
metaCrossSectional_validData[livDates_valid$Sample_ID,]["Date.of.sample"] <- livDates_valid$Date.of.sample

# Drop mismtached columns
addMetaCross <- addMetaCross[,!grepl("Diagnosis|Box|Tick",colnames(addMetaCross))]
addMetaValid <- addMetaValid[,!grepl("Diagnosis|Box|Tick",colnames(addMetaValid))]

# Add liverpool data
addMetaCross <- rbind(addMetaCross[,colnames(metaCrossSectional_discoData)], 
                      metaCrossSectional_discoData[metaCrossSectional_discoData$Site == "Liverpool",])
addMetaValid <- rbind(addMetaValid[,colnames(metaCrossSectional_validData)], 
                      metaCrossSectional_validData[metaCrossSectional_validData$Site == "Liverpool",])
table(addMetaCross$Site)
table(addMetaCross$Condition)
table(addMetaValid$Site)
table(addMetaValid$Condition)

```

### Binarise Condition

```{r}

# Add bimodal CNO vs other
metaCrossSectional_discoData["BiModal"] <- ifelse(metaCrossSectional_discoData$Condition == "CNO", "CNO", "Other")
metaCrossSectional_validData["BiModal"] <- ifelse(metaCrossSectional_validData$Condition == "CNO", "CNO", "Other")
addMeta["BiModal"] <- ifelse(addMeta$Condition == "CNO", "CNO", "Other")
addMetaCross["BiModal"] <- ifelse(addMetaCross$Condition == "CNO", "CNO", "Other")

```

### Save clean metadata

```{r}

# Save
write.table(metaCrossSectional_discoData, here::here(metaDir, "preSchool_crossSectional_tallFormatClean.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(metaCrossSectional_validData, here::here(metaDir, "preSchool_validationData_tallFormatClean_validationData.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(addMeta, here::here(metaDir, "preSchool_additionalMetadata.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(addMetaCross, here::here(metaDir, "preSchool_additionalMetadata_commonToCrossSectional.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(addMetaValid, here::here(metaDir, "preSchool_additionalMetadata_commonToCrossSectional_validationData.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Save ID converted quantification table

```{r}

# Save non-QC samples with sample ID names
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed"))
write.table(quantTable_crossSectional[,c("PG.ProteinGroups", metaCrossSectional_discoData$Sample_ID)],  
            here::here(diaDir, "processed", "crossSectional_pivotTable_PGQuantity.sampleIDs_noSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(quantTable_validData[,c("PG.ProteinGroups", metaCrossSectional_validData$Sample_ID)],  
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed", 
                       "validationData_pivotTable_PGQuantity.sampleIDs_noSerumQCs.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```







