---
title: "Data Processing, Normalisation & QC"
author: "Euan McDonnell"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
here::i_am("02_dataProcessQC.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

## Load data

### Start list for plots

```{r}

# Initialise, unknown length
plotList <- list()

```

### Metadata

```{r}

# Source metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata.R"))

```

### Quantification table

```{r}

# Read in pivot table
## Read in protein abundance table of actual samples
quantTable_crossSectional <- read.table(here::here(diaDir, "processed", 
                                                   "crossSectional_pivotTable_PGQuantity.sampleIDs_noSerumQCs.tsv"), sep="\t", quote="\"", header=TRUE)
rownames(quantTable_crossSectional) <- quantTable_crossSectional$PG.ProteinGroups

```

### Drop depleted proteins

```{r}

# Depleted prots
depletedProtData <- read.table(here::here(baseDataDir, "annotation", "depletedProteinSet.txt"), sep="\t", header=FALSE)
depletedProts <- depletedProtData$V2

# Drop proteins
nrow(quantTable_crossSectional)
quantTable_crossSectional <- quantTable_crossSectional[!quantTable_crossSectional$PG.ProteinGroups %in% depletedProts,]

```

```{r}

# Load QCs
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_serumQCTables.R"))

# Retain unfiltered data table
quantTable_crossSectional_unFilt <- quantTable_crossSectional

# Rownames
rownames(quantTable_crossSectional) <- quantTable_crossSectional$PG.ProteinGroups

# Visualise
dim(quantTable_crossSectional_serumQC); colnames(quantTable_crossSectional_serumQC)
dim(quantTable_crossSectional_instrQC); colnames(quantTable_crossSectional_instrQC)

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

### Plot abundances of depleted proteins

Test to see group-structure of depleted proteins is present prior to any real normalisation or processing.

```{r}

# Add if depleted protein
qntTable_depletedProt <- reshape2::melt(quantTable_crossSectional[c("P02768","P02763","P01023","P01011"),], id.vars="PG.ProteinGroups")
qntTable_depletedProt <- merge(metaCrossSectional[metaCrossSectional$Site != "Leiden" & metaCrossSectional$Condition != "JIA",], qntTable_depletedProt, by.x="Sample_ID", by.y="variable")

# Plot
ggplot(qntTable_depletedProt, 
       aes(y=log2(value), x=Condition, fill=Condition)) +
  geom_boxplot() + 
  geom_jitter(size=0.75) +
  facet_grid(cols=vars(PG.ProteinGroups), scale="free_y", space="free_y") +
  scale_fill_manual(values=condCols) +
  theme_classic() + theme(axis.text.x=element_blank())

```

### Get proteins detected in all Serum QC samples

```{r}

# Get
rownames(quantTable_crossSectional_serumQC) <- quantTable_crossSectional_serumQC$PG.ProteinGroups
serumProtsCross <- GetSerumQCProteins(serumQCData=quantTable_crossSectional_serumQC[,-1]); length(serumProtsCross)
instrProtsCross <- GetSerumQCProteins(serumQCData=quantTable_crossSectional_instrQC[,-1]); length(serumProtsCross)

```

## Checking for drift

```{r}

# Tall format
quantTable_crossSectional_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_crossSectional_serumQC), 
                                                                    quantTable_crossSectional_serumQC[,2:ncol(quantTable_crossSectional_serumQC)]), id.vars=c("ID"))
quantTable_crossSectional_instrQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_crossSectional_instrQC), 
                                                                    quantTable_crossSectional_instrQC[,2:ncol(quantTable_crossSectional_instrQC)]), id.vars=c("ID"))

# Add metadata
quantTable_crossSectional_serumQC_tall["PLATE"] <- "P1"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC05","QC06","QC07","QC08")),]["PLATE"] <- "P2"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC09","QC10","QC11","QC12")),]["PLATE"] <- "P3"
quantTable_crossSectional_instrQC_tall["PLATE"] <- gsub("late", "", gsub("SerumInstrument.*", "", quantTable_crossSectional_instrQC_tall$variable))
quantTable_crossSectional_instrQC_tall["variable"] <- gsub(".*SerumInstrument", "", quantTable_crossSectional_instrQC_tall$variable)

# Order
quantTable_crossSectional_serumQC_tall <- quantTable_crossSectional_serumQC_tall[order(quantTable_crossSectional_serumQC_tall$variable, decreasing=FALSE),]
quantTable_crossSectional_serumQC_tall$variable <- factor(quantTable_crossSectional_serumQC_tall$variable, levels=unique(quantTable_crossSectional_serumQC_tall$variable))

quantTable_crossSectional_instrQC_tall <- quantTable_crossSectional_instrQC_tall[order(quantTable_crossSectional_instrQC_tall$variable, decreasing=FALSE),]
quantTable_crossSectional_instrQC_tall$variable <- factor(quantTable_crossSectional_instrQC_tall$variable, levels=unique(quantTable_crossSectional_instrQC_tall$variable))

# Plot
p1 <- ggplot(quantTable_crossSectional_serumQC_tall, 
             aes(x=variable,
                 y=log2(value), 
                 group=variable)) + 
  geom_violin(fill="#99E6FF",
              width=0.75,
              alpha=0.75) + 
  geom_boxplot(outlier.size=4,
               width=0.25) + 
  ggtitle("Serum QC (Unfiltered)") +
  xlab("Serum QC Sample") + ylab("Log2 Protein\nAbundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=72)) 
p2 <- ggplot(quantTable_crossSectional_instrQC_tall, 
             aes(x=variable,
                 y=log2(value), 
                 group=variable)) + 
  geom_violin(fill="#99E6FF",
              width=0.75,
              alpha=0.75) + 
  geom_boxplot(outlier.size=4,
               width=0.25) + 
  ggtitle("Instrument QC (Unfiltered)") +
  xlab("Instrument QC Sample") + ylab("Log2 Protein\nAbundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=72)) 

# Save
ggsave(here::here(resCrossDiscoDir, "eda",
                  "crossSectional_serumQCSamples_proteinAbundance_unFiltered.png"),
       p1,
       units="px", width=9000, height=3500)
ggsave(here::here(resCrossDiscoDir, "eda",
                  "crossSectional_instrumentQCSamples_proteinAbundance_unFiltered.png"),
       p2,
       units="px", width=9000, height=3500)

# Add to list
newList <- list(p1, p2)
names(newList) <- c("UnFiltProtAbundanceSerumQC",
                    "UnFiltProtAbundanceInstrQC")
plotList <- append(plotList,
                   newList)

```

## Missing values

### Enumerate frequencies

```{r}

# Enumerate raw NAs
## Protein
perProtNA_cross <- EnumerateNAs(quantTable_crossSectional[,metaCrossSectional$Sample_ID], idxVal=1)
perProtNA_cross_serumQC <- EnumerateNAs(quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idxVal=1)

## Samples
perSmplNA_cross <- EnumerateNAs(quantTable_crossSectional[,metaCrossSectional$Sample_ID], idxVal=2)
perSmplNA_cross_serumQC <- EnumerateNAs(quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idxVal=2)

# Test dropping values on frequencies of missing proteins and samples
perProtNACross_dropProts <- TestProteinDropThresholds(perProtNA=perProtNA_cross, quantTbl=quantTable_crossSectional[,metaCrossSectional$Sample_ID], idxVal=1)
perProtNACross_dropProts_serumQCOnly <- TestProteinDropThresholds(perProtNA=perProtNA_cross_serumQC, quantTbl=quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idxVal=1)

perSmplNACross_dropProts <- TestProteinDropThresholds(perProtNA=perSmplNA_cross, quantTbl=quantTable_crossSectional[,metaCrossSectional$Sample_ID], idxVal=2)
perSmplNACross_dropProts_serumQCOnly <- TestProteinDropThresholds(perProtNA=perSmplNA_cross_serumQC, quantTbl=quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idxVal=2)

# Enumerate total proteins at each threshold
totalNProtsCross <- do.call("rbind",lapply(unique(perProtNACross_dropProts$PERCENTAGE), function(perc) 
                                    data.frame(PERCENTAGE=perc, FREQ=nrow(perProtNACross_dropProts[perProtNACross_dropProts$PERCENTAGE == perc,]))))
totalNProtsCross_serumQC <- do.call("rbind",lapply(unique(perProtNACross_dropProts_serumQCOnly$PERCENTAGE), function(perc) 
                                            data.frame(PERCENTAGE=perc, FREQ=nrow(perProtNACross_dropProts_serumQCOnly[perProtNACross_dropProts_serumQCOnly$PERCENTAGE == perc,]))))

```

### Add to metadata

```{r}

# Combine
perSmplNACross_dropProts <- merge(perSmplNACross_dropProts, metaCrossSectional, by.x="ID", by.y="Sample_ID")
perSmplNACross_dropProts_serumQCOnly <- merge(perSmplNACross_dropProts_serumQCOnly, metaCrossSectional, by.x="ID", by.y="Sample_ID")

```

### Plot

NOTE: Not filtering proteins separately by plate.

```{r, fig.width=3.5, fig.height=2}

# Cross-sectional
source(here::here("utilities", "R", "codeModules", "plotting", "plotMissingness_crossSectional.R"))

# Save
ggsave(here::here(resCrossDiscoDir, "eda", 
                  "propMissingValues_missingValueThresholds.png"),
       boxPlt,
       units="px", width=4000, height=4000)
ggsave(here::here(resCrossDiscoDir, "eda", 
                  "propMissingValues_missingValueThresholds_byPlate.png"),
       boxPltByPlate,
       units="px", width=10000, height=3000)

ggsave(here::here(resCrossDiscoDir, "eda", 
                  "propMissingValues_missingValueThresholds_justSerumQC.png"),
       boxPlt_serumQCOnly,
       units="px", width=4000, height=4000)
ggsave(here::here(resCrossDiscoDir, "eda", 
                  "propMissingValues_missingValueThresholds_justSerumQC_byPlate.png"),
       boxPltByPlate_serumQCOnly,
       units="px", width=10000, height=3000)

ggsave(here::here(resCrossDiscoDir, "eda", 
                  "nProteins_missingValueThresholds_allProts.png"),
       barPlt,
       units="px", width=4000, height=4000)
ggsave(here::here(resCrossDiscoDir, "eda", 
                  "nProteins_missingValueThresholds_serumQCProts.png"),
       barPlt_serumQCOnly,
       units="px", width=4000, height=4000)

# Add to list
newList <- list(boxPlt,
                boxPlt_serumQCOnly,
                barPlt,
                barPlt_serumQCOnly)
names(newList) <- c("PropMissingValsUnFilt",
                    "PropMissingValsSerumQCProts",
                    "NProtsAllProts",
                    "NProtsSerumQCProts")
plotList <- append(plotList,
                   newList)

```

### Checking for drift

```{r, fig.height=5, fig.width=12}

# Tall format
rownames(quantTable_crossSectional_serumQC) <- quantTable_crossSectional_serumQC$PG.ProteinGroups
rownames(quantTable_crossSectional_instrQC) <- quantTable_crossSectional_instrQC$PG.ProteinGroups
quantTable_crossSectional_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_crossSectional_serumQC[serumProtsCross,]), 
                                                                    quantTable_crossSectional_serumQC[serumProtsCross,2:ncol(quantTable_crossSectional_serumQC)]), id.vars=c("ID"))
quantTable_crossSectional_instrQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_crossSectional_instrQC[serumProtsCross,]), 
                                                                    quantTable_crossSectional_instrQC[serumProtsCross,2:ncol(quantTable_crossSectional_instrQC)]), id.vars=c("ID"))

# Get sample IDs

# Add metadata
quantTable_crossSectional_serumQC_tall["PLATE"] <- "P1"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC05","QC06","QC07","QC08")),]["PLATE"] <- "P2"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC09","QC10","QC11","QC12")),]["PLATE"] <- "P3"
quantTable_crossSectional_instrQC_tall["PLATE"] <- gsub("late", "", gsub("SerumInstrument.*", "", quantTable_crossSectional_instrQC_tall$variable))
quantTable_crossSectional_instrQC_tall["variable"] <- gsub(".*SerumInstrument", "", quantTable_crossSectional_instrQC_tall$variable)

# Order
quantTable_crossSectional_serumQC_tall <- quantTable_crossSectional_serumQC_tall[order(quantTable_crossSectional_serumQC_tall$variable, decreasing=FALSE),]
quantTable_crossSectional_serumQC_tall$variable <- factor(quantTable_crossSectional_serumQC_tall$variable, levels=unique(quantTable_crossSectional_serumQC_tall$variable))

quantTable_crossSectional_instrQC_tall <- quantTable_crossSectional_instrQC_tall[order(quantTable_crossSectional_instrQC_tall$variable, decreasing=FALSE),]
quantTable_crossSectional_instrQC_tall$variable <- factor(quantTable_crossSectional_instrQC_tall$variable, levels=unique(quantTable_crossSectional_instrQC_tall$variable))

# Plot
p1 <- ggplot(quantTable_crossSectional_serumQC_tall, aes(x=variable, 
                                                 y=log2(value),
                                                 group=variable)) + 
  xlab("Serum QC Sample") + ylab("Log2 Protein Abundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  geom_violin(fill="#99E6FF",
              width=0.75,
              alpha=0.75) + 
  geom_boxplot(outlier.size=4,
               width=0.25) + 
  ggtitle("Serum QC (Filtered)") +
  xlab("Serum QC Sample") + ylab("Log2 Protein\nAbundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=72)) 
p2 <- ggplot(quantTable_crossSectional_instrQC_tall, aes(x=variable, 
                                                 y=log2(value),
                                                 group=variable)) + 
  ggtitle("Instrument QC (Filtered)") +
  xlab("Instrument QC Sample") + ylab("Log2 Protein Abundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  geom_violin(fill="#99E6FF",
              width=0.75,
              alpha=0.75) + 
  geom_boxplot(outlier.size=4,
               width=0.25) + 
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=72)) 

# Save
ggsave(here::here(resCrossDiscoDir, "eda",
                  "crossSectional_serumQCSamples_proteinAbundance_nonMissingInAllSerumQC.png"),
       p1,
       units="px", width=9000, height=3500)
ggsave(here::here(resCrossDiscoDir, "eda",
                  "crossSectional_instrumentQCSamples_proteinAbundance_nonMissingInAllSerumQC.png"),
       p2,
       units="px", width=9000, height=3500)


# Add to list
newList <- list(p1, p2)
names(newList) <- c("SerumQCFiltProtAbundanceSerumQC",
                    "SerumQCFiltProtAbundanceInstrQC")
plotList <- append(plotList,
                   newList)

```

### Correlate N missing with abundance

```{r}

# Correlate abundance and number of missing values
MeanFreqMissing_cross <- data.frame(MeanAbundance=rowMeans(quantTable_crossSectional[,metaCrossSectional$Sample_ID], na.rm=TRUE),
                                    FreqMissing=apply(quantTable_crossSectional[,metaCrossSectional$Sample_ID], 
                                                      1, function(row) length(row[is.na(row)])))

# Plot
## Correlate cross-sectional
stats::cor(MeanFreqMissing_cross[MeanFreqMissing_cross$FreqMissing > 0,]$MeanAbundance, 
    MeanFreqMissing_cross[MeanFreqMissing_cross$FreqMissing > 0,]$FreqMissing, method="spearman", use="complete.obs")

# Plot
ggplot(MeanFreqMissing_cross[MeanFreqMissing_cross$FreqMissing > 0,], aes(x=log2(MeanAbundance), y=FreqMissing)) + 
  geom_point() + 
  xlab("Log2 Mean Protein Abundance") + ylab("Frequency of Missing") +
  theme_classic(base_size=16) + theme(axis.title=element_text(size=24))

```

### Investigate samples with high numbers of missing proteins

Define this as \> 20%

```{r}

perSmplNACross_dropProts[perSmplNACross_dropProts$PROP_MISSING > maxPropMissing & perSmplNACross_dropProts$PERCENTAGE == "20%",]

```

### Test for enrichment of missing values

#### Binarise to missing or not

```{r}

# Binarise data 
rownames(quantTable_crossSectional) <- quantTable_crossSectional$PG.ProteinGroups
quantTableCross_naBinarised_unFilt <- GenerateBinaryMissingValTable(quantTable_crossSectional[,metaCrossSectional$Sample_ID], metaCrossSectional)

# Combined
forTestMetaCross_unFilt <- merge(metaCrossSectional, t(quantTableCross_naBinarised_unFilt[,metaCrossSectional$Sample_ID]), by.x="Sample_ID", by.y="row.names")
rownames(forTestMetaCross_unFilt) <- forTestMetaCross_unFilt$Sample_ID

```

#### Test for protein-wise differences in missing values

##### Filtered samples

```{r}

# Test for associations between missingness and factors
## Cross-Sectional
allFactorCross <- TestFactors(metaData=forTestMetaCross_unFilt, factorCols=c("Age","Sex","Condition","PLATE","Site"), ids=rownames(quantTableCross_naBinarised_unFilt))
allFactorChild <- TestFactors(metaData=merge(metaCrossSectional[metaCrossSectional$Age_Group == "Child",], data.frame(t(quantTableCross_naBinarised_unFilt)), 
                              by.x="Sample_ID", by.y="row.names"), factorCols=c("Age","Sex","Condition","PLATE","Site"), ids=rownames(quantTableCross_naBinarised_unFilt))

# Sig filter
missSig_protsAll <- allFactorCross[allFactorCross$PADJ < 0.05 & allFactorCross$Factor == "Condition",]
missSig_protsChild <- allFactorChild[allFactorChild$PADJ < 0.05 & allFactorChild$Factor == "Condition",]

# Save
dir.create(here::here(resCrossDiscoDir, "eda", "missingnessSigTests"), showWarnings=FALSE, recursive=TRUE)
write.table(allFactorCross, here::here(resCrossDiscoDir, "eda", "missingnessSigTests", "binarisedMissingVals_sampleVariableSigTests_allCrossSectionalTogether_unfiltered.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(allFactorChild, here::here(resCrossDiscoDir, "eda", "missingnessSigTests", "binarisedMissingVals_sampleVariableSigTests_crossSectionalChildOnly_unfiltered.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Filter quantification sample

```{r}

# Retain only serum QC proteins
quantTable_crossSectional <- quantTable_crossSectional[serumProtsCross,]
dim(quantTable_crossSectional)

# Drop proteins with high frequencies of missing values
crossPropProts <- EnumerateNAs(quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idx=1)
quantTable_crossSectional <- quantTable_crossSectional[crossPropProts[crossPropProts$PROP_MISSING < maxPropMissing,]$ID,]
dim(quantTable_crossSectional)

```


```{r, fig.height=3, fig.width=6}

# Define samples with high frequencies of missing values
crossPropSmpls <- EnumerateNAs(quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idx=2)

crossPropSmpls <- merge(crossPropSmpls, metaCrossSectional, by.x="ID", by.y="Sample_ID")

table(crossPropSmpls[crossPropSmpls$PROP_MISSING >= 0.2,]$Condition)

# Plot
ggplot(crossPropSmpls,
       aes(x=ID,
           y=PROP_MISSING,
           fill=Condition)) + 
  geom_bar(stat="identity") +
  scale_fill_manual(values=condCols) +
  facet_grid(cols=vars(Condition),
             scale="free_x", 
             space="free_x") +
  xlab("Sample") + ylab("Proportion of Missing Proteins") +
  ylim(0, 1) +
  geom_hline(yintercept=maxPropMissing, linetype="dashed", color="darkgrey") +
  theme_classic() + 
  theme(axis.text.x=element_blank(),
        legend.position="none")

quantTable_crossSectional <- data.frame(PG.ProteinAccessions=rownames(quantTable_crossSectional), 
                                        quantTable_crossSectional[,c(crossPropSmpls[crossPropSmpls$PROP_MISSING < maxPropMissing,]$ID)])

```


### Subset metadata to match filtered columns

```{r}

# Subset metadata
addMetaCross <- addMetaCross[which(addMetaCross$Sample_ID %in% colnames(quantTable_crossSectional)),]
metaCrossSectional <- metaCrossSectional[which(metaCrossSectional$Sample_ID %in% colnames(quantTable_crossSectional)),]

```

### Filter samples

```{r}

# Check max missingness
if (max(EnumerateNAs(quantTable_crossSectional[,metaCrossSectional$Sample_ID], idx=1)$PROP_MISSING) > maxPropMissing) { warning("Protein Missingness for Cross-Sectional not controlled at desired level") }

# Filter proteins again to make sure proteins are < 20% after removing samples
crossPropProts <- EnumerateNAs(quantTable_crossSectional[serumProtsCross,metaCrossSectional$Sample_ID], idx=1)
length(c(crossPropSmpls[crossPropSmpls$PROP_MISSING < maxPropMissing,]$ID))
quantTable_crossSectional <- data.frame(PG.ProteinAccessions=rownames(quantTable_crossSectional),
                                        quantTable_crossSectional[,c(crossPropSmpls[crossPropSmpls$PROP_MISSING < maxPropMissing,]$ID)])
dim(quantTable_crossSectional)

```

### Check and filter proteins MNAR for Condition

The main idea here is to check that any proteins that may be MNAR with condition, thus possible interesting candidates wrt any one condition, we actually have robust measurements for. To do this, we enumerate the overall and per-condition missingness frequencies and apply filtering thresholds to these individually. Main idea is to set a minimum per-group missingness threshold, then an overall missingness threshold more lax than the one used above. 

```{r}

# Filter proteins by a certain proportion of missing
checkGroupWiseMissing <- function(metaData,
                                  prots,
                                  maxPropMissingPerGroup=0.5,
                                  maxPropMissingOverall=0.3) {

  # Loop over proteins
  checkVec <- unlist(sapply(seq_along(prots), function(i) {
    
    # Subset metadata to get child data
    subData <- metaData[,c(prots[[i]], 
                           "Condition")]
    
    # Get proportion of missing
    freqData <- as.matrix(table(subData))
    propData <- sweep(freqData, 
                      STATS=colSums(freqData), 
                      2, 
                      FUN="/")
    
    # Get if protein is above less stringent thresholds
    if (all(propData[rownames(propData) == "0",] > maxPropMissingPerGroup) & # All groups  at least x% non-missing
            (rowSums(freqData)["0"] / nrow(subData)) < maxPropMissingOverall) { # At most y% missing overall
      
      # Return
      return(prots[[i]])
      
    } else { return(NA) }
    
  }))

  # Return
  return(prots[!is.na(checkVec)])

}

# Check missingness
colnames(forTestMetaCross_unFilt) <- gsub("[;]","_",colnames(forTestMetaCross_unFilt))
stringentMissingCheck_cross <- checkGroupWiseMissing(metaData=forTestMetaCross_unFilt,
                                                     prots=setdiff(missSig_protsAll$ID, 
                                                                   rownames(quantTable_crossSectional)),
                                                     maxPropMissingPerGroup=maxPropMissingPerGroup,
                                                     maxPropMissingOverall=maxPropMissingOverall)
stringentMissingCheck_child <- checkGroupWiseMissing(metaData=forTestMetaCross_unFilt[forTestMetaCross_unFilt$Age_Group == "Child",],
                                                     prots=setdiff(missSig_protsChild$ID, 
                                                                   rownames(quantTable_crossSectional)),
                                                     maxPropMissingPerGroup=maxPropMissingPerGroup,
                                                     maxPropMissingOverall=maxPropMissingOverall)

# Check for any
stringentMissingCheck_cross
stringentMissingCheck_child

```

### Save

```{r}

# Create output
dir.create(here::here(diaDir, "filtered"), showWarnings=FALSE, recursive=TRUE)

# Save cross-Sectional
write.table(quantTable_crossSectional, here::here(diaDir, "filtered", paste0("crossSectional_quantTable.filtBy_serumQCProts_maxMissing",maxPropMissing,".tsv")), sep="\t", quote=FALSE, row.names=FALSE)

```

### Checking for drift

```{r, fig.height=5, fig.width=12}

# Tall format
rownames(quantTable_crossSectional_serumQC) <- quantTable_crossSectional_serumQC$PG.ProteinGroups
quantTable_crossSectional_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_crossSectional_serumQC[rownames(quantTable_crossSectional),]), 
                                                                    quantTable_crossSectional_serumQC[rownames(quantTable_crossSectional),2:ncol(quantTable_crossSectional_serumQC)]), id.vars=c("ID"))

# Get sample IDs

# Add metadata
quantTable_crossSectional_serumQC_tall["PLATE"] <- "P1"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC05","QC06","QC07","QC08")),]["PLATE"] <- "P2"
quantTable_crossSectional_serumQC_tall[which(quantTable_crossSectional_serumQC_tall$variable %in% c("QC09","QC10","QC11","QC12")),]["PLATE"] <- "P3"

# Order
quantTable_crossSectional_serumQC_tall <- quantTable_crossSectional_serumQC_tall[order(quantTable_crossSectional_serumQC_tall$variable, decreasing=FALSE),]
quantTable_crossSectional_serumQC_tall$variable <- factor(quantTable_crossSectional_serumQC_tall$variable, levels=unique(quantTable_crossSectional_serumQC_tall$variable))

# Plot
ggplot(quantTable_crossSectional_serumQC_tall, aes(x=variable, y=log2(value))) + 
  geom_boxplot() + ggtitle("Serum QC (Cross-Sectional)") +
  xlab("Serum QC Sample") + ylab("Log2 Protein Abundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=24)

```

### Cluster missing values

#### Preprocess and define colours

```{r}

# Prepare matrices
quantTableCross_naZeros <- quantTable_crossSectional[,metaCrossSectional$Sample_ID]

# Set non-zero to 1
quantTableCross_naZeros[!is.na(quantTableCross_naZeros)] <- 1

# Set NA to 0
quantTableCross_naZeros[is.na(quantTableCross_naZeros)] <- 0

# Tidy names
colnames(quantTableCross_naZeros) <- gsub(" Osteo","\nOsteo",colnames(quantTableCross_naZeros))

# Harmonize data with metadata
quantTableCross_naZeros <- quantTableCross_naZeros[,metaCrossSectional$Sample_ID]

# Define shared color
metaCrossSectional$Age[is.na(metaCrossSectional$Age)] <- 0
mainColFun = colorRamp2(c(0, 1), c("white", "mediumblue"))
ageColFun_cross  = colorRamp2(c(0,min(metaCrossSectional$Age, na.rm=T), max(metaCrossSectional$Age, na.rm=T)), c("darkgrey","white","darkred"))

```

#### Plot as heatmaps

```{r}

source(here::here("utilities", "R", "codeModules", "plotting", "plotMissingHmaps_crossSectional.R"))

```

### View distributions as boxplots

```{r, fig.height=3, fig.width=6}

# Get as tall
rawCrossTall <- melt(data.frame(ID=rownames(quantTable_crossSectional), log2(quantTable_crossSectional[,metaCrossSectional$Sample_ID])), id.vars="ID")

# Merge with
rawCrossTall <- merge(rawCrossTall, metaCrossSectional, by.x="variable", by.y="Sample_ID")

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","Age_Group","PLATE","Ethnicity","Site")) {

  # Plot
  plt <- ggplot(rawCrossTall, aes(x=variable, y=value, fill=.data[[fctr]])) + 
    geom_boxplot() + 
    scale_fill_manual(values=condCols) +
    xlab("Samples") + ylab("Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic() + theme(axis.text.x=element_blank(), 
                       legend.position="top") +
    guides(fill=guide_legend(nrow=1, byRow=TRUE))
  print(plt)
  
} 

```

### Generate Summarize Experiment for DEP

```{r}

# Generate design matrix
designCross_m <- data.frame(label=rownames(metaCrossSectional), Sex=metaCrossSectional$Sex, Age=metaCrossSectional$Age, condition=metaCrossSectional$Condition)
rownames(designCross_m) <- metaCrossSectional$Sample_ID

# Add replicate column
designCross_m["replicate"] <- NA
for (cond in unique(designCross_m$condition)) { designCross_m[designCross_m$condition == cond,]["replicate"] <- 1:nrow(designCross_m[designCross_m$condition == cond,]) }

# Make summarized experiment
uniqNames_quantTableCross  <- make_unique(quantTable_crossSectional[,c("PG.ProteinAccessions","PG.ProteinAccessions",metaCrossSectional$Sample_ID)], 
                                          names="PG.ProteinAccessions", ids="PG.ProteinAccessions", delim=";")
seCross <- make_se(uniqNames_quantTableCross, which(colnames(uniqNames_quantTableCross) %in% metaCrossSectional$Sample_ID), designCross_m)

# Rough fix I think
assay(seCross)[is.nan(assay(seCross))] <- NA

```

### Plot frequencies for missing values

```{r, fig.height=3, fig.width=6}

# Plot protein identification overlaps
plot_frequency(seCross) + theme(axis.text=element_blank())

# Plot protein identification overlaps (stacked bars)
plot_coverage(seCross) + theme(legend.position="none")

# Plot proteins per samples
names(condCols) <- gsub(" ",".",names(condCols))
plot_numbers(seCross) + scale_fill_manual(values=condCols) + 
  theme(axis.text.x=element_blank(), legend.position="none") 

```

### Scale and normalise with VSN

```{r}

# Normalise
## VSN
normCross <- normalize_vsn(seCross)

# Plot SD
meanSdPlot(normCross)

# Check mix/min for VSN - some strange things here with some parameterisations
min(assay(normCross), na.rm=T)

```

### View distributions as boxplots

```{r}

# Get as tall
normCrossTall <- melt(data.frame(ID=rownames(assay(normCross)), assay(normCross)), id.vars="ID")
seCrossTall <- melt(data.frame(ID=rownames(assay(seCross)), assay(seCross)), id.vars="ID")

# Merge with column data
normCrossTall <- merge(normCrossTall, data.frame(colData(normCross)), by.x="variable", by.y="ID")
seCrossTall <- merge(seCrossTall, data.frame(colData(seCross)), by.x="variable", by.y="ID")

# Merge with
normCrossTall <- merge(normCrossTall[,c("variable","ID","label","value")], metaCrossSectional, by.x="label", by.y="Sample_ID")
seCrossTall <- merge(seCrossTall[,c("variable","ID","label","value")], metaCrossSectional, by.x="label", by.y="Sample_ID")

```

### View distributions as boxplots

```{r}

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","Age_Group","PLATE","Ethnicity","Site")) {

  # Plot
  plt <- ggplot(normCrossTall, aes(x=variable, y=value, fill=.data[[fctr]])) + 
    geom_boxplot() + 
    xlab("Samples") + ylab("Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic() + theme(axis.text.x=element_blank(), 
                       legend.position="top") +
    guides(fill=guide_legend(nrow=1, byRow=TRUE))
  print(plt)
  
}  

```

### Plot sample distributions

```{r}

# Get number of conditions per site
freqCondPerSite <- melt(data.frame(table(metaCrossSectional[metaCrossSectional$Age_Group == "Child",][,c("Condition","Site")])))

# Get proportion of conditions per site
freqCondPerSite["Prop"] <- NA
sites <- unique(freqCondPerSite$Site)
for (i in seq_len(length(sites))) { freqCondPerSite[freqCondPerSite$Site == sites[[i]],]["Prop"] <- freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value / sum(freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value) }

# Plot freq
plt1 <- ggplot(freqCondPerSite,
       aes(x=Site, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of\nSamples") +
  theme_classic(base_size=36) + theme(axis.title=element_text(size=88),
                                 axis.text=element_text(size=42),
                                 legend.text=element_text(size=42)) +
  guides(fill=guide_legend(override.aes=list(size=12)))

# Plot freq
plt2 <- ggplot(freqCondPerSite,
       aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportio of\nSamples") +
  theme_classic(base_size=36) + theme(axis.title=element_text(size=88),
                                 axis.text=element_text(size=42),
                                 legend.text=element_text(size=42)) +
  guides(fill=guide_legend(override.aes=list(size=12)))

# Plot
plt <- plt1 + plt2 +
         plot_layout(nrow=1,
                     guides = "collect",
                     axis_titles = "collect")

# save
ggsave(here::here(resCrossDiscoDir, "eda", 
                  paste0("sampleGroupNumbers_bySite_afterFiltration.png")),
       plt, width=10000, height=4000, units="px")

```

```{r}

# Get number of conditions per site
freqCondPerSite <- melt(data.frame(table(addMetaCross_child[addMetaCross_child$Age_Group == "Child" & addMetaCross_child$Condition != "JIA" & !is.na(addMetaCross_child$Date.of.sample),][,c("Condition","Site")])))

# Get proportion of conditions per site
freqCondPerSite["Prop"] <- NA
sites <- unique(freqCondPerSite$Site)
for (i in seq_len(length(sites))) { freqCondPerSite[freqCondPerSite$Site == sites[[i]],]["Prop"] <- freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value / sum(freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value) }

# Plot freq
ggplot(freqCondPerSite,
       aes(x=Site, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of Samples") +
  theme_classic()

# Plot freq
ggplot(freqCondPerSite,
       aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportion of Samples (in Site)") +
  theme_classic()

```

```{r}

# Get number of conditions per Plate
freqCondPerPlate <- melt(data.frame(table(metaCrossSectional[metaCrossSectional$Age_Group == "Child" & metaCrossSectional$Condition != "JIA",][,c("Condition","PLATE")])))

# Get proportion of conditions per Plate
freqCondPerPlate["Prop"] <- NA
Plates <- unique(freqCondPerPlate$PLATE)
for (i in seq_len(length(Plates))) { freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]["Prop"] <- freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]$value / sum(freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]$value) }

# Plot freq
ggplot(freqCondPerPlate,
       aes(x=PLATE, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of Samples") +
  theme_classic()

# Plot prob
ggplot(freqCondPerPlate,
       aes(x=PLATE, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportion of Samples (in Plate)") +
  theme_classic()

# Plot freq 
ggplot(aggregate(value ~ Condition + variable, data=freqCondPerPlate[,c("Condition","value","variable")], FUN=sum),
       aes(x=variable, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of Samples") +
  xlab("") + 
  theme_classic() + theme(axis.text.x=element_blank())


```

### Save filtered metadata

```{r}

# Save metadata
write.table(metaCrossSectional, here::here(diaDir, "filtered", "preSchool_crossSectional_tallFormatClean_samplesFiltered_byMissingness.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(addMetaCross[which(addMetaCross$Sample_ID %in% metaCrossSectional$Sample_ID),], 
            here::here(diaDir, "filtered", "preSchool_additionalMetadata_commonToCrossSectional_samplesFiltered_byMissingness.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

# Combine Leiden CNO with child cross-sectional data
metaCrossSectional["Patient_ID"] <- metaCrossSectional$Sample_ID
metaCrossSectional["Start_Age"] <- metaCrossSectional$Age

```

### Save filtered proteins

```{r}

# Save proteisn
saveRDS(rownames(quantTable_crossSectional[,metaCrossSectional$Sample_ID]), 
        here::here(diaDir, "filtered", paste0("filteredProteinSet_crossSectional_maxMissing",maxPropMissing,".rds")))

# Save quant table
saveRDS(quantTable_crossSectional[,metaCrossSectional$Sample_ID], 
        here::here(diaDir, "filtered", paste0("filtered_unNormalised_crossSectional_maxMissing",maxPropMissing,".rds")))
write.table(quantTable_crossSectional[,metaCrossSectional$Sample_ID], 
            here::here(diaDir, "filtered", paste0("filtered_unNormalised_crossSectional_maxMissing",maxPropMissing,".tsv")), sep="\t", quote=FALSE)

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "02_dataProcessQC_figs.rds"))

```




