---
title: "Data Normalisation"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Load data

```{r}

# Metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Abundance tables
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_quantTables.R"))

# QC tables
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_serumQCTables.R"))

quantTable_crossSectional

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

## Main

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "02_dataProcessQC_figs.rds"))
 
```

### Read in filtered by 20% missingness

```{r}

# Read in filtered protein list
protsCross <- readRDS(here::here(diaDir, "filtered", paste0("filteredProteinSet_crossSectional_maxMissing",maxPropMissing,".rds")))

# Subset data
rownames(quantTable_crossSectional) <- quantTable_crossSectional$PG.ProteinGroups
quantTable_crossSectional <- quantTable_crossSectional[protsCross,c("PG.ProteinGroups",metaCrossSectional$Sample_ID)]

dim(quantTable_crossSectional)

```

### Combine with QC samples

```{r}

# Combine data
quantTable_crossWithQCs <- cbind(quantTable_crossSectional_serumQC[rownames(quantTable_crossSectional),-1], 
                                 quantTable_crossSectional_instrQC[rownames(quantTable_crossSectional),-1], 
                                 quantTable_crossSectional)

# Get QC only
quantTable_crossOnlyQCs <- quantTable_crossWithQCs[,grepl("QC",colnames(quantTable_crossWithQCs))]

```

### NormalyzerDE

#### Format metadata for NormalyzerDE

```{r}

# Sort metadata
forNormDEMeta <- rename(metaCrossSectional[colnames(quantTable_crossSectional)[-1],c("Sample_ID","Condition","Sex","Age","PLATE")], 
                         c("Sample_ID"="sample", "Condition"="group", "PLATE"="Plate"))

## Subset metadata
forNormDEMeta <- rename(metaCrossSectional[colnames(quantTable_crossSectional)[-1],c("Sample_ID","Condition")], 
                         c("Sample_ID"="sample", "Condition"="group"))
forNormDEMeta <- rbind(data.frame(sample=c(colnames(quantTable_crossSectional_serumQC)[-1], 
                                           colnames(quantTable_crossSectional_instrQC)[-1]),
                                  group=c(rep("Serum_QC", ncol(quantTable_crossSectional_serumQC)-1), 
                                          rep("Instrument_QC", ncol(quantTable_crossSectional_instrQC)-1))),
                                  forNormDEMeta)

## Save metadata
write.table(forNormDEMeta, 
            here::here(metaDir, "preSchool_crossSectional_groupIsCondition_forNormalyzerDE.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

# Read in QC ID converstion table
idConvert <- read.table(here::here(metaDir, "serumQCs_idConversionMetadata.tsv"), sep="\t", header=TRUE)

# 
## Subset metadata
forNormDEMeta <- rename(metaCrossSectional[colnames(quantTable_crossSectional)[-1],c("Sample_ID","PLATE")], 
                         c("Sample_ID"="sample", "PLATE"="group"))
forNormDEMetaQC <- rename(idConvert[idConvert$DATASET == "CrossSectional",c("META_ID","PLATE")], c("META_ID"="sample", "PLATE"="group"))
forNormDEMeta <- rbind(forNormDEMeta, forNormDEMetaQC, 
                       data.frame(sample=colnames(quantTable_crossWithQCs)[grepl("Inst",colnames(quantTable_crossWithQCs))], group="PInstr"))

## Save metadata
write.table(forNormDEMeta, 
            here::here(metaDir, "preSchool_crossSectional_groupIsPlate_forNormalyzerDE.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

# Save abundance data
write.table(cbind(PG.ProteinAccessions=rownames(quantTable_crossWithQCs), quantTable_crossWithQCs[,forNormDEMeta$sample]), 
            here::here(diaDir, "filtered", 
                       paste0("crossSectional_quantTable.filtBy_serumQCProts_maxMissing",maxPropMissing,".withSerumQCSamples.forNormalyzerDE.tsv")), 
            sep="\t", quote=FALSE, row.names=FALSE)

```

#### Run Normalyzer

Based on: https://www.bioconductor.org/packages/release/bioc/vignettes/NormalyzerDE/inst/doc/vignette.html

```{r}

# Create object for NormalyzerDE
experimentObj_byCond  <- setupRawDataObject(designPath=here::here(metaDir, "preSchool_crossSectional_groupIsCondition_forNormalyzerDE.tsv"), 
                                            dataPath=here::here(diaDir, "filtered", 
                                                                paste0("crossSectional_quantTable.filtBy_serumQCProts_maxMissing",maxPropMissing,".withSerumQCSamples.forNormalyzerDE.tsv")))
experimentObj_byPlate <- setupRawDataObject(designPath=here::here(metaDir, "preSchool_crossSectional_groupIsPlate_forNormalyzerDE.tsv"), 
                                            dataPath=here::here(diaDir, "filtered", 
                                                                paste0("crossSectional_quantTable.filtBy_serumQCProts_maxMissing",maxPropMissing,".withSerumQCSamples.forNormalyzerDE.tsv")))
normObj_byCond <- getVerifiedNormalyzerObject("metaCrossSectional_NormalyzerDE_byCondtion", 
                                              experimentObj_byCond,
                                              noLogTransform=TRUE)
normObj_byPlate <- getVerifiedNormalyzerObject("metaCrossSectional_NormalyzerDE_byPlate", 
                                               experimentObj_byPlate,
                                               noLogTransform=TRUE)

# Normalise
normResults_byCond <- normMethods(normObj_byCond)
normResults_byPlate <- normMethods(normObj_byPlate)

# Generate performance measures
normResultsWithEval_byCond <- analyzeNormalizations(normResults_byCond)
normResultsWithEval_byPlate <- analyzeNormalizations(normResults_byPlate)

# Save results
jobDir_byCond <-  setupJobDir("metaCrossSectional_NormalyzerDE_byCondition", here::here(diaDir, "normalised"))
jobDir_byPlate <-  setupJobDir("metaCrossSectional_NormalyzerDE_byPlate", here::here(diaDir, "normalised"))
writeNormalizedDatasets(normResultsWithEval_byCond, jobDir_byCond)
writeNormalizedDatasets(normResultsWithEval_byPlate, jobDir_byPlate)

# Plots
generatePlots(normResultsWithEval_byCond, jobDir_byCond)
generatePlots(normResultsWithEval_byPlate, jobDir_byPlate)

```

### Read in normalised data

```{r}

# List files
normFiles <- list.files(path=jobDir_byCond, pattern="*.txt")
normMs <- lapply(normFiles, function(file) read.table(here::here(jobDir_byCond, file), sep="\t", header=TRUE, row.names=1))
names(normMs) <- gsub("-.*|.*_|[.]txt","",normFiles)

```

### Check if normalised data with LOESS is unsupervised

Cyclic LOESS may be spliting by group-data to normalise which may be problematic for downstream modelling. Check this by manually normalising nad commparing normalised matrices. Have to round for slight variations in precision of normaliastion.

```{r}

# Check
cycLoessNorm <- limma::normalizeCyclicLoess(log2(quantTable_crossWithQCs[,forNormDEMeta$sample]), method="fast")

# Get whether LOESS is unsupervisde
checkLoessUnSupervised <- all(round(as.matrix(cycLoessNorm[,colnames(normMs$CycLoess)]), 8) == round(as.matrix(as.matrix(normMs$CycLoess)), 8), na.rm=TRUE)
if (!checkLoessUnSupervised) { warning("LOESS APPEARS TO NOT BE UNSUPERVISED, IT MIGHT BE CONSIDERING GROUP DATA")}

```

### Save

```{r}

# Save
write.table(data.frame(ProteinGroup=rownames(normMs$CycLoess), normMs$CycLoess[,!grepl("QC",colnames(normMs$CycLoess))]), 
            here::here(diaDir, "normalised", "crossSectional_quantTable_normalisedCyclicLoess.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Process normalised data

```{r}

# Make tall
normTall <- lapply(names(normMs), function(method) data.frame(reshape2::melt(data.frame(ProtID=rownames(normMs[[method]]), normMs[[method]]), id.vars="ProtID"), Method=method))
names(normTall) <- names(normMs)
  
# Combine
normTall <- do.call("rbind",normTall)

# Retain only QC samples
normTall_qcOnly <- normTall[grepl("QC",normTall$variable),]

```

### View normalised distributions as boxplots

```{r, fig.height=3, fig.width=6}

# Get as tall
rawCrossTall <- melt(data.frame(ID=rownames(quantTable_crossSectional), 
                                log2(quantTable_crossSectional[,metaCrossSectional$Sample_ID])), id.vars="ID")

# Merge with
rawCrossTall <- merge(rawCrossTall, metaCrossSectional, by.x="variable", by.y="Sample_ID")

# Get list of colours
allColors <- list(Sex=sexCols,
                  Condition=condCols,
                  Site=siteCols,
                  PLATE=plateCols)

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","PLATE","Site")) {

  # Plot
  p1 <- ggplot(rawCrossTall[rawCrossTall$Age_Group == "Child",], 
               aes(x=variable, 
                   y=value, 
                   fill=.data[[fctr]])) + 
    geom_boxplot(outlier.size=2.5) + 
    scale_fill_manual(values=allColors[[fctr]]) +
    xlab("Samples") + ylab("Log2 Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic(base_size=48) + 
    theme(axis.text.x=element_blank(), 
          axis.ticks.x=element_blank(),
          axis.title=element_text(size=72),
          strip.text=element_blank(),
          legend.text=element_text(size=28),
          legend.title=element_blank(),
          legend.position="top") +
    guides(fill=guide_legend(nrow=1, 
                             byRow=TRUE, 
                             override.aes=list(size=16)))
  
  # Save
  ggsave(here::here(diaDir, "normalised", 
                    paste0("sampleAbundanceDistribution_splitBy",fctr,"_rawAbundances_childSamples.png")),
         p1,
         units="px", width=5000, height=3500)

  if (fctr == "Condition") {
    
    # Add to list
    newList <- list(p1)
    names(newList) <- c("sampleAbundDistribution_colByCondition_raw")
    plotList <- append(plotList,
                       newList)
    
  }
    
}

```

### Plot

```{r}

# Split by Serum/instrument QC
normTall_qcOnly["QCType"] <- "Serum QC"
normTall_qcOnly[grepl("Instrument",normTall_qcOnly$variable),]["QCType"] <- "Instrument QC"

# Split into serum/instrument QC
serumQC_norm <- normTall_qcOnly[normTall_qcOnly$QCType == "Serum QC",]
instrQC_norm <- normTall_qcOnly[normTall_qcOnly$QCType == "Instrument QC",]

# Retain only proteins that had missing values to impute missing values
serumQC_norm_qcOnly <- serumQC_norm[which(serumQC_norm$ProtID %in% 
                                                    rownames(quantTable_crossOnlyQCs[,!grepl("Instrument",colnames(quantTable_crossOnlyQCs))])),]
instrQC_norm_qcOnly <- instrQC_norm[which(instrQC_norm$ProtID %in% 
                                                    rownames(quantTable_crossOnlyQCs[, grepl("Instrument",colnames(quantTable_crossOnlyQCs))])),]
  
# Plot
serumQC_plt <- ggplot(serumQC_norm_qcOnly, aes(x=variable, y=log2(value))) + 
         geom_boxplot(outlier.size=2.5) + ggtitle("Serum QC Samples Only") +
         xlab("QC Sample") + ylab("Log2 Protein Abundance") +
         facet_wrap(~ Method, ncol=4) +
         theme_classic(base_size=16) + 
         theme(axis.text=element_blank(),
               axis.title=element_text(size=48),
               plot.title=element_text(size=64),
               strip.text=element_text(size=32))
instrQC_plt <- ggplot(instrQC_norm_qcOnly, aes(x=variable, y=log2(value))) + 
         geom_boxplot(outlier.size=2.5) + ggtitle("Instrument QC Samples Only") +
         xlab("QC Sample") + ylab("Log2 Protein Abundance") +
         facet_wrap(~ Method, ncol=4) +
         theme_classic(base_size=16) + 
         theme(axis.text=element_blank(),
               axis.title=element_text(size=48),
               plot.title=element_text(size=64),
               strip.text=element_text(size=32))

# Save
dir.create(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "QC"), showWarnings=FALSE, recursive=TRUE)
ggsave(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "QC", 
                  "normalisationMethod_boxplotDistCheck_serumQCOnly_allProteins.png"), serumQC_plt, units="px", width=10000, height=6000)
ggsave(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "QC", 
                  "normalisationMethod_boxplotDistCheck_instrumentQCOnly_allProteins.png"), instrQC_plt, units="px", width=10000, height=6000)


# Retain only proteins that had missing values to impute missing values
serumQC_norm_qcOnly <- serumQC_norm[which(serumQC_norm$ProtID %in% 
                                                    rownames(quantTable_crossOnlyQCs[is.na(rowSums(quantTable_crossOnlyQCs)),!grepl("Instrument",colnames(quantTable_crossOnlyQCs))])),]
instrQC_norm_qcOnly <- instrQC_norm[which(instrQC_norm$ProtID %in% 
                                                    rownames(quantTable_crossOnlyQCs[is.na(rowSums(quantTable_crossOnlyQCs)), grepl("Instrument",colnames(quantTable_crossOnlyQCs))])),]
  
# Plot
serumQC_plt <- ggplot(serumQC_norm_qcOnly, aes(x=variable, y=log2(value))) + 
         geom_boxplot(outlier.size=2.5) + ggtitle("Serum QC Samples Only") +
         xlab("QC Sample") + ylab("Log2 Protein Abundance") +
         facet_wrap(~ Method, ncol=4) +
         theme_classic(base_size=16) + 
         theme(axis.text=element_blank(),
               axis.title=element_text(size=48),
               plot.title=element_text(size=64),
               strip.text=element_text(size=32))
instrQC_plt <- ggplot(instrQC_norm_qcOnly, aes(x=variable, y=log2(value))) + 
         geom_boxplot(outlier.size=2.5) + ggtitle("Instrument QC Samples Only") +
         xlab("QC Sample") + ylab("Log2 Protein Abundance") +
         facet_wrap(~ Method, ncol=4) +
         theme_classic(base_size=16) + 
         theme(axis.text=element_blank(),
               axis.title=element_text(size=48),
               plot.title=element_text(size=64),
               strip.text=element_text(size=32))

# Save
ggsave(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "QC", 
                  "normalisationMethod_boxplotDistCheck_serumQCOnly_protsWithNAOnly.png"), serumQC_plt, units="px", width=10000, height=6000)
ggsave(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "QC", 
                  "normalisationMethod_boxplotDistCheck_instrumentQCOnly_protsWithNAOnly.png"), instrQC_plt, units="px", width=10000, height=6000)

```

#### PCA to investigate

```{r}

# Edit Colours
tidyCondCols <- condCols
names(tidyCondCols) <- gsub("Infectious.","Inf. ",names(tidyCondCols))

# Make column categorical
metaCrossSectional$COLUMN <- as.character(metaCrossSectional$COLUMN)

# Fix infectious osteomyelitis
metaCrossSectional$Condition <- gsub("Infectious ","Inf. ",metaCrossSectional$Condition)

```

#### Prepare data

```{r}

# Extract normalised data
normData <- normMs$CycLoess

# Drop QCs
normData <- normData[,!grepl("QC",colnames(normData))]

# Drop NAs
normData_noNA <- normData[!is.na(rowSums(normData)),]
dim(normData_noNA)

```

### View normalised distributions as boxplots

```{r, fig.height=3, fig.width=6}

# Get as tall
normCrossTall <- melt(data.frame(ID=rownames(normMs$CycLoess), 
                                normMs$CycLoess[,metaCrossSectional$Sample_ID]), id.vars="ID")

# Merge with
normCrossTall <- merge(normCrossTall, metaCrossSectional, by.x="variable", by.y="Sample_ID")

# Get list of colours
allColors <- list(Sex=sexCols,
                  Condition=condCols,
                  Site=siteCols,
                  PLATE=plateCols)


# Loop over factors and color/facet
for (fctr in c("Sex","Condition","PLATE","Site")) {

  # Plot
  p1 <- ggplot(normCrossTall[normCrossTall$Age_Group == "Child",], 
               aes(x=variable, 
                   y=value, 
                   fill=.data[[fctr]])) + 
    geom_boxplot(outlier.size=2.5) + 
    scale_fill_manual(values=allColors[[fctr]]) +
    xlab("Samples") + ylab("Log2 Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic(base_size=48) + 
    theme(axis.text.x=element_blank(), 
          axis.ticks.x=element_blank(),
          axis.title=element_text(size=72),
          strip.text=element_blank(),
          legend.text=element_text(size=28),
          legend.title=element_blank(),
          legend.position="top") +
    guides(fill=guide_legend(nrow=1, 
                             byRow=TRUE, 
                             override.aes=list(size=16)))
  
  # Save
  ggsave(here::here(diaDir, "normalised", 
                    paste0("sampleAbundanceDistribution_splitBy",fctr,"_cycLoessNormAbundances_childSamples.png")),
         p1,
         units="px", width=5000, height=3500)

  if (fctr == "Condition") {
    
    # Add to list
    newList <- list(p1)
    names(newList) <- c("sampleAbundDistribution_colByCondition_cyclicLoessNorm")
    plotList <- append(plotList,
                       newList)
    
  }
  
}

```

#### All Samples

```{r}

# Run PCA
crossPCA <- prcomp(t(normData_noNA[!is.na(rowSums(normData_noNA)),metaCrossSectional$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
crossPCA_metaData <- merge(metaCrossSectional[,!grepl("^PC",colnames(metaCrossSectional))], crossPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Child Samples

```{r}

# Subset metadata
metaCrossSectional_child <- metaCrossSectional[metaCrossSectional$Age_Group == "Child",]

# Run PCA
crossChildPCA <- prcomp(t(normData_noNA[!is.na(rowSums(normData_noNA)),metaCrossSectional_child$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
crossChildPCA_metaData <- merge(metaCrossSectional_child[,!grepl("^PC",colnames(metaCrossSectional_child))], crossChildPCA$x, by.x="Sample_ID", by.y="row.names")

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "03_dataNormalisation_figs.rds"))

```




