---
title: "Exploratory Data Analysis"
author: "Euan McDonnell"
format: html
editor: source
---

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Install here
if (!require("here")) { install.packages("here") }

# Source metadata
here::i_am("04_eda.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Load data

```{r}

# Load metadatas
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Abundance tables
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_quantTables.R"))

# Read in chosen normalised data
normQuantTable_cross <- read.table(here::here(diaDir, "normalised", "crossSectional_quantTable_normalisedCyclicLoess.tsv"),
                             sep="\t", header=TRUE, row.names=1)

# Drop QC samples
normQuantTable_cross <- normQuantTable_cross[,!grepl("QC", colnames(normQuantTable_cross))]
dim(normQuantTable_cross)

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "03_dataNormalisation_figs.rds"))
 
```

## Main code

### PCA

Do this for proteins with no missing values and compare to probabilistic PCA.

#### Prep

```{r}

# Edit Colours
tidyCondCols <- condCols
names(tidyCondCols) <- gsub("Infectious.","Inf. ",names(tidyCondCols))

# Make column categorical
metaCrossSectional$COLUMN <- as.character(metaCrossSectional$COLUMN)

# Fix infectious osteomyelitis
metaCrossSectional$Condition <- gsub("Infectious ","Inf. ",metaCrossSectional$Condition)

```

#### All Samples

```{r}

# Run PCA
crossPCA <- prcomp(t(normQuantTable_cross[!is.na(rowSums(normQuantTable_cross)),]), center=TRUE, scale=TRUE)

# Merge with metadata
crossPCA_metaData <- merge(metaCrossSectional[,!grepl("^PC",colnames(metaCrossSectional))], crossPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Child Samples

```{r}

# Run PCA
crossChildPCA <- prcomp(t(normQuantTable_cross[!is.na(rowSums(normQuantTable_cross)), metaCrossSectional[metaCrossSectional$Age_Group == "Child",]$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
crossChildPCA_metaData <- merge(metaCrossSectional[,!grepl("^PC",colnames(metaCrossSectional))], crossChildPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Plot all

```{r}

# Keep the code in a sourced R script for tidyness
source(here::here("utilities", "R", "codeModules", "plotting", "plotPCA_crossSectional.R"))

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=12) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
plate1_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(1, 2), pch_size=12) + 
    scale_fill_manual(values=plateCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
plate2_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(2, 3), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
plate3_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(3, 4), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
plate4_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(4, 5), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt,
                plate1_plt,
                plate2_plt,
                plate3_plt,
                plate4_plt)
names(newList) <- c("normAbundancePCA_colBySite",
                    "normAbundancePCA_colByCondition",
                    "normAbundancePCA_colByPlatePC12",
                    "normAbundancePCA_colByPlatePC23",
                    "normAbundancePCA_colByPlatePC34",
                    "normAbundancePCA_colByPlatePC45")
plotList <- append(plotList,
                   newList)

```

### Probabilistic PCA

Base off of: https://www.bioconductor.org/packages/release/bioc/vignettes/pcaMethods/inst/doc/pcaMethods.pdf

quantTable_crossSectional \<- data.frame(PG.ProteinAccessions=rownames(quantTable_crossSectional), quantTable_crossSectional\[,c(crossPropSmpls\[crossPropSmpls$PROP_MISSING < maxPropMissing,]$ID)\]) quantTable_longitudinal \<- data.frame(PG.ProteinAccessions=rownames(quantTable_longitudinal), quantTable_longitudinal\[,c(longPropSmpls\[longPropSmpls$PROP_MISSING < maxPropMissing,]$ID)\])

#### Cross-Sectional

##### All-together

```{r}

# Fit probabilistic PCA
set.seed(35353)
resPPCA <- pcaMethods::pca(t(normQuantTable_cross[,metaCrossSectional$Sample_ID]), 
                           method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
resPPCA_data <- merge(metaCrossSectional, pcaMethods::scores(resPPCA), by.x="Sample_ID", "row.names")

# Get variance explained
varExpl <- list(variance.percent=(sDev(resPPCA) / sum(sDev(resPPCA)))  * 100)

```

##### Child-only

```{r}

# Fit probabilistic PCA
set.seed(35353)
resPPCA_child <- pcaMethods::pca(t(normQuantTable_cross[,metaCrossSectional[metaCrossSectional$Age_Group == "Child",]$Sample_ID]), 
                                 method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
resPPCA_childData <- merge(metaCrossSectional, scores(resPPCA_child), by.x="Sample_ID", "row.names")

# Get variance explained
varExpl_child <- list(variance.percent=(sDev(resPPCA_child) / sum(sDev(resPPCA_child)))  * 100)

# Set factors as categorical/numeric
resPPCA_childData$COLUMN <- as.character(resPPCA_childData$COLUMN)

```

##### Adult-only

```{r}

# If adult
if (!excludeAdultSamples) {
  
  # Fit probabilistic PCA
  set.seed(35353)
  resPPCA_adult <- pcaMethods::pca(t(normQuantTable_cross[,metaCrossSectional[metaCrossSectional$Age_Group == "Adult",]$Sample_ID]), 
                                   method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)
  
  # Combine with metadata
  resPPCA_adultData <- merge(metaCrossSectional, scores(resPPCA_adult), by.x="Sample_ID", "row.names")
  
  # Get variance explained
  varExpl_adult <- list(variance.percent=(sDev(resPPCA_adult) / sum(sDev(resPPCA_adult)))  * 100)
  
  # Set factors as categorical/numeric
  resPPCA_adultData$COLUMN <- as.character(resPPCA_adultData$COLUMN)

}  

```

##### Plot

```{r}

# Keep the code in a sourced R script for tidyness
source(here::here("utilities", "R", "codeModules", "plotting", "plotProbPCA_crossSectional.R"))

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=9) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=9) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
plate_plt <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="PLATE", label_id="Sample_ID", pcs=c(1, 2), pch_size=12) + 
    scale_fill_manual(values=plateCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt1 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(1, 2), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt2 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(2, 3), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt3 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(3, 4), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt4 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(4, 5), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt5 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(5, 6), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
condSite_plt6 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="CondSite", label_id="Sample_ID", pcs=c(6, 7), pch_size=9) + 
    scale_fill_manual(values=condSiteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
site_plt1 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=9) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
site_plt2 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="Site", label_id="Sample_ID", pcs=c(2, 3), pch_size=9) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
site_plt3 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="Site", label_id="Sample_ID", pcs=c(3, 4), pch_size=9) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
site_plt4 <- Plot2DPCA(meta=resPPCA_childData, var=varExpl_child, color_by="Site", label_id="Sample_ID", pcs=c(4, 5), pch_size=9) + 
    scale_fill_manual(values=siteCols) +
    theme_minimal(base_size=48) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=64),
          legend.text=element_text(size=56),
          axis.title=element_text(size=64),
          plot.title=element_text(size=72)) +
    labs(fill="Plate") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt,
                plate_plt,
                condSite_plt1,
                condSite_plt2,
                condSite_plt3,
                condSite_plt4,
                condSite_plt5,
                condSite_plt6,
                site_plt1,
                site_plt2,
                site_plt3,
                site_plt4)
names(newList) <- c("normAbundancePPCA_colBySite",
                    "normAbundancePPCA_colByCondition",
                    "normAbundancePPCA_colByPlate",
                    "normAbundancePPCA_colByCondSitePC12",
                    "normAbundancePPCA_colByCondSitePC23",
                    "normAbundancePPCA_colByCondSitePC34",
                    "normAbundancePPCA_colByCondSitePC45",
                    "normAbundancePPCA_colByCondSitePC67",
                    "normAbundancePPCA_colByCondSitePC78",
                    "normAbundancePPCA_colBySitePC12",
                    "normAbundancePPCA_colBySitePC23",
                    "normAbundancePPCA_colBySitePC34",
                    "normAbundancePPCA_colBySitePC45")
plotList <- append(plotList,
                   newList)

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "04_eda_figs.rds"))

```




