---
title: "Data Imputation"
format: revealjs
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source parameters, functions and install packages
here::i_am("05_dataImputation.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "04_eda_figs.rds"))
 
```


### Source functions for imputation

# Source specific functions for 
source(here::here("utilities", "pancreatic_cancer_biomarkers", "code", "Imput_funcs.r")) # trKNN

### Load data

```{r}

# Metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Abundance tables
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_quantTables.R"))

# QC tables
source(here::here("utilities", "R", "codeModules", "data", "unFiltered_serumQCTables.R"))

# Abundance tables (filtered)
# source(here::here("utilities", "R", "codeModules", "data", "filtered_quantTables.R"))

# Read in cyclic Loess-normalised data
crossSectional_normQuantTable <- read.table(here::here(diaDir, "normalised", "metaCrossSectional_NormalyzerDE_byCondition", "CycLoess-normalized.txt"), sep="\t", header=TRUE, row.names=1)

```

## Main

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

Below is adapted from code by Dr Emily Johnson: https://github.com/CBFLivUni/pancreatic_cancer_biomarkers/blob/HedrichCNO/06_imputation_optimisation.qmd

### Filter by 20% missingness

```{r}

# Filter by missingess
source(here::here("utilities", "R", "codeModules", "data", "filterByMissingness.R"))

```

### Prep data for imputation

```{r}

# Subset metadata
imputeMetaData <- metaCrossSectional[colnames(quantTable_crossSectional)[-1],c("Sample_ID","Condition")]
imputeMetaData <- rbind(data.frame(Sample_ID=c(colnames(quantTable_crossSectional_serumQC)[-1], 
                                               colnames(quantTable_crossSectional_instrQC)[-1]),
                                   Condition=c(rep("Serum_QC", ncol(quantTable_crossSectional_serumQC)-1), 
                                               rep("Instrument_QC", ncol(quantTable_crossSectional_instrQC)-1))),
                                   imputeMetaData)

# Rename to data_impute to fit emily's code
data_impute <- crossSectional_normQuantTable[,imputeMetaData$Sample_ID]

# Get QC-only
data_impute_QCOnly <- data_impute[,grepl("QC",colnames(data_impute))]

# Retain only proteins with non-NAs
imputDataCross_noNA <- crossSectional_normQuantTable[!is.na(rowSums(crossSectional_normQuantTable)),metaCrossSectional$Sample_ID]
imputDataCross_onlyMissing <- crossSectional_normQuantTable[is.na(rowSums(crossSectional_normQuantTable)),metaCrossSectional$Sample_ID]
dim(imputDataCross_noNA)

```

### Imputation with QC samples

#### Perform imputation

```{r}

# Toggle whether to read in if file exists
impFileName <- here::here(impDir, "impMethodsTrialled.rds")
if (!file.exists(impFileName) | reGenerateFiles) {

  # Run imputation
  impAllData   <- suppressWarnings(RunAllImputationMethods(data_impute, imputeMetaData, group="Condition"))
  saveRDS(impAllData, impFileName)

# Else read in from file
} else {
  
  impAllData <- readRDS(impFileName)
  
}

# Drop NULLs
impAllData <- impAllData[!sapply(impAllData, is.null)]
  
# Make tall
imputDataTall <- lapply(names(impAllData), function(method) data.frame(reshape2::melt(data.frame(ProtID=rownames(impAllData[[method]]), impAllData[[method]]), id.vars="ProtID"), Method=method))
names(imputDataTall) <- names(impAllData)
  
# Combine
imputDataTall <- do.call("rbind",imputDataTall)

# Retain only QC samples
imputDataTall_qcOnly <- imputDataTall[grepl("QC",imputDataTall$variable),]

```

### Save imputed data

```{r}

# Loop over imputation methods
for (impType in names(impAllData)) {

  # Save
  print(impType)
  write.table(data.frame(Protein=rownames(impAllData[[impType]]), impAllData[[impType]][,!grepl("QC", colnames(impAllData[[impType]]))]),
              here::here(impDir, paste0("quantTable_imputedWith",impType,".tsv")),
              sep="\t", quote=FALSE)
  
}

```

### Perform PCAs

```{r}

# Loop over imputation methods
for (impType in names(impAllData)) {

  # Save
  print(impType)
  subImpMat <- impAllData[[impType]][,metaCrossSectional$Sample_ID]
  
  # Run PCA
  crossChildPCA <- prcomp(t(subImpMat[!is.na(rowSums(subImpMat)), metaCrossSectional[metaCrossSectional$Age_Group == "Child",]$Sample_ID]), center=TRUE, scale=TRUE)
  
  # Merge with metadata
  crossChildPCA_metaData <- merge(metaCrossSectional[,!grepl("^PC",colnames(metaCrossSectional))], crossChildPCA$x, by.x="Sample_ID", by.y="row.names")
  
  # Loop over PCs
  ## All together
  for (i in 1:4) {
    
    # Plot
    # Technical
    plate_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
      scale_fill_manual(values=plateCols) +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      labs(fill="Plate") +
      ggtitle("Child Samples") +
      guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    row_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      labs(fill="Row") +
      ggtitle("Child Samples") +
      guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    col_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      labs(fill="Column") +
      ggtitle("Child Samples") +
      guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    
    ## Biological
    cond_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
      scale_fill_manual(values=tidyCondCols) +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      labs(fill="Condition") +
      ggtitle("Child Samples") +
      guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    sex_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
      scale_fill_manual(values=sexCols) +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      labs(fill="Sex") +
      ggtitle("Child Samples") +
      guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    age_plt <- Plot2DPCA(meta=crossChildPCA_metaData, var=get_eigenvalue(crossChildPCA), color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
      scale_fill_continuous(high="darkred", low="white") +
      theme_minimal(base_size=24) +
      theme(legend.position="right", 
            legend.direction="vertical",
            legend.title=element_text(size=32),
            legend.text=element_text(size=24),
            axis.title=element_text(size=42),
            plot.title=element_text(size=64)) +
      ggtitle("Child Samples") +
      guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
    
    # Save
    dir.create(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType),
               recursive=TRUE)
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colByPlate.crossSectional_childSamples.png")), plate_plt, width=6000, height=4000, units="px")
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colByPlateRow.crossSectional_childSamples.png")), row_plt, width=6000, height=4000, units="px")
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colByPlateColumn.crossSectional_childSamples.png")), col_plt, width=6000, height=4000, units="px")
    
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colByCondition.crossSectional_childSamples.png")), cond_plt, width=6000, height=4000, units="px")
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colBySex.crossSectional_childSamples.png")), sex_plt, width=6000, height=4000, units="px")
    ggsave(here::here(resCrossDiscoDir, "eda", "PCA", "childSamples", "imputed", impType, paste0("pcaBiplot.pc",i,"to",i+1,".colByAge.crossSectional_childSamples.png")), age_plt, width=6000, height=4000, units="px")
    
  }
  
}

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "05_dataImputation_figs.rds"))

```




