---
title: "Model Evalutaion"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Ensure correction location
if (!require(here)) { install.packages("here") }
here::i_am("09_modelEvaluation.qmd")

# Source parameters, functions and install packages
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

# Source grid search parameters
source(here::here("params", "biomarkerParams_gridSearch.R"))

```

### Load data

#### Abundance data

```{r}

# Read in cyclic Loess-normalised data
quantTable_allSamples <- read.table(here::here(diaDir, "normalised", 
                                    "metaCrossSectional_NormalyzerDE_byCondition", "CycLoess-normalized.txt"), sep="\t", header=TRUE, row.names=1)
# Match metadata
trainMetaData <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "biomarkers",
                                       "allSamplesMetadata_takenForSelection.tsv"), sep="\t", header=TRUE)

# Chosen imputed data
quantTable_chosenMethod <- read.table(here::here(impDir, paste0("quantTable_imputedWithbPCA.tsv")), sep="\t", header=TRUE, row.names=1)
quantTable_chosenMethodScaled <- t(apply(quantTable_chosenMethod[,-1], 1, function(row) scale(row, center=TRUE, scale=TRUE)))
colnames(quantTable_chosenMethodScaled) <- colnames(quantTable_chosenMethod)[-1]

# Get common
quantTable <- quantTable_allSamples[,trainMetaData$Sample_ID]

# Keep only selected metadata
quantTable <- quantTable[,!grepl("QC",colnames(quantTable))]
quantTable_noMissing <- quantTable[!is.na(rowSums(quantTable)),]

# raed in differential abundance data
diffData <- read.table(here::here(resCrossDiscoDir, "trainSplitProp0.8", "sampleRandom", "differentialAbundance",
                                  "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData_justDresdenSamples.tsv"),
                       sep="\t",
                       header=TRUE)

# Read in nomrlaised, unimputed data
quantTableNorm <- read.table(here::here(diaDir, "normalised", 
                                        "crossSectional_quantTable_normalisedCyclicLoess.tsv"),
                             sep="\t",
                             header=TRUE,
                             row.names=1)

# Take sig
diffDataSig <- diffData[diffData$adj.P.Val < candidatePThr,]

```

#### Metadata

```{r}

# Metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Get relevant samples
metaCrossSectional_child <- metaCrossSectional[metaCrossSectional$Age_Group == "Child",]
metaCrossSectional_noJIA <- metaCrossSectional_child[!metaCrossSectional_child$Condition %in% c("JIA"),]

# Get test data
testMetaData <- metaCrossSectional_noJIA[!metaCrossSectional_noJIA$Sample_ID %in% trainMetaData$Sample_ID,]

# Add years
addMetaCross_child["Years"] <- ( max(as.numeric(ymd(addMetaCross_child$Date.of.sample)), na.rm=TRUE) - as.numeric(ymd(addMetaCross_child$Date.of.sample)) ) / 365 

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

## Main

### List all relevant models

```{r}

# Get possible files
grpModelFiles <- list.files(pattern = "mg_lassoRFModel.rds", 
                            recursive = TRUE)
bimModelFiles <- list.files(pattern = "bm_lassoRFModel.rds", 
                            recursive = TRUE)

# Keep run name
grpModelFiles <- grpModelFiles[grepl(runName,grpModelFiles)]
bimModelFiles <- bimModelFiles[grepl(runName,bimModelFiles)]
grpModelFiles <- grpModelFiles[!grepl("old",grpModelFiles)]
bimModelFiles <- bimModelFiles[!grepl("old",bimModelFiles)]

grpModelFiles <- grpModelFiles[grepl("lfc0.1|lfc0.2|lfc0.3|lfc0.4|lfc0.5", grpModelFiles)]
bimModelFiles <- bimModelFiles[grepl("lfc0.1|lfc0.2|lfc0.3|lfc0.4|lfc0.5", bimModelFiles)]

grpModelFiles <- grpModelFiles[grepl("DresdenOnly", grpModelFiles)]
bimModelFiles <- bimModelFiles[grepl("DresdenOnly", bimModelFiles)]

length(grpModelFiles)
length(bimModelFiles)

```

### Load models

```{r}

# Loop over models 
grpModels <- list()
bimModels <- list()
for (modelFilename in grpModelFiles) {
  
  # Append models
  grpModels <- rlist::list.append(grpModels, readRDS(here::here(modelFilename)))
  
}
for (modelFilename in bimModelFiles) {
  
  # Append models
  bimModels <- rlist::list.append(bimModels, readRDS(here::here(modelFilename)))
  
}

gc()


```

### Add names

```{r}

# Get parameters
grpModelAttrib <- lapply(grpModels, function(mod) mod[[1]]$Parameterisation)
bimModelAttrib <- lapply(bimModels, function(mod) mod[[1]]$Parameterisation)

# Add names to models
names(grpModels) <- sapply(grpModelAttrib, paste, collapse="_")
names(bimModels) <- sapply(bimModelAttrib, paste, collapse="_")

```

### Get numbers of proteins

```{r}

multiGrp_modelProts <- data.frame()
biModal_modelProts <- data.frame()

# Loop over models
for (i in seq_len(length(grpModels))) {
  
  # Get model details
  grpModelProts <- data.frame(ModelID=names(grpModels[i]),
                              SelectedProteins=grpModels[[i]][[1]]$varSel)
  bimModelProts <- data.frame(ModelID=names(bimModels[i]),
                              SelectedProteins=bimModels[[i]][[1]]$varSel)
  
  # Combine to output
  multiGrp_modelProts <- rbind(multiGrp_modelProts,
                               grpModelProts)
  biModal_modelProts  <- rbind(biModal_modelProts,
                               bimModelProts)
  
}

# Get parameters
multiGrp_modelProts <- cbind(multiGrp_modelProts, stringr::str_split_fixed(multiGrp_modelProts$ModelID, pattern="_", n=Inf))
biModal_modelProts  <- cbind(biModal_modelProts,  stringr::str_split_fixed(biModal_modelProts$ModelID, pattern="_", n=Inf))
colnames(multiGrp_modelProts) <- c("ModelID","SelectedProteins","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                   "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")
colnames(biModal_modelProts)  <- c("ModelID","SelectedProteins","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                   "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")

# FreqProts
multiGrp_protFreq <- aggregate(SelectedProteins ~ ModelID, data=multiGrp_modelProts[,1:2], FUN=length)
biModal_protFreq  <- aggregate(SelectedProteins ~ ModelID,  data=biModal_modelProts[,1:2], FUN=length)

# Get parameters
multiGrp_protFreq <- cbind(multiGrp_protFreq, stringr::str_split_fixed(multiGrp_protFreq$ModelID, pattern="_", n=Inf))
biModal_protFreq  <- cbind(biModal_protFreq,  stringr::str_split_fixed(biModal_protFreq$ModelID, pattern="_", n=Inf))
colnames(multiGrp_protFreq) <- c("ModelID","Frequency","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                   "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")
colnames(biModal_protFreq)  <- c("ModelID","Frequency","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                   "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")

```

### Extract model MCCs

```{r}

# Define outputs
multiGrp_modelMCC <- data.frame()
multiGrp_allModelMCC <- data.frame()
biModal_modelMCC  <- data.frame()
biModal_allModelMCC <- data.frame()

# Loop over models
for (i in seq_len(length(grpModels))) {
  
  # Get MCC
  mccsGrp <- sapply(seq_len(length(grpModels[[i]])), function(j) grpModels[[i]][[j]]$rfMCC)
  mccsBim <- sapply(seq_len(length(bimModels[[i]])), function(j) bimModels[[i]][[j]]$rfMCC)
  
  # Get coefficient of variation
  mccGrpCoefVar <- sd(mccsGrp, na.rm=TRUE) / mean(mccsGrp, na.rm=TRUE)
  mccBimCoefVar <- sd(mccsBim, na.rm=TRUE) / mean(mccsBim, na.rm=TRUE)
  
  # Get model details
  grpModelMCC <- data.frame(ModelID=names(grpModels[i]),
                            ModelFeatures=grpModels[[i]][[1]]$Parameterisation,
                            Mean=mean(mccsGrp, na.rm=TRUE),
                            SD=sd(mccsGrp, na.rm=TRUE),
                            CoeffVar=mccGrpCoefVar)
  bimModelMCC <- data.frame(ModelID=names(bimModels[i]),
                            ModelFeatures=bimModels[[i]][[1]]$Parameterisation,
                            Mean=mean(mccsBim, na.rm=TRUE),
                            SD=sd(mccsBim, na.rm=TRUE),
                            CoeffVar=mccBimCoefVar)
  
  # Combine to output
  multiGrp_modelMCC <- rbind(multiGrp_modelMCC,
                             grpModelMCC)
  biModal_modelMCC <- rbind(biModal_modelMCC,
                             bimModelMCC)
  multiGrp_allModelMCC <- rbind(multiGrp_allModelMCC,
                                data.frame(ModelID=names(grpModels[i]), 
                                           Split=seq_len(length(mccsGrp)),
                                           MCC=mccsGrp))
  biModal_allModelMCC <- rbind(biModal_allModelMCC,
                                data.frame(ModelID=names(bimModels[i]), 
                                           Split=seq_len(length(mccsBim)),
                                           MCC=mccsBim))
  
}

### Get model params for MCC
multiGrp_allModelMCC <- cbind(multiGrp_allModelMCC, stringr::str_split_fixed(multiGrp_allModelMCC$ModelID, pattern="_", n=Inf))
biModal_allModelMCC <- cbind(biModal_allModelMCC, stringr::str_split_fixed(biModal_allModelMCC$ModelID, pattern="_", n=Inf))
colnames(multiGrp_allModelMCC) <- c("ModelID","Split","MCC","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                    "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")
colnames(biModal_allModelMCC) <- c("ModelID","Split","MCC","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                    "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")

# Sort factor levels
multiGrp_allModelMCC$Log2Thr <- factor(multiGrp_allModelMCC$Log2Thr, levels=c("0","0.1","0.2","0.3","0.4","0.5"))
multiGrp_allModelMCC$PThr <- factor(multiGrp_allModelMCC$PThr, levels=c("0.1","0.05","0.01"))
biModal_allModelMCC$Log2Thr <- factor(biModal_allModelMCC$Log2Thr, levels=c("0","0.1","0.2","0.3","0.4","0.5"))
biModal_allModelMCC$PThr <- factor(multiGrp_allModelMCC$PThr, levels=c("0.1","0.05","0.01"))

```

### Check MCC and Group Frequency correlation

```{r}

multiGrpTrain_corRes <- data.frame()
multiGrpTest_corRes  <- data.frame()
bimTrain_corRes <- data.frame()
bimTest_corRes  <- data.frame()

# Define all conitions
allConditions <- unique(metaCrossSectional_child$Condition)

getCondFreq <- function(model,
                        conds,
                        dataSet,
                        bimodalModel=FALSE) {
  
    # Get list of frequencies
    out <- lapply(seq_len(length(model)), function(j) {
      
      # If running a bimodal model, group non-CNO conditions
      if (bimodalModel) {
        
        freq <- table(ifelse(model[[j]]$SampleConditions[[dataSet]] == conds[[1]], conds[[1]], conds[[2]]))

      # Else consider individual 
      } else {
      
        # Enumerate frequencies
        freq <- unlist(table(model[[j]]$SampleConditions[[dataSet]]))
        
      }
      
      # return
      return(freq[conds])
      
    })
    
    # Return
    return(out)
    
}

# Loop over models
for (i in seq_len(length(grpModels))) {
  
  # Get MCCs
  mccsGrp <- sapply(seq_len(length(grpModels[[i]])), function(j) grpModels[[i]][[j]]$rfMCC)
  mccsBim <- sapply(seq_len(length(bimModels[[i]])), function(j) bimModels[[i]][[j]]$rfMCC)
  
  # Get condiiton frequency
  grpFrq_grpTrain <- do.call("rbind", getCondFreq(model=grpModels[[i]], conds=allConditions, dataSet="Train"))
  grpFrq_grpTest  <- do.call("rbind", getCondFreq(model=grpModels[[i]], conds=allConditions, dataSet="Test"))
  grpFrq_bimTrain <- do.call("rbind", getCondFreq(model=bimModels[[i]], conds=c("CNO", "Other"), dataSet="Train", bimodalMode=TRUE))
  grpFrq_bimTest  <- do.call("rbind", getCondFreq(model=bimModels[[i]], conds=c("CNO", "Other"), dataSet="Train", bimodalModel=TRUE))
  
  # Correlate
  corRes_grpTrain <- cor(scale(mccsGrp, center=TRUE, scale=TRUE), apply(grpFrq_grpTrain, 2, scale, center=TRUE, scale=TRUE))
  corRes_grpTest <- cor(scale(mccsGrp, center=TRUE, scale=TRUE), apply(grpFrq_grpTest, 2, scale, center=TRUE, scale=TRUE))
  
  corRes_bimTrain <- cor(scale(mccsBim, center=TRUE, scale=TRUE), apply(grpFrq_bimTrain, 2, scale, center=TRUE, scale=TRUE))
  corRes_bimTest <- cor(scale(mccsBim, center=TRUE, scale=TRUE), apply(grpFrq_bimTest, 2, scale, center=TRUE, scale=TRUE))
  
  # Melt
  corRes_grpTrainTall <- reshape2::melt(cbind(Group=rownames(grpFrq_grpTrain), grpFrq_grpTrain), id.vars="Group")
  corRes_grpTestTall <- reshape2::melt(cbind(Group=rownames(grpFrq_grpTest), grpFrq_grpTest), id.vars="Group")

  # Generate output
  multiGrpTrain_corRes <- rbind(multiGrpTrain_corRes,
                                data.frame(ModelID=names(grpModels[i]), 
                                           data.frame(t(corRes_grpTrain))))
  multiGrpTest_corRes  <- rbind(multiGrpTest_corRes,
                                data.frame(ModelID=names(grpModels[i]), 
                                           data.frame(t(corRes_grpTest))))
  bimTrain_corRes <- rbind(bimTrain_corRes,
                                data.frame(ModelID=names(bimModels[i]), 
                                           data.frame(t(corRes_bimTrain))))
  bimTest_corRes  <- rbind(bimTest_corRes,
                                data.frame(ModelID=names(bimModels[i]), 
                                           data.frame(t(corRes_bimTest))))
  
}

### Add group
multiGrpTrain_corRes <- cbind(multiGrpTrain_corRes, "Group"=gsub("[0-9]*$","",rownames(multiGrpTrain_corRes)))
bimTrain_corRes <- cbind(bimTrain_corRes, "Group"=gsub("[0-9]*$","",rownames(bimTrain_corRes)))

### Get model params for MCC
multiGrpTrain_corRes <- cbind(multiGrpTrain_corRes, stringr::str_split_fixed(multiGrpTrain_corRes$ModelID, pattern="_", n=Inf))
bimTrain_corRes <- cbind(bimTrain_corRes, stringr::str_split_fixed(bimTrain_corRes$ModelID, pattern="_", n=Inf))
colnames(multiGrpTrain_corRes) <- c("ModelID","Correlation","Group","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                    "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")
colnames(bimTrain_corRes) <- c("ModelID","Correlation","Group","ImputationMethod","BatchAdjusted","FeatureTypesSelected","CommonProteins",
                                    "SigFilter","FreqRuns","FreqSplit","PThr","Log2Thr","DEMethod","ProtSet")

# Sort factor levels
multiGrpTrain_corRes$Log2Thr <- factor(multiGrpTrain_corRes$Log2Thr, levels=c("0","0.1","0.2","0.5"))
multiGrpTrain_corRes$PThr <- factor(multiGrpTrain_corRes$PThr, levels=c("0.1","0.05","0.01"))
bimTrain_corRes$Log2Thr <- factor(bimTrain_corRes$Log2Thr, levels=c("0","0.1","0.2","0.5"))
bimTrain_corRes$PThr <- factor(bimTrain_corRes$PThr, levels=c("0.1","0.05","0.01"))

```

### Plot correlation of MCC vs group frequencies

```{r}

# Get groups to loop over
groups <- unique(multiGrpTrain_corRes$Group)

```

### Get overlap matrix

```{r}

# Get proteins
selectProts_grpModel <- lapply(seq_len(length(grpModels)), function(i) grpModels[[i]][[1]]$varSel)
selectProts_bimModel <- lapply(seq_len(length(bimModels)), function(i) bimModels[[i]][[1]]$varSel)

# Function to compute the number of common elements
jsiCommonProts <- function(x, y) {
  length(intersect(x, y)) / length(union(x, y))
}
intersectCommonProts <- function(x, y) { length(intersect(x, y)) }
    
# Generate the symmetrical matrix using outer and sapply
jsiProtM_grp <- outer(selectProts_grpModel, selectProts_grpModel, Vectorize(jsiCommonProts))
jsiProtM_bim <- outer(selectProts_bimModel, selectProts_bimModel, Vectorize(jsiCommonProts))
rownames(jsiProtM_grp) <- names(grpModels)
colnames(jsiProtM_grp) <- names(grpModels)
rownames(jsiProtM_bim) <- names(bimModels)
colnames(jsiProtM_bim) <- names(bimModels)

```

### Get intersections

```{r}

# Generate the symmetrical matrix using outer and sapply
intProtM_grp <- outer(selectProts_grpModel, selectProts_grpModel, Vectorize(intersectCommonProts))
intProtM_bim <- outer(selectProts_bimModel, selectProts_bimModel, Vectorize(intersectCommonProts))
rownames(intProtM_grp) <- names(grpModels)
colnames(intProtM_grp) <- names(grpModels)
rownames(intProtM_bim) <- names(bimModels)
colnames(intProtM_bim) <- names(bimModels)

```

### Plot heatmap

```{r}
    
# Color function
cellColFun <- colorRamp2(c(0, 1), c('white','deepskyblue'))

dir.create(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts", "jsi", "biModal"),
           recursive=TRUE)
dir.create(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts", "jsi", "multiGroup"),
           recursive=TRUE)

bimModel_metaData <- biModal_modelProts[!duplicated(biModal_modelProts$ModelID),]
grpModel_metaData <- biModal_modelProts[!duplicated(biModal_modelProts$ModelID),]

```

### Investigate frequent proteins

```{r}

pThr <- "0.05"

# Relevant columns
relCols <- c("ModelID","SelectedProteins","ImputationMethod","PThr","Log2Thr")

# FreqProts
multiGrp_protFreq <- aggregate(ModelID ~ SelectedProteins + ImputationMethod + Log2Thr + PThr,
                               data=multiGrp_modelProts[, relCols], FUN=length)
biModal_protFreq  <- aggregate(ModelID ~ SelectedProteins + ImputationMethod + Log2Thr + PThr, 
                               data=biModal_modelProts[, relCols], FUN=length)

# Order
multiGrp_protFreq <- multiGrp_protFreq[order(multiGrp_protFreq$ModelID, decreasing=TRUE),]
biModal_protFreq <- biModal_protFreq[order(biModal_protFreq$ModelID, decreasing=TRUE),]

# Set if proteins have missing values
biModal_protFreq["Imputed"] <- "Yes"
biModal_protFreq[which(biModal_protFreq$SelectedProteins %in% rownames(quantTable_noMissing)),]["Imputed"] <- "No"
multiGrp_protFreq["Imputed"] <- "Yes"
multiGrp_protFreq[which(multiGrp_protFreq$SelectedProteins %in% rownames(quantTable_noMissing)),]["Imputed"] <- "No"

```

```{r}

# Generate all possible combinations of factor2, factor3, and factor1
bim_allCombinations <- expand.grid(
  Imputed = unique(biModal_protFreq$Imputed),
  Log2Thr = unique(biModal_protFreq$Log2Thr),
  PThr = unique(biModal_protFreq$PThr),
  SelectedProteins = unique(biModal_protFreq$SelectedProteins),
  ImputationMethod = unique(biModal_protFreq$ImputationMethod)
)
grp_allCombinations <- expand.grid(
  Imputed = unique(multiGrp_protFreq$Imputed),
  Log2Thr = unique(multiGrp_protFreq$Log2Thr),
  PThr = unique(multiGrp_protFreq$PThr),
  SelectedProteins = unique(multiGrp_protFreq$SelectedProteins),
  ImputationMethod = unique(multiGrp_protFreq$ImputationMethod)
)

# Merge original data with all combinations, filling missing values with 0
biModal_protFreqFull <- merge(bim_allCombinations, biModal_protFreq, by = c("Imputed", "Log2Thr", "PThr", "SelectedProteins", "ImputationMethod"), all.x = TRUE)
multiGrp_protFreqFull <- merge(grp_allCombinations, multiGrp_protFreq, by = c("Imputed", "Log2Thr", "PThr", "SelectedProteins", "ImputationMethod"), all.x = TRUE)
biModal_protFreqFull$ModelID[is.na(biModal_protFreqFull$ModelID)] <- 0
multiGrp_protFreqFull$ModelID[is.na(multiGrp_protFreqFull$ModelID)] <- 0

# Make heatmap
bim_mat2Plot <- reshape(
  
  biModal_protFreqFull,
  timevar = "ImputationMethod", # Becomes columns
  idvar = c("Imputed","Log2Thr","PThr","SelectedProteins"), # Becomes rows
  direction = "wide" # Specify wide
  
)
grp_mat2Plot <- reshape(
  
  multiGrp_protFreqFull,
  timevar = "ImputationMethod", # Becomes columns
  idvar = c("Imputed","Log2Thr","PThr","SelectedProteins"), # Becomes rows
  direction = "wide" # Specify wide
  
)
colnames(bim_mat2Plot) <- gsub("ModelID[.]","",colnames(bim_mat2Plot))
colnames(grp_mat2Plot) <- gsub("ModelID[.]","",colnames(grp_mat2Plot))

```

```{r}

# Get all unique selected prorteins
bim_allProts <- unique(biModal_modelProts$SelectedProteins)
grp_allProts <- unique(multiGrp_modelProts$SelectedProteins)

# Split by pvalue threshold
bim_mat2Plot$PThr <- as.character(bim_mat2Plot$PThr)
grp_mat2Plot$PThr <- as.character(grp_mat2Plot$PThr)
bim_mat2PlotWide <- do.call("cbind", lapply(unique(bim_mat2Plot$PThr), function(pThr) {
  
  bim_mat2Plot[bim_mat2Plot$PThr == pThr, imputationMethod]
  
}))
grp_mat2PlotWide <- do.call("cbind", lapply(unique(grp_mat2Plot$PThr), function(pThr) {
  
  grp_mat2Plot[grp_mat2Plot$PThr == pThr, imputationMethod]
  
}))

# Add relevant metadata
bim_rowMetadata <- bim_mat2Plot[!duplicated(bim_mat2Plot[,c("Imputed", "Log2Thr","SelectedProteins")]),c("Imputed", "Log2Thr","SelectedProteins")]
grp_rowMetadata <- grp_mat2Plot[!duplicated(grp_mat2Plot[,c("Imputed", "Log2Thr","SelectedProteins")]),c("Imputed", "Log2Thr","SelectedProteins")]

# Add rownames
rownames(bim_rowMetadata) <- paste(bim_rowMetadata$Imputed, bim_rowMetadata$Log2Thr, bim_rowMetadata$SelectedProteins, sep="_")
rownames(grp_rowMetadata) <- paste(grp_rowMetadata$Imputed, grp_rowMetadata$Log2Thr, grp_rowMetadata$SelectedProteins, sep="_")
rownames(bim_mat2PlotWide) <- rownames(bim_rowMetadata)
rownames(grp_mat2PlotWide) <- rownames(grp_rowMetadata)

# Fix imputed
bim_rowMetadata["Imputed"] <- "Yes"
grp_rowMetadata["Imputed"] <- "Yes"
bim_rowMetadata[which(bim_rowMetadata$SelectedProteins %in% rownames(quantTable_noMissing)),]["Imputed"] <- "No"
grp_rowMetadata[which(grp_rowMetadata$SelectedProteins %in% rownames(quantTable_noMissing)),]["Imputed"] <- "No"

# De-duplicate
bim_rowMetadata <- bim_rowMetadata[!duplicated(bim_rowMetadata),]
grp_rowMetadata <- grp_rowMetadata[!duplicated(grp_rowMetadata),]

# Add rownames
rownames(bim_rowMetadata) <- paste(bim_rowMetadata$Imputed, bim_rowMetadata$Log2Thr, bim_rowMetadata$SelectedProteins, sep="_")
rownames(grp_rowMetadata) <- paste(grp_rowMetadata$Imputed, grp_rowMetadata$Log2Thr, grp_rowMetadata$SelectedProteins, sep="_")

# Filter to remove empty rows
# rowMetadata <- rowMetadata[rowSums(mat2PlotWide) > 0,]
bim_mat2PlotWide <- bim_mat2PlotWide[rownames(bim_rowMetadata),]
grp_mat2PlotWide <- grp_mat2PlotWide[rownames(grp_rowMetadata),]
rownames(bim_mat2PlotWide) <- rownames(bim_rowMetadata)
rownames(grp_mat2PlotWide) <- rownames(grp_rowMetadata)

# Order by row sums
bim_mat2PlotWide <- bim_mat2PlotWide[order(rowSums(bim_mat2PlotWide), decreasing=TRUE),]
grp_mat2PlotWide <- grp_mat2PlotWide[order(rowSums(grp_mat2PlotWide), decreasing=TRUE),]

# Remove with zero
bim_mat2PlotWide <- bim_mat2PlotWide[rowSums(bim_mat2PlotWide) > 0,]
grp_mat2PlotWide <- grp_mat2PlotWide[rowSums(grp_mat2PlotWide) > 0,]

# Re-sort
bim_rowMetadata <- bim_rowMetadata[rownames(bim_mat2PlotWide),]
grp_rowMetadata <- grp_rowMetadata[rownames(grp_mat2PlotWide),]

```

```{r}

filtGrpCandidatesNone <- grp_mat2PlotWide[grepl("No_0.5_",rownames(grp_mat2PlotWide)),
                                          grepl("None", colnames(grp_mat2PlotWide))][,2,drop=FALSE]
filtBimCandidatesNone <- bim_mat2PlotWide[grepl("No_0.5_",rownames(bim_mat2PlotWide)),
                                          grepl("None", colnames(bim_mat2PlotWide))][,2,drop=FALSE]
filtGrpCandidatesNone <- filtGrpCandidatesNone[filtGrpCandidatesNone >= 9,,drop=FALSE]
filtBimCandidatesNone <- filtBimCandidatesNone[filtBimCandidatesNone >= 9,,drop=FALSE]
filtGrpCandidatesNone <- unique(gsub(".*_", "", rownames(filtGrpCandidatesNone)))
filtBimCandidatesNone <- unique(gsub(".*_", "", rownames(filtBimCandidatesNone)))

# Keep any proteins detected at least > 7 times in all models in all except "None"
filtGrpCandidates <- grp_mat2PlotWide[apply(grp_mat2PlotWide[,!grepl("None", colnames(grp_mat2PlotWide))], 1, function(row) all(row >= 7)),]
filtGrpCandidates <- unique(gsub(".*_", "", rownames(filtGrpCandidates)))
filtBimCandidates <- bim_mat2PlotWide[apply(bim_mat2PlotWide[,!grepl("None", colnames(bim_mat2PlotWide))], 1, function(row) all(row >= 7)),]
filtBimCandidates <- unique(gsub(".*_", "", rownames(filtBimCandidates)))

# # Get common
filtGrpCandidates <- union(filtGrpCandidates,
                           filtGrpCandidatesNone)
filtBimCandidates <- union(filtBimCandidates,
                           filtBimCandidatesNone)

# Compare number of proteins
length(filtGrpCandidates)
length(filtBimCandidates)
length(intersect(filtGrpCandidates,
                 filtBimCandidates))

# Get output dataframe
finalTargets <- data.frame(Protein=union(filtBimCandidates, 
                                         filtGrpCandidates))

# Get models
finalTargets["Model"] <- "MultiGroup"
finalTargets[which(finalTargets$Protein %in% filtBimCandidates),]["Model"] <- "BiModal"
finalTargets[which(finalTargets$Protein %in% filtBimCandidates & finalTargets$Protein %in% filtGrpCandidates),]["Model"] <- "Both"
table(finalTargets$Model)

```

### Add correlated proteins

```{r}

# Get top correlated proteins
corrM <- sapply(finalTargets$Protein, function(protID) cor(unlist(quantTableNorm[protID,]), 
                                                           t(quantTableNorm[setdiff(rownames(quantTable_chosenMethod), finalTargets$Protein),]),
                                                           use="pairwise.complete.obs"))
rownames(corrM) <- setdiff(rownames(quantTable_chosenMethod), finalTargets$Protein)

# Loop over candidates
allCors <- vector(mode="list",
                  length=ncol(corrM))
for (i in seq_len(ncol(corrM))) {

  # Sort most correlated and get any with corr > 0.7
  topCorrs <- sort(abs(corrM[,i]),
                   decreasing=TRUE)
  topCorrs <- topCorrs[topCorrs > 0.7]

  # Add to all
  allCors[[i]] <- topCorrs[topCorrs > 0.7]

}
names(allCors) <- colnames(corrM)

# Get all unique
allUniq <- unique(names(unlist(unname(allCors))))

# Loop over unique proteins
protGroups <- vector(mode="list",
                     length=length(allUniq))
names(protGroups) <- allUniq
for (protID in allUniq) {
  
  # Check prots
  checkProts <- sapply(allCors, function(prots) protID %in% names(prots))
  chosenProts <- names(checkProts[checkProts])
  
  # Check if bimodal, group or both
  protGroups[[protID]] <- ifelse(any(chosenProts %in% filtBimCandidates & chosenProts %in% filtGrpCandidates),
                                 "Both",
                                 ifelse(any(chosenProts %in% filtGrpCandidates),
                                        "MultiGroup",
                                        "BiModal"))
  
}

# Get if significant in stringent-filtered set
topDEHits <- intersect(allUniq,
                       diffDataSig[apply(diffDataSig[,c(2,3)], 1, function(row) any(abs(row) > 3)),]$PG_ID)


```

```{r}

# Add correlators
finalTargets <- rbind(finalTargets,
                      data.frame(Protein=topDEHits,
                                 Model=unlist(protGroups[topDEHits])))

```

### Save

```{r}

# Save final targets
write.table(finalTargets,
            here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "biomarkers", "modelComparison",
                       "finalProteinList_biModal_multiGroupModels.tsv"),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)

```

### Plot heatmaps dictating selection

```{r}

bimImp <- bim_rowMetadata[bim_rowMetadata$Imputed == "Yes",]
  hmap1 <- Heatmap(bim_mat2PlotWide[bim_rowMetadata$Imputed == "Yes",],
          
          # Colour
          col = colorRamp2(c(0, max(bim_mat2PlotWide)), c('white','salmon')),
          
          # Row/column splits/clustering
          cluster_columns=FALSE,
          cluster_rows=FALSE,
          row_split=bim_rowMetadata[bim_rowMetadata$Imputed == "Yes",]$Log2Thr,
          column_split=colnames(bim_mat2PlotWide),
          row_title_rot=0,
          row_title_gp = gpar(fontsize=172),
          column_title_gp = gpar(fontsize=172),
          
          # Borders and splits
          border=TRUE,
          row_gap=unit(2.5, "cm"),
          column_gap=unit(2.5, "cm"),
  
          # Row/column names
          row_labels=as.character(bim_rowMetadata[bim_rowMetadata$Imputed == "Yes",]$SelectedProteins),
          column_labels=rep(candidatePThr, length(imputationMethod)),
          row_names_gp=gpar(fontsize=48),
          column_names_gp=gpar(fontsize=84),
          column_names_rot=0,
          column_names_centered=TRUE,
          row_names_side = "left",
          column_names_side = "bottom",
          
          # Add cell functions
          cell_fun = function(j, i, x, y, width, height, fill) {
            
            # If more than zero overlaps
            if (bim_mat2PlotWide[bim_rowMetadata$Imputed == "Yes",][i,j] > 0) {
            
              # Add number of overlapping models
              grid.text(sprintf("%d", bim_mat2PlotWide[bim_rowMetadata$Imputed == "Yes",][i, j]), x, y, gp = gpar(fontsize = 72))
              
            }
            
          },
          
          # Legend
          heatmap_legend_param = list(title="Frequency of Models",
                                          legend_direction="horizontal",
                                          grid_height = unit(45, "mm"),
                                          legend_width = unit(350, "mm"),
                                          labels_gp = gpar(fontsize=72),
                                          title_gp = gpar(fontsize=96,
                                                          fontface="bold"),
                                          border="black")
          
  )
  
bimUnImp <- bim_rowMetadata[bim_rowMetadata$Imputed == "No",]
  hmap2 <- Heatmap(bim_mat2PlotWide[bim_rowMetadata$Imputed == "No",],
          
          # Colour
          col = colorRamp2(c(0, max(bim_mat2PlotWide)), c('white','royalblue1')),
                  
          # Row/column splits/clustering
          cluster_columns=FALSE,
          cluster_rows=FALSE,
          row_split=bim_rowMetadata[bim_rowMetadata$Imputed == "No",]$Log2Thr,
          column_split=colnames(bim_mat2PlotWide),
          row_title_rot=0,
          row_title_gp = gpar(fontsize=172),
          column_title_gp = gpar(fontsize=172),
          
          # Borders and splits
          border=TRUE,
          row_gap=unit(2.5, "cm"),
          column_gap=unit(2.5, "cm"),
  
          # Row/column names
          row_labels=as.character(bim_rowMetadata[bim_rowMetadata$Imputed == "No",]$SelectedProteins),
          column_labels=rep(candidatePThr, length(imputationMethod)),
          row_names_gp=gpar(fontsize=48),
          column_names_gp=gpar(fontsize=84),
          column_names_rot=0,
          column_names_centered=TRUE,
          row_names_side = "left",
          column_names_side = "bottom",
          
          # Add cell functions
          cell_fun = function(j, i, x, y, width, height, fill) {
            
            # If more than zero overlaps
            if (bim_mat2PlotWide[bim_rowMetadata$Imputed == "No",][i,j] > 0) {
            
              # Add number of overlapping models
              grid.text(sprintf("%d", bim_mat2PlotWide[bim_rowMetadata$Imputed == "No",][i, j]), x, y, gp = gpar(fontsize = 64))
              
            }
            
          },
          
          # Legend
          heatmap_legend_param = list(title="Frequency of Models",
                                          legend_direction="horizontal",
                                          grid_height = unit(45, "mm"),
                                          legend_width = unit(350, "mm"),
                                          labels_gp = gpar(fontsize=72),
                                          title_gp = gpar(fontsize=96,
                                                          fontface="bold"),
                                          border="black")
          
  )
  
  # Save
  png(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts",
                 paste0("proteinSelectFreq_overlapBetweenSelectParams_imputedProteins_biModalModel.png")), units="px", width=3500, height=2500)
  draw(hmap1, padding=unit(c(5,5,5,5), "cm"), heatmap_legend_side = "top")
  dev.off()
  png(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts",
                 paste0("proteinSelectFreq_overlapBetweenSelectParams_unimputedProteins_biModalModel.png")), units="px", width=3500, height=3500)
  draw(hmap2, padding=unit(c(5,5,5,5), "cm"), heatmap_legend_side = "top")
  dev.off()

```

```{r}

grpImp <- grp_rowMetadata[grp_rowMetadata$Imputed == "Yes",]
  hmap1 <- Heatmap(grp_mat2PlotWide[grp_rowMetadata$Imputed == "Yes",],
          
          # Colour
          col = colorRamp2(c(0, max(grp_mat2PlotWide)), c('white','salmon')),
          
          # Row/column splits/clustering
          cluster_columns=FALSE,
          cluster_rows=FALSE,
          row_split=grp_rowMetadata[grp_rowMetadata$Imputed == "Yes",]$Log2Thr,
          column_split=colnames(grp_mat2PlotWide),
          row_title_rot=0,
          row_title_gp = gpar(fontsize=172),
          column_title_gp = gpar(fontsize=172),
          
          # Borders and splits
          border=TRUE,
          row_gap=unit(2.5, "cm"),
          column_gap=unit(2.5, "cm"),
  
          # Row/column names
          row_labels=as.character(grp_rowMetadata[grp_rowMetadata$Imputed == "Yes",]$SelectedProteins),
          column_labels=rep(candidatePThr, length(imputationMethod)),
          row_names_gp=gpar(fontsize=48),
          column_names_gp=gpar(fontsize=84),
          column_names_rot=0,
          column_names_centered=TRUE,
          row_names_side = "left",
          column_names_side = "bottom",
          
          # Add cell functions
          cell_fun = function(j, i, x, y, width, height, fill) {
            
            # If more than zero overlaps
            if (grp_mat2PlotWide[grp_rowMetadata$Imputed == "Yes",][i,j] > 0) {
            
              # Add number of overlapping models
              grid.text(sprintf("%d", grp_mat2PlotWide[grp_rowMetadata$Imputed == "Yes",][i, j]), x, y, gp = gpar(fontsize = 64))
              
            }
            
          },
          
          
          # Legend
          heatmap_legend_param = list(title="Frequency of Models",
                                          legend_direction="horizontal",
                                          grid_height = unit(45, "mm"),
                                          legend_width = unit(350, "mm"),
                                          labels_gp = gpar(fontsize=72),
                                          title_gp = gpar(fontsize=96,
                                                          fontface="bold"),
                                          border="black")
          
  )
  
  
grpUnImp <- grp_rowMetadata[grp_rowMetadata$Imputed == "No",]
  hmap2 <- Heatmap(grp_mat2PlotWide[grp_rowMetadata$Imputed == "No",],
          
          # Colour
          col = colorRamp2(c(0, max(grp_mat2PlotWide)), c('white','royalblue1')),
                  
          # Row/column splits/clustering
          cluster_columns=FALSE,
          cluster_rows=FALSE,
          row_split=grp_rowMetadata[grp_rowMetadata$Imputed == "No",]$Log2Thr,
          column_split=colnames(grp_mat2PlotWide),
          row_title_rot=0,
          row_title_gp = gpar(fontsize=172),
          column_title_gp = gpar(fontsize=172),
          
          # Borders and splits
          border=TRUE,
          row_gap=unit(2.5, "cm"),
          column_gap=unit(2.5, "cm"),
  
          # Row/column names
          row_labels=as.character(grp_rowMetadata[grp_rowMetadata$Imputed == "No",]$SelectedProteins),
          column_labels=rep(candidatePThr, length(imputationMethod)),
          row_names_gp=gpar(fontsize=48),
          column_names_gp=gpar(fontsize=84),
          column_names_rot=0,
          column_names_centered=TRUE,
          row_names_side = "left",
          column_names_side = "bottom",
          
          # Add cell functions
          cell_fun = function(j, i, x, y, width, height, fill) {
            
            # If more than zero overlaps
            if (grp_mat2PlotWide[grp_rowMetadata$Imputed == "No",][i,j] > 0) {
            
              # Add number of overlapping models
              grid.text(sprintf("%d", grp_mat2PlotWide[grp_rowMetadata$Imputed == "No",][i, j]), x, y, gp = gpar(fontsize = 48))
              
            }
            
          },
          
          # Legend
          heatmap_legend_param = list(title="Frequency of Models",
                                          legend_direction="horizontal",
                                          grid_height = unit(45, "mm"),
                                          legend_width = unit(350, "mm"),
                                          labels_gp = gpar(fontsize=72),
                                          title_gp = gpar(fontsize=96,
                                                          fontface="bold"),
                                          border="black")
          
  )
    
  # Save
  png(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts",
                 paste0("proteinSelectFreq_overlapBetweenSelectParams_imputedProteins_multiGroupModel.png")), units="px", width=3500, height=6500)
  draw(hmap1, padding=unit(c(5,5,5,5), "cm"), heatmap_legend_side = "top")
  dev.off()
  png(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp),"sampleRandom", "biomarkers", "modelComparison", "overlapProts",
                 paste0("proteinSelectFreq_overlapBetweenSelectParams_unimputedProteins_multiGroupModel.png")), units="px", width=3500, height=6500)
  draw(hmap2, padding=unit(c(5,5,5,5), "cm"), heatmap_legend_side = "top")
  dev.off()

```




