---
title: "Validation Data Processing, Normalisation & QC"
author: "Euan McDonnell"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
here::i_am("17_figurePreparation.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "10_finalModel_figs.rds"))
 
```


## Load data

### Metadata

```{r}

# Source metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata.R"))

# Read in protein annotation
proteinAnnotData_validData <- read.table(here::here(baseDataDir, "validation", "DIAProteomics", "directDIA_spectronautProcessed", "spectronaut", "20250409_PreSchool_2_20250312_ValidationData", 
                                                    "2022_0105d_PreSchool_2_20250312_ValidationData_Report_BGS Analysis Grid View Report (Pivot).tsv"), sep="\t", quote="\"", header=TRUE)
proteinAnnotData_validData <- proteinAnnotData_validData[!duplicated(proteinAnnotData_validData$PG.ProteinAccessions),c("PG.ProteinAccessions","PG.ProteinDescriptions","PG.ProteinNames")]
proteinAnnotData_validData$PG.ProteinNames <- gsub("_HUMAN","",proteinAnnotData_validData$PG.ProteinNames)

```

### Metadata

```{r}

# Read in pivot table
## Read in protein abundance table of actual samples
metaCrossSectional_validData <- read.table(here::here(metaDir, "preSchool_validationData_tallFormatClean_validationData.tsv"), sep="\t", header=TRUE, quote="") 
table(metaCrossSectional_validData$Condition)

# Keep just child cross-sectional
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",]
table(metaCrossSectional_validData$Condition)

# Exclude JIA
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Condition != "JIA",] #  & metaCrossSectional_validData$Condition != "Healthy"
table(metaCrossSectional_validData$Condition)

# Just conditions of interest
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Condition != "HPP",]

# Drop ethnicity outliers
# metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Ethnicity == "C" & !is.na(metaCrossSectional_validData$Ethnicity),]

```

### Quantification data

```{r}

# Read in pivot table
## Read in protein abundance table of actual samples
quantTable_validData <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed",
                                              "validationData_pivotTable_PGQuantity.sampleIDs_noSerumQCs.tsv"), sep="\t", quote="\"", header=TRUE)
rownames(quantTable_validData) <- quantTable_validData$PG.ProteinGroups

```

### Drop depleted proteins

```{r}

# Depleted prots
depletedProtData <- read.table(here::here(baseDataDir, "annotation", "depletedProteinSet.txt"), sep="\t", header=FALSE)
depletedProts <- depletedProtData$V2

# Drop proteins
nrow(quantTable_validData)
quantTable_validData <- quantTable_validData[!quantTable_validData$PG.ProteinGroups %in% depletedProts,]
nrow(quantTable_validData)

```

```{r}

# Read in QC
quantTable_validDataQCs <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed", "QCs",
                                                "validData_pivotTable_PGQuantity.sampleIDs_justSerumQCs.tsv"), sep="\t", quote="\"", header=TRUE)
rownames(quantTable_validDataQCs) <- quantTable_validDataQCs$PG.ProteinGroups
quantTable_validDataQCs <- quantTable_validDataQCs[rownames(quantTable_validData),]

# Visualise
dim(quantTable_validDataQCs)

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

### Get proteins detected in all Serum QC samples

```{r}

# Get
rownames(quantTable_validDataQCs) <- quantTable_validDataQCs$PG.ProteinGroups
serumProtsValid <- GetSerumQCProteins(serumQCData=quantTable_validDataQCs); length(serumProtsValid)

```

## Checking for drift

```{r}

# Tall format
quantTable_validData_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_validDataQCs), 
                                                                    quantTable_validDataQCs[,2:ncol(quantTable_validDataQCs)]), id.vars=c("ID"))

# Rename QCs
quantTable_validData_serumQC_tall["variable"] <- gsub("PooledPlasma|_S.*","",quantTable_validData_serumQC_tall$variable)

# Add metadata
quantTable_validData_serumQC_tall["PLATE"] <- "P1"
quantTable_validData_serumQC_tall[which(quantTable_validData_serumQC_tall$variable %in% c("QC3","QC4")),]["PLATE"] <- "P2"

# Order
quantTable_validData_serumQC_tall <- quantTable_validData_serumQC_tall[order(quantTable_validData_serumQC_tall$variable, decreasing=FALSE),]
quantTable_validData_serumQC_tall$variable <- factor(quantTable_validData_serumQC_tall$variable, levels=unique(quantTable_validData_serumQC_tall$variable))

# Plot
p1 <- ggplot(quantTable_validData_serumQC_tall, 
             aes(x=variable, 
                 y=log2(value), 
                 group=variable)) + 
  geom_violin(fill="#99E6FF",
              alpha=0.75) + 
  geom_boxplot(outlier.size=1.5,
               width=0.125) + 
  ggtitle("Serum QC (Unfiltered)") +
  xlab("Serum QC Sample") + ylab("Log2 Protein\nAbundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=56))

# Save
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "validationData_serumQCSamples_proteinAbundance_unFiltered.png"),
       p1,
       units="px", width=5000, height=3500)

# Add to list
newList <- list(p1)
names(newList) <- c("UnFiltProtAbundanceSerumQC_validData")
plotList <- append(plotList,
                   newList)

```

## Missing values

### Enumerate frequencies of proteins missing in samples

```{r, fig.height=5, fig.width=12}

# Enumerate raw NAs
## Protein
perProtNAValid <- EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idxVal=1)
## Samples
perSmplNAValid <- EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idxVal=2)

# Histogram
protPlt <- ggplot(perProtNAValid,
       aes(x=N_MISSING)) +
  geom_histogram(fill="mediumblue") + 
  theme_classic(base_size=16)
smplPlt <- ggplot(perProtNAValid,
       aes(x=N_MISSING)) +
  geom_histogram(fill="mediumblue") + 
  theme_classic(base_size=16)

protPlt + smplPlt

```

### Enumerate frequencies of samples missing proteins

```{r, fig.height=4, fig.width=12}
 
# Test dropping values on frequencies of missing proteins and samples
perProtNAValid_dropProts <- TestProteinDropThresholds(perProtNA=perProtNAValid, quantTbl=quantTable_validData[,metaCrossSectional_validData$Sample_ID], idxVal=1)
perSmplNAValid_dropProts <- TestProteinDropThresholds(perProtNA=perSmplNAValid, quantTbl=quantTable_validData[,metaCrossSectional_validData$Sample_ID], idxVal=2)

# Keep only serum QC proteins
perProtNAValid_dropProts_serumQCOnly <- TestProteinDropThresholds(perProtNA=perProtNAValid[serumProtsValid,], 
                                                                  quantTbl=quantTable_validData[serumProtsValid,metaCrossSectional_validData$Sample_ID], idxVal=1)
perSmplNAValid_dropProts_serumQCOnly <- TestProteinDropThresholds(perProtNA=EnumerateNAs(quantTable_validData[serumProtsValid,metaCrossSectional_validData$Sample_ID], idxVal=2),
                                                                  quantTbl=quantTable_validData[serumProtsValid,metaCrossSectional_validData$Sample_ID], idxVal=2)

# Histogram
protPlt <- ggplot(perProtNAValid_dropProts,
       aes(x=N_MISSING)) +
  geom_histogram(fill="mediumblue") + 
  facet_grid(cols=vars(PERCENTAGE)) +
  theme_classic(base_size=16)
smplPlt <- ggplot(perSmplNAValid_dropProts,
       aes(x=N_MISSING)) +
  geom_histogram(fill="mediumblue") + 
  facet_grid(cols=vars(PERCENTAGE)) +
  theme_classic(base_size=16)

protPlt
smplPlt

```

### Add to metadata

```{r}

# Combine
totalNProtsValid <- do.call("rbind",lapply(unique(perProtNAValid_dropProts$PERCENTAGE), function(perc) 
                                    data.frame(PERCENTAGE=perc, FREQ=nrow(perProtNAValid_dropProts[perProtNAValid_dropProts$PERCENTAGE == perc,]))))
totalNProtsValid_serumQC <- do.call("rbind",lapply(unique(perProtNAValid_dropProts_serumQCOnly$PERCENTAGE), function(perc) 
                                    data.frame(PERCENTAGE=perc, FREQ=nrow(perProtNAValid_dropProts_serumQCOnly[perProtNAValid_dropProts_serumQCOnly$PERCENTAGE == perc,]))))

```

### Correlate N missing with abundance

```{r}

# Correlate abundance and number of missing values
MeanFreqMissing_valid <- data.frame(MeanAbundance=rowMeans(quantTable_validData[,metaCrossSectional_validData$Sample_ID], na.rm=TRUE),
                                    FreqMissing=apply(quantTable_validData[,metaCrossSectional_validData$Sample_ID], 
                                                      1, function(row) length(row[is.na(row)])))

# Plot
## Correlate valid-sectional
stats::cor(MeanFreqMissing_valid[MeanFreqMissing_valid$FreqMissing > 0,]$MeanAbundance, 
    MeanFreqMissing_valid[MeanFreqMissing_valid$FreqMissing > 0,]$FreqMissing, method="spearman", use="complete.obs")

# Plot
ggplot(MeanFreqMissing_valid[MeanFreqMissing_valid$FreqMissing > 0,], aes(x=base::log2(MeanAbundance), y=FreqMissing)) + 
  geom_point() + 
  xlab("base::log2 Mean Protein Abundance") + ylab("Frequency of Missing") +
  theme_classic(base_size=16) + theme(axis.title=element_text(size=24))

```

### Keep only serum proteins

```{r}

# Keep serum proteins
quantTable_validData <- quantTable_validData[serumProtsValid,]
quantTable_validDataQCs <- quantTable_validDataQCs[serumProtsValid,]

```

## Checking for drift

```{r}

# Tall format
quantTable_validData_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_validDataQCs), 
                                                                    quantTable_validDataQCs[,2:ncol(quantTable_validDataQCs)]), id.vars=c("ID"))

# Rename QCs
quantTable_validData_serumQC_tall["variable"] <- gsub("PooledPlasma|_S.*","",quantTable_validData_serumQC_tall$variable)

# Add metadata
quantTable_validData_serumQC_tall["PLATE"] <- "P1"
quantTable_validData_serumQC_tall[which(quantTable_validData_serumQC_tall$variable %in% c("QC3","QC4")),]["PLATE"] <- "P2"

# Order
quantTable_validData_serumQC_tall <- quantTable_validData_serumQC_tall[order(quantTable_validData_serumQC_tall$variable, decreasing=FALSE),]
quantTable_validData_serumQC_tall$variable <- factor(quantTable_validData_serumQC_tall$variable, levels=unique(quantTable_validData_serumQC_tall$variable))

# Plot
p1 <- ggplot(quantTable_validData_serumQC_tall, 
             aes(x=variable, 
                 y=log2(value), 
                 group=variable)) + 
  geom_violin(fill="#99E6FF",
              alpha=0.75) + 
  geom_boxplot(outlier.size=1.5,
               width=0.125) + 
  ggtitle("Serum QC (Filtered for Missingness)") +
  xlab("Serum QC Sample") + ylab("Log2 Protein\nAbundance") +
  facet_grid(cols=vars(PLATE), scales="free", space="free") +
  theme_classic(base_size=48) +
  theme(axis.title=element_text(size=72),
        strip.text=element_text(size=72),
        plot.title=element_text(size=56))

# Save
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed",
                  "validationData_serumQCSamples_proteinAbundance_filteredForNoNaMissingInSerumQCProts.png"),
       p1,
       units="px", width=5000, height=3500)

# Add to list
newList <- list(p1)
names(newList) <- c("SerumQCFiltProtAbundanceSerumQC_validData")
plotList <- append(plotList,
                   newList)

```

### Add to metadata

```{r}

# Combine with metadata
perSmplNAValid_dropProts <- merge(perSmplNAValid_dropProts, metaCrossSectional_validData, by.x="ID", by.y="Sample_ID")
perSmplNAValid_dropProts_serumQCOnly <- merge(perSmplNAValid_dropProts_serumQCOnly, metaCrossSectional_validData, by.x="ID", by.y="Sample_ID")

```

### Plot

NOTE: Not filtering proteins separately by plate.

```{r, fig.width=3.5, fig.height=2}

# Cross-sectional
source(here::here("utilities", "R", "codeModules", "plotting", "plotMissingness_crossSectional_validation.R"))

# Save
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "propMissingValues_missingValueThresholds.png"),
       boxPlt,
       units="px", width=4000, height=4000)
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "propMissingValues_missingValueThresholds_byPlate.png"),
       boxPltByPlate,
       units="px", width=10000, height=3000)

ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "propMissingValues_missingValueThresholds_justSerumQC.png"),
       boxPlt_serumQCOnly,
       units="px", width=4000, height=4000)
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "propMissingValues_missingValueThresholds_justSerumQC_byPlate.png"),
       boxPltByPlate_serumQCOnly,
       units="px", width=10000, height=3000)
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "nProteins_missingValueThresholds.png"),
       barPlt + barPlt_serumQCOnly +
         plot_layout(nrow=2,
                     axis_titles = "collect") &
         theme_classic(base_size=48) &
         theme(axis.title=element_text(size=56),
               axis.text.x=element_text(size=32)),
       units="px", width=4000, height=4000)

# Make big
bigPlt <- barPlt + barPlt_serumQCOnly +
          plot_layout(nrow=2,
                      guides = "collect",
                      axis_titles = "collect")
ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                  "nProteins_missingValueThresholds.png"),
       bigPlt,
       units="px", width=4000, height=4000)

# Add to list
newList <- list(boxPlt,
                boxPlt_serumQCOnly,
                barPlt + theme_classic(base_size=48) + theme(axis.title=element_text(size=56),
                                                        axis.text.x=element_text(size=32)),
                barPlt_serumQCOnly + theme_classic(base_size=48) + theme(axis.title=element_text(size=56),
                                                                    axis.text.x=element_text(size=32)))
names(newList) <- c("PropMissingValsUnFilt_validData",
                    "PropMissingValsSerumQCProts_validData",
                    "NProtsAllProts_validData",
                    "NProtsSerumQCProts_validData")
plotList <- append(plotList,
                   newList)

```

### Investigate samples with high numbers of missing proteins

Define this as \> 50%

```{r}

# Check proportion missing
perSmplNAValid_dropProts[perSmplNAValid_dropProts$PROP_MISSING > 0.5,]

```

### Filter quantification sample

Dont filter samples except for those with high missingness

```{r}

# Ensure consistent
serumProtsValid <- intersect(rownames(quantTable_validData),
                             serumProtsValid)

# Retain only serum QC proteins
quantTable_validData <- quantTable_validData[serumProtsValid,]
dim(quantTable_validData)

# Drop proteins with high frequencies of missing values
validPropProts <- EnumerateNAs(quantTable_validData[serumProtsValid,], idx=1)
quantTable_validData <- quantTable_validData[validPropProts[validPropProts$PROP_MISSING < maxPropMissing,]$ID,]

# Define samples with high frequencies of missing values
validPropSmpls <- EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idx=2)
validPropSmpls <- merge(validPropSmpls, metaCrossSectional_validData, by.x="ID", by.y="Sample_ID")
table(validPropSmpls[validPropSmpls$PROP_MISSING >= maxPropMissing_samples,]$Condition)

# Order by missingness
validPropSmpls[order(validPropSmpls$PROP_MISSING, decreasing=TRUE),]

# Subset
quantTable_validData <- data.frame(PG.ProteinAccessions=rownames(quantTable_validData), 
                                   quantTable_validData[,c(validPropSmpls[validPropSmpls$PROP_MISSING < maxPropMissing_samples,]$ID)])

# Ensure consistent
serumProtsValid <- intersect(rownames(quantTable_validData),
                             serumProtsValid)

rownames(metaCrossSectional_validData) <- metaCrossSectional_validData$Sample_ID
table(metaCrossSectional_validData[colnames(quantTable_validData)[-1],]$Condition)
sum(table(metaCrossSectional_validData[colnames(quantTable_validData)[-1],]$Condition))

```

### Enumerate missing values by sample again

Check proper filtration

```{r}

# Enumerate more missing values per sample
validPropSmpls <- EnumerateNAs(quantTable_validData, idx=2)

# Check proportion of missing
validPropSmpls[validPropSmpls$PROP_MISSING >= maxPropMissing_samples,]

```

### Subset metadata to match filtered columns

```{r}

# Subset metadata
metaCrossSectional_validData <- metaCrossSectional_validData[which(metaCrossSectional_validData$Sample_ID %in% colnames(quantTable_validData)),]

```

### Check protein and sample missing rates

```{r}

# Check max missingness
if (base::max(EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idx=2)$PROP_MISSING) >= maxPropMissing_samples) { 
  
  warning("Sample Missingness for Validation data not controlled at desired level") 
  
}

# Check max missingness
if (base::max(EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idx=1)$PROP_MISSING) >= maxPropMissing) { 
  
  warning("Protein Missingness for Validation data not controlled at desired level") 
  
}

```

### Report missingness now

```{r}

# SUbset metadta
metaCrossSectional_validData <- metaCrossSectional_validData[colnames(quantTable_validData)[-1],]

# Check mean missingess by protein and samples
max(EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idx=1)$PROP_MISSING)
max(EnumerateNAs(quantTable_validData[,metaCrossSectional_validData$Sample_ID], idx=2)$PROP_MISSING)

```

### Save

```{r}

# Create output
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered"), showWarnings=FALSE, recursive=TRUE)

# Save cross-Sectional
write.table(quantTable_validData, here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered", 
                                             paste0("validationData_quantTable.filtBy_serumQCProts_maxMissing",maxPropMissing,".tsv")), sep="\t", quote=FALSE, row.names=FALSE)

```

### Checking for drift

```{r, fig.width=10}

# Subset
quantTable_validDataQCs <- quantTable_validDataQCs[rownames(quantTable_validData),]

# Tall format
quantTable_validData_serumQC_tall <- reshape2::melt(data.frame(ID=rownames(quantTable_validDataQCs), 
                                                               quantTable_validDataQCs[,2:ncol(quantTable_validDataQCs)]), id.vars=c("ID"))

# Order
quantTable_validData_serumQC_tall <- quantTable_validData_serumQC_tall[order(quantTable_validData_serumQC_tall$variable, decreasing=FALSE),]
quantTable_validData_serumQC_tall$variable <- factor(quantTable_validData_serumQC_tall$variable, levels=unique(quantTable_validData_serumQC_tall$variable))

# Plot
ggplot(quantTable_validData_serumQC_tall, aes(x=variable, y=base::log2(value))) + 
  geom_boxplot() + ggtitle("Serum QC (Cross-Sectional)") +
  xlab("Serum QC Sample") + ylab("base::log2 Protein Abundance") +
  theme_classic()

```

### Cluster missing values

#### Preprocess and define colours

```{r}

# Prepare matrices
quantTableValid_naZeros <- quantTable_validData[,metaCrossSectional_validData$Sample_ID]

# Set non-zero to 1
quantTableValid_naZeros[!is.na(quantTableValid_naZeros)] <- 1

# Set NA to 0
quantTableValid_naZeros[is.na(quantTableValid_naZeros)] <- 0

# Tidy names
colnames(quantTableValid_naZeros) <- gsub(" Osteo","\nOsteo",colnames(quantTableValid_naZeros))

# Harmonize data with metadata
quantTableValid_naZeros <- quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID]

# Define shared color
metaCrossSectional_validData$Age[is.na(metaCrossSectional_validData$Age)] <- 0
mainColFun = colorRamp2(c(0, 1), c("white", "mediumblue"))
ageColFun_cross  = colorRamp2(c(0,min(metaCrossSectional_validData$Age, na.rm=T), max(metaCrossSectional_validData$Age, na.rm=T)), c("darkgrey","white","darkred"))

```

# Empty chart

png(file=here::here(diaEdaDir, "conditionLegenedPlot.png"), units="px", height=45, width=500) \# No margin around chart par(mar=c(0,0,0,0)) plot(0, 0, type = "n", xlim = c(0, 1), ylim = c(0, 1), axes = FALSE, xlab = "", ylab = "")

# Settings

line \<- 1 col \<- length(condCols)

# Add color background

rect(\
rep((0:(col - 1)/col),line) ,\
sort(rep((0:(line - 1)/line),col),decreasing=T),\
rep((1:col/col),line) , sort(rep((1:line/line),col),decreasing=T),\
border = "white" , col=condCols\[seq(1,line\*col)\])

# Color names

text(\
rep((0:(col - 1)/col),line)+0.1 ,\
sort(rep((0:(line - 1)/line),col),decreasing=T)+0.5 , names(condCols)\[seq(1,line\*col)\] , cex=1)

dev.off()

#### Plot as heatmaps

```{r}

# Generate output
dir.create(here::here(resDir, "validation", "eda"))

# Heatmap
crossPlt <- Heatmap(quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID],
                    
                    # Names
                    # name = "Child",
                    
                    # Colours
                    col = mainColFun,
                    
                    # Legend direction
                    heatmap_legend_param = list(title = "base::log2 Abundance",
                                                direction = "horizontal",
                                                legend_width = unit(10, "cm"),
                                                grid_height = unit(1.25, "cm"),
                                                labels_gp = gpar(fontsize=38),
                                                legend_gp = gpar(fontsize=72)),
                    show_heatmap_legend =FALSE,
                    
                    # Clustering
                    clustering_distance_rows = "manhattan",
                    clustering_distance_columns = "manhattan",
                    
                    # Columns
                    column_title_rot = 65,
                    column_split = metaCrossSectional_validData[colnames(quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID]),]$PLATE,
                    top_annotation = HeatmapAnnotation(Condition = anno_simple(metaCrossSectional_validData$Condition,
                                                                               simple_anno_size = unit(2, "cm"), col=condCols),
                                                       Age = anno_simple(metaCrossSectional_validData$Age, col = ageColFun_cross,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Sex = anno_simple(metaCrossSectional_validData$Sex, col=sexCols,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Plate = anno_simple(as.character(metaCrossSectional_validData$PLATE), col=plateCols,
                                                                           simple_anno_size = unit(2, "cm")),
                                                       Site = anno_simple(as.character(metaCrossSectional_validData$Site), col=siteCols,
                                                                          simple_anno_size = unit(2, "cm")),
                                                       annotation_name_gp= gpar(fontsize = 48)),
                    column_title_gp = gpar(fontsize = 42, fontface = "bold"),
                    
                    # Rows
                    show_row_names = FALSE,
                    border = TRUE
)

png(file=here::here(resDir, "crossSectional", "validation", "eda", "imputationCheck_naZeroHeatmap_splitByPlate_validationData.png"), units="px", width=2500, height=2500)
draw(crossPlt, ht_gap = unit(5, "cm"), heatmap_legend_side = "bottom", legend_title_gp = gpar(fontsize=72),
     column_title="Validation Data (Split by Plate)")
dev.off()

# Heatmap
crossPlt <- Heatmap(quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID],
                    
                    # Names
                    # name = "Child",
                    
                    # Colours
                    col = mainColFun,
                    
                    # Legend direction
                    heatmap_legend_param = list(title = "base::log2 Abundance",
                                                direction = "horizontal",
                                                legend_width = unit(10, "cm"),
                                                grid_height = unit(1.25, "cm"),
                                                labels_gp = gpar(fontsize=38),
                                                legend_gp = gpar(fontsize=72)),
                    show_heatmap_legend =FALSE,
                    
                    # Clustering
                    clustering_distance_rows = "manhattan",
                    clustering_distance_columns = "manhattan",
                    
                    # Columns
                    column_title_rot = 65,
                    column_split = metaCrossSectional_validData[colnames(quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID]),]$Condition,
                    top_annotation = HeatmapAnnotation(Condition = anno_simple(metaCrossSectional_validData$Condition,
                                                                               simple_anno_size = unit(2, "cm"), col=condCols),
                                                       Age = anno_simple(metaCrossSectional_validData$Age, col = ageColFun_cross,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Sex = anno_simple(metaCrossSectional_validData$Sex, col=sexCols,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Plate = anno_simple(as.character(metaCrossSectional_validData$PLATE), col=plateCols,
                                                                           simple_anno_size = unit(2, "cm")),
                                                       Site = anno_simple(as.character(metaCrossSectional_validData$Site), col=siteCols,
                                                                          simple_anno_size = unit(2, "cm")),
                                                       annotation_name_gp= gpar(fontsize = 48)),
                    column_title_gp = gpar(fontsize = 42, fontface = "bold"),
                    
                    # Rows
                    show_row_names = FALSE,
                    border = TRUE
)

png(file=here::here(resDir, "crossSectional", "validation", "eda", "imputationCheck_naZeroHeatmap_splitByCondition_validationData.png"), units="px", width=2500, height=2500)
draw(crossPlt, ht_gap = unit(5, "cm"), heatmap_legend_side = "bottom", legend_title_gp = gpar(fontsize=72),
     column_title="Validation Data (Split by Condition)")
dev.off()

# Heatmap
crossPlt <- Heatmap(quantTableValid_naZeros[,metaCrossSectional_validData$Sample_ID],
                    
                    # Names
                    # name = "Child",
                    
                    # Colours
                    col = mainColFun,
                    
                    # Legend direction
                    heatmap_legend_param = list(title = "base::log2 Abundance",
                                                direction = "horizontal",
                                                legend_width = unit(10, "cm"),
                                                grid_height = unit(1.25, "cm"),
                                                labels_gp = gpar(fontsize=38),
                                                legend_gp = gpar(fontsize=72)),
                    show_heatmap_legend =FALSE,
                    
                    # Clustering
                    clustering_distance_rows = "manhattan",
                    clustering_distance_columns = "manhattan",
                    
                    # Columns
                    column_title_rot = 65,
                    top_annotation = HeatmapAnnotation(Condition = anno_simple(metaCrossSectional_validData$Condition,
                                                                               simple_anno_size = unit(2, "cm"), col=condCols),
                                                       Age = anno_simple(metaCrossSectional_validData$Age, col = ageColFun_cross,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Sex = anno_simple(metaCrossSectional_validData$Sex, col=sexCols,
                                                                         simple_anno_size = unit(2, "cm")),
                                                       Plate = anno_simple(as.character(metaCrossSectional_validData$PLATE), col=plateCols,
                                                                           simple_anno_size = unit(2, "cm")),
                                                       Site = anno_simple(as.character(metaCrossSectional_validData$Site), col=siteCols,
                                                                          simple_anno_size = unit(2, "cm")),
                                                       annotation_name_gp= gpar(fontsize = 48)),
                    column_title_gp = gpar(fontsize = 42, fontface = "bold"),
                    
                    # Rows
                    show_row_names = FALSE,
                    border = TRUE
)

png(file=here::here(resDir, "crossSectional", "validation", "eda", "imputationCheck_naZeroHeatmap_validationData.png"), units="px", width=2500, height=2500)
draw(crossPlt, ht_gap = unit(5, "cm"), heatmap_legend_side = "bottom", legend_title_gp = gpar(fontsize=72),
     column_title="Validation Data")
dev.off()


```

### View distributions as boxplots

```{r}

# Get as tall
rawValidTall <- melt(data.frame(ID=rownames(quantTable_validData), 
                                base::log2(quantTable_validData[,metaCrossSectional_validData$Sample_ID])), id.vars="ID")

# Merge with
rawValidTall <- merge(rawValidTall, metaCrossSectional_validData, by.x="variable", by.y="Sample_ID")

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","Age_Group","PLATE","Site")) {

  # Plot
  plt <- ggplot(rawValidTall, aes(x=variable, y=value, fill=.data[[fctr]])) + 
    geom_boxplot() + 
    xlab("Samples") + ylab("Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic() + theme(axis.text.x=element_blank(), 
                       legend.position="top") +
    guides(fill=guide_legend(nrow=1, byRow=TRUE))
  print(plt)
  
}  

```

### Generate Summarize Experiment for DEP

```{r}

# Generate design matrix
designValid_m <- data.frame(label=rownames(metaCrossSectional_validData), 
                            Sex=metaCrossSectional_validData$Sex, 
                            Age=metaCrossSectional_validData$Age, 
                            condition=metaCrossSectional_validData$Condition)
rownames(designValid_m) <- metaCrossSectional_validData$Sample_ID

# Add replicate column
designValid_m["replicate"] <- NA
for (cond in unique(designValid_m$condition)) { designValid_m[designValid_m$condition == cond,]["replicate"] <- 1:nrow(designValid_m[designValid_m$condition == cond,]) }

# Make summarized experiment
uniqNames_quantTableValid  <- make_unique(quantTable_validData[,c("PG.ProteinAccessions","PG.ProteinAccessions",metaCrossSectional_validData$Sample_ID)], 
                                          names="PG.ProteinAccessions", ids="PG.ProteinAccessions", delim=";")
seValid <- make_se(uniqNames_quantTableValid, which(colnames(uniqNames_quantTableValid) %in% metaCrossSectional_validData$Sample_ID), designValid_m)

# Rough fix I think
assay(seValid)[is.nan(assay(seValid))] <- NA

```

### Plot frequencies for missing values

```{r}

# Plot protein identification overlaps
plot_frequency(seValid) + theme(axis.text=element_blank())

# Plot protein identification overlaps (stacked bars)
plot_coverage(seValid) + theme(legend.position="none")

# Plot proteins per samples
plot_numbers(seValid) + scale_fill_manual(values=condCols) + 
  theme(axis.text.x=element_blank(), legend.position="none") 

```

### Scale and normalise with VSN

```{r}

# Normalise
## VSN
normValid <- normalize_vsn(seValid)

# Plot SD
meanSdPlot(normValid)

# Check mix/min for VSN - some strange things here with some parameterisations
min(assay(normValid), na.rm=T)

```

### View distributions as boxplots

```{r}

# Get as tall
normValidTall <- melt(data.frame(ID=rownames(assay(normValid)), assay(normValid)), id.vars="ID")
seValidTall <- melt(data.frame(ID=rownames(assay(seValid)), assay(seValid)), id.vars="ID")

# Merge with column data
normValidTall <- merge(normValidTall, data.frame(colData(normValid)), by.x="variable", by.y="ID")
seValidTall <- merge(seValidTall, data.frame(colData(seValid)), by.x="variable", by.y="ID")

# Merge with
normValidTall <- merge(normValidTall[,c("variable","ID","label","value")], metaCrossSectional_validData, by.x="label", by.y="Sample_ID")
seValidTall <- merge(seValidTall[,c("variable","ID","label","value")], metaCrossSectional_validData, by.x="label", by.y="Sample_ID")

```

### View distributions as boxplots

```{r}

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","Age_Group","PLATE","Site")) { # ,"Ethnicity"

  # Plot
  plt <- ggplot(normValidTall, aes(x=variable, y=value, fill=.data[[fctr]])) + 
    geom_boxplot() + 
    xlab("Samples") + ylab("Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic() + theme(axis.text.x=element_blank(), 
                       legend.position="top") +
    guides(fill=guide_legend(nrow=1, byRow=TRUE))
  print(plt)
  
}

```

### Plot sample distributions

```{r}

# Get number of conditions per site
freqCondPerSite <- melt(data.frame(table(metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",][,c("Condition","Site")])))

# Get proportion of conditions per site
freqCondPerSite["Prop"] <- NA
sites <- unique(freqCondPerSite$Site)
for (i in seq_len(length(sites))) { freqCondPerSite[freqCondPerSite$Site == sites[[i]],]["Prop"] <- freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value / sum(freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value) }

# Plot freq
plt1 <- ggplot(freqCondPerSite,
       aes(x=Site, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of\nSamples") +
  theme_classic()

# Plot freq
plt2 <- ggplot(freqCondPerSite,
       aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportio of Samples (in Site)") +
  theme_classic()

plt1
plt2

# save
ggsave(here::here(resDir, "crossSectional", "validation", "eda",
                  paste0("proteinAbundanceDistribution_validationData_missingnormonMissing_empiricalDistributions.png")),
       plt1, width=5000, height=2750, units="px")

```


```{r}

# Get number of conditions per Plate
freqCondPerPlate <- melt(data.frame(table(metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child" & metaCrossSectional_validData$Condition != "JIA",][,c("Condition","PLATE")])))

# Get proportion of conditions per Plate
freqCondPerPlate["Prop"] <- NA
Plates <- unique(freqCondPerPlate$PLATE)
for (i in seq_len(length(Plates))) { freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]["Prop"] <- freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]$value / sum(freqCondPerPlate[freqCondPerPlate$PLATE == Plates[[i]],]$value) }

# Plot freq
ggplot(freqCondPerPlate,
       aes(x=PLATE, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of Samples") +
  theme_classic()

# Plot prob
ggplot(freqCondPerPlate,
       aes(x=PLATE, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportion of Samples (in Plate)") +
  theme_classic()

# Plot freq 
ggplot(aggregate(value ~ Condition + variable, data=freqCondPerPlate[,c("Condition","value","variable")], FUN=sum),
       aes(x=variable, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of Samples") +
  xlab("") + 
  theme_classic() + theme(axis.text.x=element_blank())


```

### Plot sample distributions

```{r}

# Get number of conditions per site
freqCondPerSite <- melt(data.frame(table(metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",][,c("Condition","Site")])))

# Get proportion of conditions per site
freqCondPerSite["Prop"] <- NA
sites <- unique(freqCondPerSite$Site)
for (i in seq_len(length(sites))) { freqCondPerSite[freqCondPerSite$Site == sites[[i]],]["Prop"] <- freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value / sum(freqCondPerSite[freqCondPerSite$Site == sites[[i]],]$value) }

# Plot freq
plt1 <- ggplot(freqCondPerSite,
       aes(x=Site, y=value, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Number of\nSamples") +
  theme_classic(base_size=36) + theme(axis.title=element_text(size=88),
                                 axis.text=element_text(size=42),
                                 legend.text=element_text(size=42)) +
  guides(fill=guide_legend(override.aes=list(size=12)))

# Plot freq
plt2 <- ggplot(freqCondPerSite,
       aes(x=Site, y=Prop, fill=Condition)) + 
  geom_bar(stat="identity", color="black", position="stack") +
  scale_fill_manual(values=condCols) +
  ylab("Proportio of\nSamples") +
  theme_classic(base_size=36) + theme(axis.title=element_text(size=88),
                                 axis.text=element_text(size=42),
                                 legend.text=element_text(size=42)) +
  guides(fill=guide_legend(override.aes=list(size=12)))

# Plot
plt <- plt1 + plt2 +
         plot_layout(nrow=1,
                     guides = "collect",
                     axis_titles = "collect")

# save
ggsave(here::here(resValidDir, "eda", 
                  paste0("sampleGroupNumbers_bySite_afterFiltration.png")),
       plt, width=10000, height=4000, units="px")

# Add to list
newList <- list(plt1)
names(newList) <- c("sampleConditionGroupNumbers_validData")
plotList <- append(plotList,
                   newList)

```

### Save filtered metadata

```{r}

# Save metadata
write.table(metaCrossSectional_validData, here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                                                     "preSchool_validationData_tallFormatClean_samplesFiltered_byMissingness.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Plot abundances of depleted proteins

Test to see group-structure of depleted proteins is present prior to any real normalisation or processing.

```{r}

# Add if depleted protein
qntTable_depletedProt <- reshape2::melt(quantTable_validData[c("P02768","P02763","P01023","P01011"),], id.vars="PG.ProteinAccessions")
qntTable_depletedProt <- merge(metaCrossSectional_validData[metaCrossSectional_validData$Site != "Leiden" & metaCrossSectional_validData$Condition != "JIA",], qntTable_depletedProt, by.x="Sample_ID", by.y="variable")

# Plot
ggplot(qntTable_depletedProt, 
       aes(y=base::log2(value), x=Condition, fill=Condition)) +
  geom_boxplot() + 
  geom_jitter(size=0.75) +
  facet_grid(cols=vars(PG.ProteinAccessions), scale="free_y", space="free_y") +
  scale_fill_manual(values=condCols) +
  theme_classic() + theme(axis.text=element_blank())

```

### Save filtered proteins

```{r}

# Save proteisn
saveRDS(rownames(quantTable_validData[,metaCrossSectional_validData$Sample_ID]), 
        here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                   paste0("filteredProteinSet_validationData_maxMissing",maxPropMissing,".rds")))

# Save quant table
saveRDS(quantTable_validData[,metaCrossSectional_validData$Sample_ID], 
        here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                   paste0("filtered_unNormalised_validationData_maxMissing",maxPropMissing,".rds")))
write.table(quantTable_validData[,metaCrossSectional_validData$Sample_ID], 
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered", 
                       paste0("filtered_unNormalised_validationData_maxMissing",maxPropMissing,".tsv")), sep="\t", quote=FALSE)

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "11_validationDataProcessQC_figs.rds"))

```






