---
title: "Validation Data Normalisation & EDA"
author: "Euan McDonnell"
format: html
editor: source
---

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
here::i_am("12_validationDataNormEDA.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "11_validationDataProcessQC_figs.rds"))
 
```

## Load data

### Metadata

```{r}

# Source BModal CNO Discovery study metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata.R"))

# Read in pivot table
## Read in protein abundance table of actual samples
metaCrossSectional_validData <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                                                     "preSchool_validationData_tallFormatClean_samplesFiltered_byMissingness.tsv"), sep="\t", header=TRUE) 
rownames(metaCrossSectional_validData) <- metaCrossSectional_validData$Sample_ID

# Keep just child cross-sectional
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",]
table(metaCrossSectional_validData$Condition)

# Exclude JIA
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Condition != "JIA",]
table(metaCrossSectional_validData$Condition)

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

### Quantification data

```{r}

# Read in pivot table
## Read in protein abundance table 
quantTable_validData <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered", 
                                              paste0("filtered_unNormalised_validationData_maxMissing",maxPropMissing,".tsv")), 
                                   sep="\t", quote="\"", header=TRUE)
quantTable_discoData <- read.table(here::here(diaDir, "normalised", "crossSectional_quantTable_normalisedCyclicLoess.tsv"), 
                                   sep="\t", quote="\"", header=TRUE, row.names=1)

# Read in QCs
quantTable_validData_serumQC <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "processed", "QCs",
                                                "validData_pivotTable_PGQuantity.sampleIDs_justSerumQCs.tsv"),
                                           sep="\t", quote="\"", header=TRUE)
rownames(quantTable_validData_serumQC) <- quantTable_validData_serumQC$PG.ProteinGroups
quantTable_validData_serumQC <- quantTable_validData_serumQC[,-1]

# Combine data
quantTable_validWithQCs <- cbind(quantTable_validData_serumQC[rownames(quantTable_validData),], 
                                 quantTable_validData)
                                 

# Read BModal CNO Discovery imputed data
discoveryProts_qntTable <- read.table(here::here(diaDir, "normalised", paste0("crossSectional_quantTable_normalisedCyclicLoess.tsv")), sep="\t", header=TRUE, row.names=1)

```

#### Format metadata for NormalyzerDE

```{r}

# Sort metadata
forNormDEMeta <- plyr::rename(metaCrossSectional_validData[colnames(quantTable_validData),c("Sample_ID","Condition","Sex","Age","PLATE")], 
                              c("Sample_ID"="sample", "Condition"="group", "PLATE"="Plate"))

## Subset metadata
forNormDEMeta <- plyr::rename(metaCrossSectional_validData[colnames(quantTable_validData),c("Sample_ID","Condition")], 
                              c("Sample_ID"="sample", "Condition"="group"))
forNormDEMeta <- rbind(data.frame(sample=colnames(quantTable_validData_serumQC)[-1],
                                  group=rep("Serum_QC", ncol(quantTable_validData_serumQC)-1)),
                                  forNormDEMeta)

## Save metadata
write.table(forNormDEMeta, 
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                       "preSchool_validationData_groupIsCondition_forNormalyzerDE.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

## Save metadata
write.table(forNormDEMeta, 
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                       "preSchool_validationData_groupIsPlate_forNormalyzerDE.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

# Save abundance data
write.table(cbind(PG.ProteinAccessions=rownames(quantTable_validWithQCs), quantTable_validWithQCs[,forNormDEMeta$sample]),
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                       paste0("validationData_quantTable.filtBy_serumQCProts_maxMissing",0.3,".withSerumQCSamples.forNormalyzerDE.tsv")),
            sep="\t", quote=FALSE, row.names=FALSE)

```

#### Run Normalyzer

Based on: https://www.bioconductor.org/packages/release/bioc/vignettes/NormalyzerDE/inst/doc/vignette.html

```{r}

# Create object for NormalyzerDE
experimentObj_byCond  <- setupRawDataObject(designPath=here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                                                                  "preSchool_validationData_groupIsCondition_forNormalyzerDE.tsv"), 
                                            dataPath=here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered", 
                                                                paste0("validationData_quantTable.filtBy_serumQCProts_maxMissing",0.3,".withSerumQCSamples.forNormalyzerDE.tsv")))
experimentObj_byPlate <- setupRawDataObject(designPath=here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                                                                  "preSchool_validationData_groupIsPlate_forNormalyzerDE.tsv"), 
                                            dataPath=here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered", 
                                                                paste0("validationData_quantTable.filtBy_serumQCProts_maxMissing",0.3,".withSerumQCSamples.forNormalyzerDE.tsv")))
normObj_byCond <- getVerifiedNormalyzerObject("metaValidData_NormalyzerDE_byCondition", 
                                              experimentObj_byCond,
                                              noLogTransform=TRUE)
normObj_byPlate <- getVerifiedNormalyzerObject("metaValidData_NormalyzerDE_byPlate", 
                                               experimentObj_byPlate,
                                               noLogTransform=TRUE)

# Normalise
normResults_byCond <- normMethods(normObj_byCond)
normResults_byPlate <- normMethods(normObj_byPlate)

# Generate performance measures
normResultsWithEval_byCond <- analyzeNormalizations(normResults_byCond)
normResultsWithEval_byPlate <- analyzeNormalizations(normResults_byPlate)

# Save results
jobDir_byCond <-  setupJobDir("metaValidData_NormalyzerDE_byCondition", 
                              here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised"))
jobDir_byPlate <-  setupJobDir("metaValidData_NormalyzerDE_byPlate", 
                              here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised"))
writeNormalizedDatasets(normResultsWithEval_byCond, jobDir_byCond)
writeNormalizedDatasets(normResultsWithEval_byPlate, jobDir_byPlate)

# Plots
generatePlots(normResultsWithEval_byCond, jobDir_byCond)
generatePlots(normResultsWithEval_byPlate, jobDir_byPlate)

```

### Read in normalised data

```{r}

# List files
normFiles <- list.files(path=jobDir_byCond, pattern="*.txt")
normMs <- lapply(normFiles, function(file) read.table(here::here(jobDir_byCond, file), sep="\t", header=TRUE, row.names=1))
names(normMs) <- gsub("-.*|.*_|[.]txt","",normFiles)

```

### Check if normalised data with LOESS is unsupervised

Cyclic LOESS may be spliting by group-data to normalise which may be problematic for downstream modelling. Check this by manually normalising nad commparing normalised matrices. Have to round for slight variations in precision of normaliastion.

```{r}

# Check
cycLoessNorm <- limma::normalizeCyclicLoess(base::log2(as.matrix(quantTable_validWithQCs[,forNormDEMeta$sample])), method="fast")

# Get whether LOESS is unsupervisde
checkLoessUnSupervised <- all(round(as.matrix(cycLoessNorm[,colnames(normMs$CycLoess)]), 8) == round(as.matrix(as.matrix(normMs$CycLoess)), 8), na.rm=TRUE)
if (!checkLoessUnSupervised) { warning("LOESS APPEARS TO NOT BE UNSUPERVISED, IT MIGHT BE CONSIDERING GROUP DATA")}

```

### Save

```{r}

# Save
write.table(data.frame(ProteinGroup=rownames(normMs$CycLoess), normMs$CycLoess), 
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised",
                       "validationData_quantTable_normalisedCyclicLoess.tsv"), sep="\t", quote=FALSE, row.names=FALSE)
write.table(data.frame(ProteinGroup=rownames(normMs$CycLoess), normMs$CycLoess[,!grepl("QC",colnames(normMs$CycLoess))]), 
            here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised",
                       "validationData_quantTable_normalisedCyclicLoess_noQCSamples.tsv"), sep="\t", quote=FALSE, row.names=FALSE)

```

### Process normalised data

```{r}

# Make tall
normTall <- lapply(names(normMs), function(method) data.frame(reshape2::melt(data.frame(ProtID=rownames(normMs[[method]]), normMs[[method]]), id.vars="ProtID"), Method=method))
names(normTall) <- names(normMs)
  
# Combine
normTall <- do.call("rbind",normTall)

# Retain only QC samples
normTall_qcOnly <- normTall[grepl("QC",normTall$variable),]

```

### Plot

```{r}

# Split by Serum/instrument QC
normTall_qcOnly["QCType"] <- "Serum QC"

# Rename instrument QC
# normTall_qcOnly$variable <- gsub("Serum.*Instrument|late","",normTall_qcOnly$variable)

# Split into serum/instrument QC
serumQC_norm <- normTall_qcOnly[normTall_qcOnly$QCType == "Serum QC",]

# Retain only proteins that had missing values to impute missing values
serumQC_norm_qcOnly <- serumQC_norm[which(serumQC_norm$ProtID %in% 
                                                    rownames(quantTable_validData_serumQC[,!grepl("Instrument",colnames(quantTable_validData_serumQC))])),]
  
# Plot
serumQC_plt <- ggplot(serumQC_norm_qcOnly, aes(x=variable, y=log2(value))) + 
         geom_boxplot() + ggtitle("Serum QC Samples Only") +
         geom_jitter() +
         xlab("QC Sample") + ylab("Log2 Protein Abundance") +
         facet_wrap(~ Method, ncol=4,
                    scale="free_y") +
         theme_classic(base_size=16) + 
         theme(axis.text=element_blank(),
               axis.title=element_text(size=48),
               plot.title=element_text(size=64),
               strip.text=element_text(size=32))

```

### Get data without normalisation

```{r}

# Extract normalised data
normData <- normMs$CycLoess

# Drop QCs
normData <- normData[,!grepl("QC",colnames(normData))]

# Drop NAs
normData_noNA <- normData[!is.na(rowSums(normData)),]
dim(normData_noNA)

```

### View raw distributions as boxplots

```{r, fig.height=3, fig.width=6}

# Get as tall
rawCrossTall <- melt(data.frame(ID=rownames(quantTable_validData), 
                                log2(quantTable_validData[,metaCrossSectional_validData$Sample_ID])), id.vars="ID")

# Merge with
rawCrossTall <- merge(rawCrossTall, metaCrossSectional_validData, by.x="variable", by.y="Sample_ID")

# Get list of colours
allColors <- list(Sex=sexCols,
                  Condition=condCols,
                  Site=siteCols,
                  PLATE=plateCols)

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","PLATE","Site")) {

  # Plot
  p1 <- ggplot(rawCrossTall[rawCrossTall$Age_Group == "Child",], 
               aes(x=variable, 
                   y=value, 
                   fill=.data[[fctr]])) + 
    geom_boxplot() + 
    scale_fill_manual(values=allColors[[fctr]]) +
    xlab("Samples") + ylab("Log2 Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic(base_size=48) + 
    theme(axis.text.x=element_blank(), 
          axis.ticks.x=element_blank(),
          axis.title.x=element_text(size=72),
          axis.title.y=element_text(size=60),
          strip.text=element_blank(),
          legend.text=element_text(size=28),
          legend.title=element_blank(),
          legend.position="top") +
    guides(fill=guide_legend(nrow=1, 
                             byRow=TRUE, 
                             override.aes=list(size=16)))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised",
                    paste0("sampleAbundanceDistribution_splitBy",fctr,"_rawAbundances_childSamples.png")),
         p1,
         units="px", width=5000, height=2750)

  if (fctr == "Condition") {
    
    # Add to list
    newList <- list(p1)
    names(newList) <- c("sampleAbundDistribution_colByCondition_raw_validData")
    plotList <- append(plotList,
                       newList)
    
  }
  
}

```

### View normalised distributions as boxplots

```{r, fig.height=3, fig.width=6}

# Get as tall
normCrossTall <- melt(data.frame(ID=rownames(normMs$CycLoess), 
                                normMs$CycLoess[,metaCrossSectional_validData$Sample_ID]), id.vars="ID")
# Merge with
normCrossTall <- merge(normCrossTall, metaCrossSectional_validData, by.x="variable", by.y="Sample_ID")

# Loop over factors and color/facet
for (fctr in c("Sex","Condition","PLATE","Site")) {

  # Plot
  p1 <- ggplot(normCrossTall[normCrossTall$Age_Group == "Child",], 
               aes(x=variable, 
                   y=value, 
                   fill=.data[[fctr]])) + 
    geom_boxplot() + 
    scale_fill_manual(values=allColors[[fctr]]) +
    xlab("Samples") + ylab("Log2 Abundance") +
    facet_grid(cols=vars(.data[[fctr]]), scale="free", space="free") +
    theme_classic(base_size=48) + 
    theme(axis.text.x=element_blank(), 
          axis.ticks.x=element_blank(),
          axis.title.x=element_text(size=72),
          axis.title.y=element_text(size=60),
          strip.text=element_blank(),
          legend.text=element_text(size=28),
          legend.title=element_blank(),
          legend.position="top") +
    guides(fill=guide_legend(nrow=1, 
                             byRow=TRUE, 
                             override.aes=list(size=16)))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised",
                    paste0("sampleAbundanceDistribution_splitBy",fctr,"_cycLoessNormAbundances_childSamples.png")),
         p1,
         units="px", width=5000, height=2750)

  if (fctr == "Condition") {
    
    # Add to list
    newList <- list(p1)
    names(newList) <- c("sampleAbundDistribution_colByCondition_cyclicLoessNorm_validData")
    plotList <- append(plotList,
                       newList)
    
  }
  
}

```

### PCA

#### Run PCA

```{r}

# Run PCA
validPCA <- prcomp(t(normData_noNA[!is.na(rowSums(normData_noNA)),metaCrossSectional_validData$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
validPCA_metaData <- merge(metaCrossSectional_validData[,!grepl("^PC",colnames(metaCrossSectional_validData))], validPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins"), recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:9) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  condLab_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top")) +
          geom_text_repel(data=validPCA_metaData[which(validPCA_metaData$Sample_ID %in% c("PS_DR_095", "PS_DR_110", "PS_DR_111", "PS_DR_112")),],
                          aes(x=.data[[paste0("PC", i)]], 
                              y=.data[[paste0("PC", i + 1)]], 
                              label=Sample_ID),
                          size=12,
                          max.overlaps=10)
  cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlate.cycLoessNorm_validationData_allSamples.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateRow.cycLoessNorm_validationData_allSamples.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateColumn.cycLoessNorm_validationData_allSamples.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByCondition.cycLoessNorm_validationData_allSamples.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByCondition_labBySampleID.cycLoessNorm_validationData_allSamples.png")), condLab_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colBySex.cycLoessNorm_validationData_allSamples.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAge.cycLoessNorm_validationData_allSamples.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "allProteins",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAgeGroup.cycLoessNorm_validationData_allSamples.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePCA_colByCondition_validData",
                    "normAbundancePCA_colBySite_validData")
plotList <- append(plotList,
                   newList)

```

### Read in desired proteins

```{r}

# Read in chosen prots
chosenProts <- data.frame(read.table(here::here(resDir, "crossSectional", "discovery", "trainSplitProp0.8", "sampleRandom", "biomarkers", "finalModel",
                                                "finalTargetProteins.tsv"), sep="\t", header=TRUE))
rownames(chosenProts) <- chosenProts$Protein

# Check common proteins
commonProts <- intersect(chosenProts$Protein, rownames(quantTable_validData))
chosenProts[commonProts,]

```

### PCA on multi-group selected proteins

#### Run PCA

```{r}

# Get BModal CNO Discovery proteins
discoveryProts_qntTable <- normData_noNA[commonProts,]

# Run PCA
validPCA <- prcomp(t(discoveryProts_qntTable[!is.na(rowSums(discoveryProts_qntTable)),metaCrossSectional_validData$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
validPCA_metaData <- merge(metaCrossSectional_validData[,!grepl("^PC",colnames(metaCrossSectional_validData))], validPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup"), recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:7) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlate.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateRow.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateColumn.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByCondition.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colBySex.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAge.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "multiGroup",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAgeMultiGroupSelected.cycLoessNorm_validationData_allSamples_justMultiGroupSelectedProtein.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePCA_multiGroupProts_colByCondition_validData",
                    "normAbundancePCA_multiGroupProts_colBySite_validData")
plotList <- append(plotList,
                   newList)

```

### PCA on bi-modal selected proteins

#### Run PCA

```{r}

# Get CNO-specific models
biModalCommonProts <- intersect(chosenProts[chosenProts$Model == "Both" | chosenProts$Model == "BiModal",]$Protein, 
                                rownames(quantTable_validData))

# Get BModal CNO Discovery proteins
discoveryProts_qntTable <- normData_noNA[intersect(rownames(normData_noNA), biModalCommonProts),]

# Run PCA
validPCA <- prcomp(t(discoveryProts_qntTable[!is.na(rowSums(discoveryProts_qntTable)),metaCrossSectional_validData$Sample_ID]), center=TRUE, scale=TRUE)

# Merge with metadata
validPCA_metaData <- merge(metaCrossSectional_validData[,!grepl("^PC",colnames(metaCrossSectional_validData))], validPCA$x, by.x="Sample_ID", by.y="row.names")

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal"), recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:3) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=36)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlate.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateRow.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByPlateColumn.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByCondition.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colBySex.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAge.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "pca", "justCandidateProteins", "BiModal",
                    paste0("PCA_biplot.pc",i,"to",i+1,".colByAgeBiModalCNOSelected.cycLoessNorm_validationData_allSamples_justBiModalCNOSelectedProtein.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=validPCA_metaData, var=get_eigenvalue(validPCA), color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePCA_biModalProts_colByCondition_validData",
                    "normAbundancePCA_biModalProts_colBySite_validData")
plotList <- append(plotList,
                   newList)

```

### Probabilistic PCA 

Base off of: https://www.bioconductor.org/packages/release/bioc/vignettes/pcaMethods/inst/doc/pcaMethods.pdf

quantTable_crossSectional \<- data.frame(PG.ProteinAccessions=rownames(quantTable_crossSectional), quantTable_crossSectional\[,c(crossPropSmpls\[crossPropSmpls$PROP_MISSING < maxPropMissing,]$ID)\]) quantTable_longitudinal \<- data.frame(PG.ProteinAccessions=rownames(quantTable_longitudinal), quantTable_longitudinal\[,c(longPropSmpls\[longPropSmpls$PROP_MISSING < maxPropMissing,]$ID)\])

#### Run PCA

```{r}

# Fit probabilistic PCA
set.seed(35353)
resPPCA <- pcaMethods::pca(t(normData[,metaCrossSectional_validData$Sample_ID]), 
                           method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
resPPCA_data <- merge(metaCrossSectional_validData, pcaMethods::scores(resPPCA), by.x="Sample_ID", "row.names")

# Get variance explained
varExpl <- list(variance.percent=(sDev(resPPCA) / sum(sDev(resPPCA)))  * 100)

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins"),
           recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:9) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  condLab_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top")) +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top")) +
          geom_text_repel(data=resPPCA_data[which(resPPCA_data$Sample_ID %in% c("PS_DR_095", "PS_DR_110", "PS_DR_111", "PS_DR_112")),],
                          aes(x=.data[[paste0("PC", i)]], 
                              y=.data[[paste0("PC", i + 1)]], 
                              label=Sample_ID),
                          size=12,
                          max.overlaps=10)
  sex_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlate.validationData_allSamples.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateRow.validationData_allSamples.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateColumn.validationData_allSamples.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByCondition.validationData_allSamples.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByCondition_labBySampleID.validationData_allSamples.png")), condLab_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colBySex.validationData_allSamples.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAge.validationData_allSamples.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "allProteins",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAgeGroup.validationData_allSamples.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePPCA_colByCondition_validData",
                    "normAbundancePPCA_colBySite_validData")
plotList <- append(plotList,
                   newList)

```

### Probabilistic PCA: Mulit-group proteins

#### Run PCA

```{r}

# Fit probabilistic PCA
set.seed(35353)
resPPCA <- pcaMethods::pca(t(normData[commonProts,metaCrossSectional_validData$Sample_ID]), 
                           method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
resPPCA_data <- merge(metaCrossSectional_validData, pcaMethods::scores(resPPCA), by.x="Sample_ID", "row.names")

# Get variance explained
varExpl <- list(variance.percent=(sDev(resPPCA) / sum(sDev(resPPCA)))  * 100)

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup"),
           recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:9) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: Multi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlate.validationData_allSamples_justMultiGroupSelectedProteins.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateRow.validationData_allSamples_justMultiGroupSelectedProteins.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateColumn.validationData_allSamples_justMultiGroupSelectedProteins.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByCondition.validationData_allSamples_justMultiGroupSelectedProteins.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colBySex.validationData_allSamples_justMultiGroupSelectedProteins.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAge.validationData_allSamples_justMultiGroupSelectedProteins.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "multiGroup",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAgeGroup.validationData_allSamples_justMultiGroupSelectedProteins.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePPCA_multiGroupProts_colByCondition_validData",
                    "normAbundancePPCA_multiGroupProts_colBySite_validData")
plotList <- append(plotList,
                   newList)

```


### Probabilistic PCA: BiModal proteins

#### Run PCA

```{r}

# Fit probabilistic PCA
set.seed(35353)
resPPCA <- pcaMethods::pca(t(normData[biModalCommonProts,metaCrossSectional_validData$Sample_ID]), 
                           method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
resPPCA_data <- merge(metaCrossSectional_validData, pcaMethods::scores(resPPCA), by.x="Sample_ID", "row.names")

# Get variance explained
varExpl <- list(variance.percent=(sDev(resPPCA) / sum(sDev(resPPCA)))  * 100)

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal"),
           recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:4) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Plate") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  row_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="ROW", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Row") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  col_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="COLUMN", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Column") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  ageGroup_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Age_Group", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=ageGroupCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Age Group") +
    ggtitle("Validation Cohort: Bi-Modal Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlate.validationData_allSamples_justMultiGroupSelectedProteins.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateRow.validationData_allSamples_justMultiGroupSelectedProteins.png")), row_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlateColumn.validationData_allSamples_justMultiGroupSelectedProteins.png")), col_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByCondition.validationData_allSamples_justMultiGroupSelectedProteins.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colBySex.validationData_allSamples_justMultiGroupSelectedProteins.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAge.validationData_allSamples_justMultiGroupSelectedProteins.png")), age_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "probPCA", "justCandidateProteins", "biModal",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAgeGroup.validationData_allSamples_justMultiGroupSelectedProteins.png")), ageGroup_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r}

# Plot
site_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, shape_by="Age_Group", color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Site") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=resPPCA_data, var=varExpl, color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(shape="Age Group") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt)
names(newList) <- c("normAbundancePPCA_biModalProts_colByCondition_validData",
                    "normAbundancePPCA_biModalProts_colBySite_validData")
plotList <- append(plotList,
                   newList)

```

### Combine validation/BModal CNO Discovery

```{r}


# Combine data
quantTable_combStudy <- merge(cycLoessNorm,
                              quantTable_discoData,
                              by="row.names")
rownames(quantTable_combStudy) <- quantTable_combStudy$Row.names
quantTable_combStudy <- quantTable_combStudy[,!grepl("Row[.]names", colnames(quantTable_combStudy))]

```

#### PPCA on data

```{r}

# Run PCA
set.seed(35353)
combPCA <- pcaMethods::pca(t(quantTable_combStudy), 
                           method="ppca", center=TRUE, scale="pareto", cv="q2", nPcs=10)

# Combine with metadata
combPCA_metaData <- merge(rbind(metaCrossSectional_validData[,c("Sample_ID","Age","Sex","PLATE","Condition","Site", "Data_Group")],
                                metaCrossSectional[metaCrossSectional$Age_Group == "Child",c("Sample_ID","Age","Sex","PLATE","Condition","Site", "Data_Group")]), 
                           scores(combPCA), by.x="Sample_ID", by.y="row.names")
combPCA_metaData[combPCA_metaData$Data_Group == "TrainTest",]["Data_Group"] <- "Discovery"

# Get variance explained
varExpl <- list(variance.percent=(sDev(combPCA) / sum(sDev(combPCA)))  * 100)

```

#### Plot

```{r}

# Create output directories
dir.create(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "combined"),
           recursive=TRUE, showWarnings=FALSE)

# Loop over PCs
## All together
for (i in 1:9) {
  
  # Plot
  # Technical
  plate_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="PLATE", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=plateCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    labs(fill="Plate") +
    ggtitle("Validation & Discovery\nCohorts: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  ## Biological
  cond_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Condition", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation & Discovery\nCohorts: All Proteins") +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  sex_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Sex", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_manual(values=sexCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation & Discovery\nCohorts: All Proteins") +
    guides(fill=guide_legend(ncol=1,title.position="top", override.aes=list(shape=21)), shape=guide_legend(ncol=1,title.position="top"))
  age_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Age", label_id="Sample_ID", pcs=c(i, i+1), pch_size=7) + 
    scale_fill_continuous(high="darkred", low="white") +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    ggtitle("Validation & Discovery\nCohorts: All Proteins") +
    guides(fill=guide_colourbar(barheight=10, title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
  
  # Save
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "combined",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByPlate.validationData_allSamples.png")), plate_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "combined",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByCondition.validationData_allSamples.png")), cond_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "combined",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colBySex.validationData_allSamples.png")), sex_plt, width=6000, height=4000, units="px")
  ggsave(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised", "combined",
                    paste0("probPCA_biplot.pc",i,"to",i+1,".colByAge.validationData_allSamples.png")), age_plt, width=6000, height=4000, units="px")
  
}

```

#### Plot required for paper

```{r, fig.width=32, fig.height=32}

# Plot
site_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Site", label_id="Sample_ID", pcs=c(1, 2), pch_size=7) + 
    scale_shape_manual(values=c(21, 24)) + 
    scale_fill_manual(values=siteCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
cond_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Condition", label_id="Sample_ID", pcs=c(1,2), pch_size=7) + 
    scale_fill_manual(values=tidyCondCols) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.title=element_text(size=42),
          plot.title=element_text(size=64)) +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))
study_plt <- Plot2DPCA(meta=combPCA_metaData, var=varExpl, color_by="Data_Group", pcs=c(1,2), pch_size=12) + 
    scale_fill_manual(values=c("Validation"="salmon",
                               "Discovery"="firebrick4")) +
    theme_classic(base_size=24) +
    theme(legend.position="right", 
          legend.direction="vertical",
          legend.title=element_text(size=32),
          legend.text=element_text(size=24),
          axis.text=element_text(size=64),
          axis.title=element_text(size=72),
          plot.title=element_text(size=64)) +
    guides(fill=guide_legend(ncol=1,override.aes=list(shape=21), title.position="top"), shape=guide_legend(ncol=1,title.position="top"))

# Add to list
newList <- list(site_plt,
                cond_plt,
                study_plt)
names(newList) <- c("normAbundancePPCA_colBySite_combValidDiscoData",
                    "normAbundancePPCA_colByCondition_combValidDiscoData",
                    "normAbundancePPCA_colByStudyCohort_combValidDiscoData")
plotList <- append(plotList,
                   newList)

study_plt

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "12_validationDataNormEDA_figs.rds"))

```




