---
title: "Validation Data Differential Abundance"
format: revealjs
editor: source
---

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
here::i_am("14_validationDataDiffAbundance.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

```

### Read in plot list

```{r}

# Save as RDS
plotList <- readRDS(here::here(figDir, 
                    "13_validationDataImputation_figs.rds"))
 
```

## Load data

### Metadata

```{r}

# Source discovery study metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata.R"))

# Read in pivot table
## Read in protein abundance table of actual samples
metaCrossSectional_validData <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "filtered",
                                                     "preSchool_validationData_tallFormatClean_samplesFiltered_byMissingness.tsv"), sep="\t", header=TRUE) 
rownames(metaCrossSectional_validData) <- metaCrossSectional_validData$Sample_ID

# Keep just child cross-sectional
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Age_Group == "Child",]
table(metaCrossSectional_validData$Condition)

# Exclude JIA
metaCrossSectional_validData <- metaCrossSectional_validData[metaCrossSectional_validData$Condition != "JIA",]
table(metaCrossSectional_validData$Condition)

# Read in additional metadata (with dates)
addMetaValid <- read.table(here::here(metaDir, "preSchool_additionalMetadata_commonToCrossSectional_validationData.tsv"), 
                           sep="\t", header=TRUE)

# Keep child drop JIA
addMetaValid <- addMetaValid[addMetaValid$Age_Group == "Child" & addMetaValid$Condition != "JIA",]
table(addMetaValid$Condition)

# Scale
addMetaValid <- addMetaValid[!is.na(addMetaValid$Age) & !is.na(addMetaValid$Date.of.sample),]
addMetaValid["Z_Age"] <- scale(addMetaValid$Age, scale=TRUE, center=TRUE)
addMetaValid["Z_Date"] <- scale(as.numeric(as.Date(addMetaValid$Date.of.sample)), scale=TRUE, center=TRUE)

# Drop filtered samples
addMetaValid <- addMetaValid[which(addMetaValid$Sample_ID %in% metaCrossSectional_validData$Sample_ID),]

```

### Read in abundance data

```{r}

# Read in normalised data
qntTableNorm <- read.table(here::here(validDir, "DIAProteomics", "directDIA_spectronautProcessed", "normalised",
                                      "validationData_quantTable_normalisedCyclicLoess.tsv"), sep="\t", header=TRUE)
rownames(qntTableNorm) <- qntTableNorm$ProteinGroup

```

### Read in desired proteins

```{r}

# Read in chosen proteins
finalTargets <- read.table(here::here(resCrossDiscoDir, "trainSplitProp0.8", "sampleRandom", "biomarkers", "finalModel",
                                      "finalTargetProteins.tsv"), sep="\t", header=TRUE)
rownames(finalTargets) <- finalTargets$Protein

# Get proteins present in all sets
bestProts <- finalTargets[finalTargets == "No" & finalTargets$inDateAdjustedSet == "Yes",]$Protein

# Check common proteins
commonProts <- intersect(finalTargets$Protein,
                         rownames(qntTableNorm))

```

### Read in train/test data

```{r}

# Read in
childLimmaRes_ANOVACondition_trainTest <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                     "limmaResults_ANOVAwrtCondition_crossSectionalData.tsv"), sep="\t", header=TRUE)
childLimmaRes_ANOVACondition_dresden_trainTest <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                             "limmaResults_ANOVAwrtCondition_crossSectionalData_justDresdenSamples.tsv"), sep="\t", header=TRUE)

childLimmaRes_vsCNO_ANOVA_trainTest <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                  "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData.tsv"), sep="\t", header=TRUE)
childLimmaRes_vsCNO_ANOVA_dresden_trainTest <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                          "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData_justDresdenSamples.tsv"), sep="\t", header=TRUE)

```

### Source colours

```{r}

source(here::here("params", "colors.R"))

```

### Perform all vs all

#### Subset metadata

```{r}

# Subset to site
metaCrossSectional_validData_dresden <- metaCrossSectional_validData[metaCrossSectional_validData$Site == "Dresden",]

```

#### Scale age

```{r}

# Scale age
metaCrossSectional_validData["Z_Age"] <- scale(metaCrossSectional_validData$Age, center=TRUE, scale=TRUE)
metaCrossSectional_validData_dresden["Z_Age"] <- scale(metaCrossSectional_validData_dresden$Age, center=TRUE, scale=TRUE)

```

#### Define models

```{r}

# Define formulas for models
childValid_modelFormulas <- list(childValidFormulas_condition=" ~ Condition + Z_Age + PLATE + Sex",
                                 childValidFormulas_site=" ~ Site + Z_Age + PLATE + Sex",
                                 childValidFormula_condition_noIntercept=" ~ 0 + Condition + Z_Age + Sex + PLATE")
childValidDate_modelFormulas <- list(childValidFormulas_condition=" ~ Condition + Z_Age + PLATE + Z_Date + Sex",
                                     childValidFormulas_site=" ~ Site + Z_Age + PLATE + Z_Date + Sex",
                                     childValidFormula_condition_noIntercept=" ~ 0 + Condition + Z_Age + Sex + Z_Date + PLATE")

# Make designs
childValidDesign         <- model.matrix(as.formula(childValid_modelFormulas$childValidFormulas_condition), 
                                         data=metaCrossSectional_validData)
childValidDesignSite     <- model.matrix(as.formula(childValid_modelFormulas$childValidFormulas_site), 
                                         data=metaCrossSectional_validData)

childValidDesign_noInt   <- model.matrix(as.formula(childValid_modelFormulas$childValidFormula_condition_noIntercept), 
                                         data=metaCrossSectional_validData)
childValidesign_noInt_dresden <- model.matrix(as.formula(childValid_modelFormulas$childValidFormula_condition_noIntercept), 
                                         data=metaCrossSectional_validData_dresden)

rownames(addMetaValid) <- addMetaValid$Sample_ID
childValidDateDesign_noInt   <- model.matrix(as.formula(childValidDate_modelFormulas$childValidFormula_condition_noIntercept), 
                                             data=addMetaValid)
childValidDatesign_noInt_dresden <- model.matrix(as.formula(childValidDate_modelFormulas$childValidFormula_condition_noIntercept), 
                                                 data=addMetaValid[addMetaValid$Site == "Dresden",])

# Make contrasts
childValid_contrastM <- makeContrasts(CNOvsCrohn   = ConditionCNO-ConditionCrohn,
                                      CNOvsOnc      = ConditionCNO-ConditionOncology,
                                      CNOvsHealthy  = ConditionCNO-ConditionHealthy,
                                      CNOvsInfOsteo = ConditionCNO-ConditionInfOsteitis,
                                      levels=childValidDesign_noInt)
  
# Make contrasts
childValidDate_contrastM <- makeContrasts(CNOvsCrohn   = ConditionCNO-ConditionCrohn,
                                          CNOvsOnc      = ConditionCNO-ConditionOncology,
                                          CNOvsHealthy  = ConditionCNO-ConditionHealthy,
                                          CNOvsInfOsteo = ConditionCNO-ConditionInfOsteitis,
                                          levels=childValidDateDesign_noInt)

childValid_contrastM <- makeContrasts(CNOvsCrohn    = ConditionCNO-ConditionCrohn,
                                      CNOvsOnc      = ConditionCNO-ConditionOncology,
                                      CNOvsInfOsteo = ConditionCNO-ConditionInfOsteitis,
                                      CNOvsHealthy  = ConditionCNO-ConditionHealthy,
                                      levels=childValidDesign_noInt)
childValid_contrastM_CNOvsIOH <- makeContrasts(CNOvsInfOsteo = ConditionCNO-ConditionInfOsteitis,
                                               CNOvsHealthy = ConditionCNO-ConditionHealthy,
                                               levels=childValidDesign_noInt)
childValid_contrastM_dresden <- makeContrasts(CNOvsCrohn   = ConditionCNO-ConditionCrohn,
                                              CNOvsOnc      = ConditionCNO-ConditionOncology,
                                              levels=childValidesign_noInt_dresden)
childValid_contrastM_dresdenDate <- makeContrasts(CNOvsCrohn   = ConditionCNO-ConditionCrohn,
                                                  CNOvsOnc      = ConditionCNO-ConditionOncology,
                                                  levels=childValidDatesign_noInt_dresden)

# Make design
childValidDesign <- model.matrix(as.formula(childValid_modelFormulas$childValidFormulas_condition), 
                                  data=metaCrossSectional_validData)

# Check rank
is.fullrank(childValidDesign)
is.fullrank(childValidDesignSite)
is.fullrank(childValidDateDesign_noInt)
is.fullrank(childValidDesign_noInt)
is.fullrank(childValidesign_noInt_dresden)

```

#### Test

```{r}

# Define levels to test
levels2Test <- paste0("Condition",unique(metaCrossSectional_validData$Condition))
levels2Test_site <- paste0("Site",unique(metaCrossSectional_validData$Site))

# Run Limma models
childLimmaRes_ANOVASite              <- runLimma(qntTableNorm[commonProts,rownames(childValidDesignSite)], 
                                                 designM=childValidDesignSite,
                                                 levels2Test=levels2Test_site[which(levels2Test_site %in% colnames(childValidDesignSite))])
childLimmaRes_ANOVACondition         <- runLimma(qntTableNorm[commonProts,rownames(childValidDesign)], 
                                                 designM=childValidDesign,
                                                 levels2Test=levels2Test[which(levels2Test %in% colnames(childValidDesign))])

# No intercept (CNO contrasts)
childLimmaRes_vsCNO_ANOVA_Condition  <- runLimma(qntTableNorm[commonProts,rownames(childValidDesign_noInt)], 
                                                 designM=childValidDesign_noInt,
                                                 levels2Test=colnames(childValid_contrastM),
                                                 contrasts2Fit=childValid_contrastM)
childLimmaRes_vsCNO_ANOVA_Condition_Dresden  <- runLimma(qntTableNorm[commonProts,rownames(childValidesign_noInt_dresden)], 
                                                         designM=childValidesign_noInt_dresden,
                                                         levels2Test=colnames(childValid_contrastM_dresden),
                                                         contrasts2Fit=childValid_contrastM_dresden)

# With date
childLimmaRes_vsCNO_ANOVA_Condition_date  <- runLimma(qntTableNorm[commonProts,
                                                                   rownames(childValidDateDesign_noInt)], 
                                                      designM=childValidDateDesign_noInt,
                                                      levels2Test=colnames(childValidDate_contrastM),
                                                      contrasts2Fit=childValidDate_contrastM)
childLimmaRes_vsCNO_ANOVA_Condition_Dresden_date  <- runLimma(qntTableNorm[commonProts,
                                                                           rownames(childValidDatesign_noInt_dresden)], 
                                                              designM=childValidDatesign_noInt_dresden,
                                                              levels2Test=colnames(childValid_contrastM_dresdenDate),
                                                              contrasts2Fit=childValid_contrastM_dresdenDate)


# Get sig
siteProt <- childLimmaRes_ANOVASite[childLimmaRes_ANOVASite$adj.P.Val < 0.05,] # Just CRP apparently associated with site.
childLimmaRes_ANOVACondition_sig <- childLimmaRes_ANOVACondition[childLimmaRes_ANOVACondition$adj.P.Val < 0.05,]

childLimmaRes_vsCNO_ANOVA_Condition_sig <- childLimmaRes_vsCNO_ANOVA_Condition[childLimmaRes_vsCNO_ANOVA_Condition$adj.P.Val < 0.05,]
childLimmaRes_vsCNO_ANOVA_Condition_Dresden_sig <- childLimmaRes_vsCNO_ANOVA_Condition_Dresden[childLimmaRes_vsCNO_ANOVA_Condition_Dresden$adj.P.Val < 0.05,]

childLimmaRes_vsCNO_ANOVA_Condition_date_sig <- childLimmaRes_vsCNO_ANOVA_Condition_date[childLimmaRes_vsCNO_ANOVA_Condition_date$adj.P.Val < 0.05,]
childLimmaRes_vsCNO_ANOVA_Condition_Dresden_date_sig <- childLimmaRes_vsCNO_ANOVA_Condition_Dresden_date[childLimmaRes_vsCNO_ANOVA_Condition_Dresden_date$adj.P.Val < 0.05,]

# Overlap selected proteins
finalTargets[intersect(childLimmaRes_ANOVACondition_sig$PG_ID, finalTargets$Protein),]

finalTargets[intersect(childLimmaRes_vsCNO_ANOVA_Condition_sig$PG_ID, finalTargets$Protein),]
finalTargets[intersect(childLimmaRes_vsCNO_ANOVA_Condition_Dresden_sig$PG_ID, finalTargets$Protein),]

finalTargets[intersect(childLimmaRes_vsCNO_ANOVA_Condition_date_sig$PG_ID, finalTargets$Protein),]
finalTargets[intersect(childLimmaRes_vsCNO_ANOVA_Condition_Dresden_date_sig$PG_ID, finalTargets$Protein),]

# Overlap all proteins
length(intersect(childLimmaRes_vsCNO_ANOVA_Condition_sig$PG_ID, 
                 childLimmaRes_vsCNO_ANOVA_trainTest[childLimmaRes_vsCNO_ANOVA_trainTest$adj.P.Val < 0.05,]$PG_ID))

# Chcek between date adjusted and non-date adjusted
## What is in the non date-adjusted that is not in the date adjusted set?
setdiff(childLimmaRes_vsCNO_ANOVA_Condition_sig$PG_ID,
        childLimmaRes_vsCNO_ANOVA_Condition_date_sig$PG_ID)
## What is in date-adjusted that is not in the non-date adjusted?
setdiff(childLimmaRes_vsCNO_ANOVA_Condition_date_sig$PG_ID,
        childLimmaRes_vsCNO_ANOVA_Condition_sig$PG_ID)

```

#### Pairwise tests

```{r}

# Loop over possible contratss
allPW_limmaRes <- data.frame()
for (contrastName in colnames(childValid_contrastM)) {

  # Test
  limmaRes <- runLimma(qntTableNorm[commonProts,rownames(childValidDesign_noInt)], 
                       designM=childValidDesign_noInt,  
                       levels2Test=contrastName,
                       contrasts2Fit=childValid_contrastM)
  limmaRes["CONTRAST"] <- contrastName
  allPW_limmaRes <- rbind(allPW_limmaRes, limmaRes)
  
}
iohPW_limmaRes <- data.frame()
for (contrastName in colnames(childValid_contrastM_CNOvsIOH)) {

  # Test
  limmaRes <- runLimma(qntTableNorm[commonProts,rownames(childValidDesign_noInt)], 
                       designM=childValidDesign_noInt,  
                       levels2Test=contrastName,
                       contrasts2Fit=childValid_contrastM_CNOvsIOH)
  limmaRes["CONTRAST"] <- contrastName
  iohPW_limmaRes <- rbind(iohPW_limmaRes, limmaRes)
  
}

# Enumerate hits
## Padj
nConsistentProts_allPW_padj <- c(table(allPW_limmaRes[allPW_limmaRes$adj.P.Val < candidatePThr,]$PG_ID))
nConsistentProts_iohPW_padj <- c(table(iohPW_limmaRes[iohPW_limmaRes$adj.P.Val < candidatePThr,]$PG_ID))
consitentProts_allPW_padj <- names(nConsistentProts_allPW_padj[nConsistentProts_allPW_padj == ncol(childValid_contrastM)])
consitentProts_iohPW_padj <- names(nConsistentProts_iohPW_padj[nConsistentProts_iohPW_padj == ncol(childValid_contrastM_CNOvsIOH)])

## PVal
nConsistentProts_allPW_pval <- c(table(allPW_limmaRes[allPW_limmaRes$P.Value < candidatePThr,]$PG_ID))
nConsistentProts_iohPW_pval <- c(table(iohPW_limmaRes[iohPW_limmaRes$P.Value < candidatePThr,]$PG_ID))
consitentProts_allPW_pval <- names(nConsistentProts_allPW_pval[nConsistentProts_allPW_pval == ncol(childValid_contrastM)])
consitentProts_iohPW_pval <- names(nConsistentProts_iohPW_pval[nConsistentProts_iohPW_pval == ncol(childValid_contrastM_CNOvsIOH)])

```

#### Add column indicating protein selected as a biomarker or not

```{r}

# Add label
childLimmaRes_vsCNO_ANOVA_Condition_Dresden["SelectedProtein"] <- "No"
childLimmaRes_vsCNO_ANOVA_Condition_Dresden[which(childLimmaRes_vsCNO_ANOVA_Condition_Dresden$PG_ID %in% commonProts),]["SelectedProtein"] <- "Yes"
childLimmaRes_ANOVACondition_trainTest["SelectedProtein"] <- "No"
childLimmaRes_ANOVACondition_trainTest[which(childLimmaRes_ANOVACondition_trainTest$PG_ID %in% commonProts),]["SelectedProtein"] <- "Yes"

```

#### Add protein names

```{r}

childLimmaRes_vsCNO_ANOVA_Condition_Dresden <- merge(finalTargets, 
                                             childLimmaRes_vsCNO_ANOVA_Condition_Dresden,
                                             by.x="Protein", by.y="PG_ID")
childLimmaRes_vsCNO_ANOVA_trainTest <- merge(finalTargets, 
                                             childLimmaRes_vsCNO_ANOVA_trainTest,
                                             by.x="Protein", by.y="PG_ID")

```

#### Compare

```{r}

calcOrthoganolDist <- function(train, valid) {
  
  # Check if both vectors have the same length.
  if (length(train) != length(valid)) {
    stop("The 'train' and 'valid' vectors must be of equal length.")
  }
  
  diffCalc <- function(x, y) { (x - y) - mean(x, y) }
  
  # Calculate the mean deviation
  distance <- mapply(diffCalc, valid, train)
  
  # Return
  return(distance)
}

# Get contrasts
contrasts2Check <- colnames(childValid_contrastM)
selectProts <- commonProts

# Get data 
colnames(childLimmaRes_vsCNO_ANOVA_Condition)[[1]] <- "Protein"
validData <- childLimmaRes_vsCNO_ANOVA_Condition
trainData <- childLimmaRes_vsCNO_ANOVA_trainTest
rownames(validData) <- validData$Protein
rownames(trainData) <- trainData$Protein

# Generate output
compareFCs <- data.frame()

# Loop over
for (i in seq_along(contrasts2Check)) {
    
    print(paste0("// Plotting FCs for ",contrasts2Check[[i]]))
    
    # Get FCs for validation and train/test data
    validFC <- validData[contrasts2Check[[i]]]
    trainFC <- trainData[contrasts2Check[[i]]]
    rownames(validFC) <- validData$Protein
    rownames(trainFC) <- trainData$Protein
    
    # Get common
    consistentProts <- intersect(validData$Protein,
                                 trainData$Protein)
    
    # Subset to common
    validFC <- validFC[consistentProts,1]
    trainFC <- trainFC[consistentProts,1]
    
    # Correlate - spearmans
    corrVal <- cor(validFC, 
                   trainFC, 
                   use="pairwise.complete.obs", 
                   method="spearman")
    
    # Get sig proteins
    sigValidProts <- childLimmaRes_vsCNO_ANOVA_Condition[childLimmaRes_vsCNO_ANOVA_Condition$adj.P.Val < 0.05,]
    rownames(childLimmaRes_vsCNO_ANOVA_Condition) <- childLimmaRes_vsCNO_ANOVA_Condition$Protein
    consistentProtNames <- childLimmaRes_vsCNO_ANOVA_Condition[consistentProts,]$Protein
    
    # Plot
    checkLabelProt <- consistentProts %in% sigValidProts$Protein & consistentProts %in% selectProts
    toPlot <- data.frame(Protein=finalTargets[consistentProtNames,]$SYMBOL,
                         SelectedProteins=factor(ifelse(consistentProts %in% selectProts, "Yes", "No"),
                                                 levels=c("No", "Yes")),
                         SigInValidation=factor(ifelse(checkLabelProt, "FDR < 0.05", "ns"),
                                                levels=c("ns", "FDR < 0.05")),
                         Validation=validFC,
                         TrainTest=trainFC)
    rownames(toPlot) <- consistentProts
    toPlot <- toPlot[consistentProts,]
    
    # Add label
    checkLabelProt[sign(toPlot$Validation) != sign(toPlot$TrainTest)] <- FALSE # If FC < 1
    toPlot["LabelProtein"] <- ifelse(checkLabelProt, gsub(" .*","",toPlot$Protein), NA)
    
    # Calculate orthoganol distance
    toPlot["OrthoDist"] <- NA
    toPlot[toPlot$SelectedProteins == "Yes",]["OrthoDist"] <- calcOrthoganolDist(scale(toPlot[toPlot$SelectedProteins == "Yes",]$Validation, 
                                                                                       center=TRUE, scale=TRUE),
                                                                                 scale(toPlot[toPlot$SelectedProteins == "Yes",]$TrainTest, 
                                                                                       center=TRUE, scale=TRUE))
    
    # Add to dataframe
    compareFCs <- rbind(compareFCs,
                        cbind(CONTRAST=contrasts2Check[[i]], 
                              FacetLabel=contrasts2Check[[i]],
                              toPlot))
  
}

# Simplify protein name
compareFCs["ProteinSimple"] <- gsub(" .*","",compareFCs$Protein)

# Add if discovery/valid FC agree
compareFCs["SignCheck"] <- "No"
compareFCs[sign(compareFCs$Validation) == sign(compareFCs$TrainTest),]["SignCheck"] <- "Yes"

# Replace 
compareFCs["Condition"] <- gsub("CNOvs","",compareFCs$CONTRAST)
compareFCs["Condition"] <- gsub("InfOsteo","InfOsteitis",compareFCs$Condition)
compareFCs["Condition"] <- gsub("Onc","Oncology",compareFCs$Condition)
compareFCs["FacetLabel"] <- gsub("CNOvs","",compareFCs$FacetLabel)
compareFCs["FacetLabel"] <- gsub("InfOsteo","InfOsteitis",compareFCs$FacetLabel)
compareFCs["FacetLabel"] <- gsub("Onc","Oncology",compareFCs$FacetLabel)

```

#### Plot scatterplot

```{r, fig.height=5.25, fig.width=16}
  
# Get max value
maxVal <- ceiling(base::max(abs(c(compareFCs$TrainTest, 
                                  compareFCs$Validation))))
  
# Plot
plt <- ggplot() +
    geom_segment(data = compareFCs[compareFCs$SelectedProteins == "Yes",],
                 mapping = aes(x = Validation,
                               y = TrainTest,
                               xend = (Validation + TrainTest) / 2,
                               yend = (Validation + TrainTest) / 2),
                 color = "darkgrey",
                 # linetype = "dashed",
                 size = 1,
                 alpha = 0.7) +
    geom_point(data=compareFCs,
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               alpha=0.65,
               color="darkgrey") +
    geom_point(data=compareFCs[compareFCs$SelectedProteins == "Yes",],
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               alpha=0,
               shape=21,
               fill="white") +
    geom_point(data=compareFCs[compareFCs$SelectedProteins == "Yes",],
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               alpha=0.75,
               shape=21,
               fill="mediumblue") +
    geom_vline(xintercept=0,
               
               color="darkgrey") +
    geom_hline(yintercept=0,
               
               color="darkgrey") +
    geom_abline(slope=1, 
                intercept=0,
                
                color="darkred") +
    geom_text_repel(data=compareFCs[compareFCs$Protein == "CRP",],
                    mapping=aes(x=Validation, 
                                y=TrainTest,
                                label=LabelProtein),
                    size=20,
                    nudge_x=3,
                    force=2.5,
                    color="black",
                    seed=1) +
    facet_grid(cols=vars(FacetLabel)) +
    scale_size_manual(values=c("ns"=2, 
                               "FDR < 0.05"=8)) +
    scale_fill_manual(values=c("ns"="darkgrey", 
                               "FDR < 0.05"="mediumblue")) +
    ylab("Discovery") + xlab("Validation") +
    xlim(-maxVal, maxVal) +
    ylim(-maxVal, maxVal) +
    theme_classic(base_size=24) +
    theme(legend.position="top",
          legend.direction="horizontal",
          legend.title.position="left",
          legend.text=element_text(size=56),
          legend.title=element_text(size=64),
          axis.title=element_text(size=80),
          axis.text=element_text(size=52),
          strip.text=element_text(size=64))
  
# Save
ggsave(here::here(resValidDir, "differentialAbundance",
                  "differentialAbundance_discoveryVsValidation_justCRP.png"),
       plt,
       units="px",
       width=10000,
       height=3250)

# Add to list
newList <- list(plt)
names(newList) <- c("diffAbundanceComparison_justCRP")
plotList <- append(plotList,
                   newList)

```

#### Plot scatterplot excluding CRP

```{r, fig.height=5.25, fig.width=16}
  
# Get max value
maxVal <- 3
  
# Plot
plt <- ggplot() +
    geom_segment(data = compareFCs[compareFCs$SelectedProteins == "Yes" &
                                   compareFCs$Protein != "CRP",],
                 mapping = aes(x = Validation,
                               y = TrainTest,
                               xend = (Validation + TrainTest) / 2,
                               yend = (Validation + TrainTest) / 2),
                 color = "darkgrey",
                 size = 1,
                 alpha = 0.7) +
    geom_point(data=compareFCs[compareFCs$Protein != "CRP",],
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               alpha=0.65,
               color="darkgrey") +
    geom_point(data=compareFCs[compareFCs$SelectedProteins == "Yes" &
                               compareFCs$Protein != "CRP",],
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               shape=21,
               fill="white") +
    geom_point(data=compareFCs[compareFCs$SelectedProteins == "Yes" &
                               compareFCs$Protein != "CRP",],
               mapping=aes(x=Validation, 
                           y=TrainTest,
                           size=SigInValidation),
               alpha=0.75,
               shape=21,
               fill="mediumblue") +
    geom_vline(xintercept=0,
               
               color="darkgrey") +
    geom_hline(yintercept=0,
               
               color="darkgrey") +
    geom_abline(slope=1, 
                intercept=0,
                
                color="darkred") +
    geom_text_repel(data=compareFCs[compareFCs$Protein != "CRP",],
                    mapping=aes(x=Validation, 
                                y=TrainTest,
                                label=LabelProtein),
                    min.segment.length = unit(0, 'lines'),
                    size=7.5,
                    force=75,
                    fill=NA,
                    color="black",
                    seed=7) +
    facet_grid(cols=vars(FacetLabel)) +
    scale_size_manual(values=c("ns"=2, 
                               "FDR < 0.05"=8)) +
    scale_fill_manual(values=c("ns"="darkgrey", 
                               "FDR < 0.05"="mediumblue")) +
    ylab("Discovery") + xlab("Validation") + 
    labs(size="Significant") +
    xlim(-maxVal, maxVal) +
    ylim(-maxVal, maxVal) +
    theme_classic(base_size=24) +
    theme(legend.position="top",
          legend.direction="horizontal",
          legend.title.position="left",
          legend.text=element_text(size=56),
          legend.title=element_text(size=64),
          axis.title=element_text(size=80),
          axis.text=element_text(size=52),
          strip.text=element_text(size=64))
  
# Save
ggsave(here::here(resValidDir, "differentialAbundance",
                  "differentialAbundance_discoveryVsValidation_allCandidatesLabelled.png"),
       plt,
       units="px",
       width=10000,
       height=3250)

# Add to list
newList <- list(plt)
names(newList) <- c("diffAbundanceComparison_noCRP")
plotList <- append(plotList,
                   newList)

```


#### Barbell plot of logFCs

##### Modify data

```{r}

# Make CRP separate 
compareFCs["CRP"] <- " "
compareFCs[compareFCs$Protein == "CRP",]["CRP"] <- "CRP"

# Plot
compareFCs <- compareFCs[order(compareFCs$SigInValidation,
                               decreasing=TRUE),]
compareFCs$Protein <- factor(compareFCs$Protein,
                             levels=unique(compareFCs$Protein))

# Copy
compareFCsCopy <- compareFCs

# Make tall
compareFCsTall <- reshape2::melt(compareFCsCopy,
                                 id.vars=setdiff(colnames(compareFCsCopy),
                                                 c("Validation", "TrainTest")))

# Add significance flag
compareFCsTall$SigInValidation <- as.character(compareFCsTall$SigInValidation)
compareFCsTall[compareFCsTall$SigInValidation == "FDR < 0.05",]["SigInValidation"] <- "FDR < 0.05"
compareFCsTall[compareFCsTall$SigInValidation == "ns",]["SigInValidation"] <- "ns"
compareFCsTall$SigInValidation <- factor(compareFCsTall$SigInValidation,
                                     levels=c("FDR < 0.05", "ns", " "))
compareFCsCopy$SigInValidation <- factor(compareFCsCopy$SigInValidation,
                                     levels=c("FDR < 0.05", "ns", " "))

```

#### Add signficance

```{r}

# Add significance flag
compareFCs$SigInValidation <- as.character(compareFCs$SigInValidation)
compareFCs[compareFCs$SigInValidation == "FDR < 0.05",]["SigInValidation"] <- "*"
compareFCs[compareFCs$SigInValidation == "ns",]["SigInValidation"] <- " "
compareFCs$SigInValidation <- factor(compareFCs$SigInValidation,
                                     levels=c("*", " "))

```


```{r, fig.height=6, fig.width=18}

ggplot(compareFCs,
              aes(y = ProteinSimple,
                  x = abs(OrthoDist),
                  fill = Condition,
                  pattern = SignCheck)) +
  geom_col_pattern(
    pattern_fill = "white",
    pattern_colour = "white",
    colour = "white"
  ) +
  geom_text(data = compareFCs,
            mapping = aes(y = ProteinSimple,
                          x = abs(OrthoDist) + 0.15,
                          label = SigInValidation,
                          size = SigInValidation),
            vjust = 0.75,
            alpha = 1) +
  facet_grid(cols = vars(FacetLabel)) +
  scale_pattern_manual(values = c("No" = "stripe", 
                                  "Yes" = "none"),
                       breaks = c("No", "Yes")) +
   scale_fill_manual(values = condCols) +
  scale_alpha_manual(values = c("Yes" = 1, 
                                "No" = 0.35)) +
  scale_size_manual(values = c(" " = 2, 
                               "*" = 16)) +
  ylab("Selected Protein") +
  xlab("Scaled Distance") +
  labs(fill = "Condition") +
  scale_x_continuous(breaks = c(0, 2, 4),
                     limits  = c(0, 4.5),
                     labels  = c(0, 2, 4))


```

#### Calculate orthoganol distances

```{r, fig.height=5.25, fig.width=16}

# put pattern mapping at top-level
plt <- ggplot(compareFCs[compareFCs$SelectedProteins == "Yes",],
              aes(y = ProteinSimple,
                  x = abs(OrthoDist),
                  fill = Condition,
                  pattern = SignCheck)) +
  geom_col_pattern(
    pattern_fill = "white",
    pattern_colour = "white",
    colour = "white"
  ) +
  geom_text(data = compareFCs,
            mapping = aes(y = ProteinSimple,
                          x = abs(OrthoDist) + 0.25,
                          label = SigInValidation,
                          size = SigInValidation),
            vjust = 0.75,
            alpha = 1) +
  facet_grid(cols = vars(FacetLabel)) +
  scale_pattern_manual(values = c("No" = "stripe", 
                                  "Yes" = "none"),
                       breaks = c("No", "Yes")) +
  scale_fill_manual(values = condCols) +
  scale_alpha_manual(values = c("Yes" = 1, "No" = 0.35)) +
  scale_size_manual(values = c(" " = 2, "*" = 16)) +
  ylab("Selected Protein") +
  xlab("Scaled Distance") +
  labs(fill = "Condition") +
  scale_x_continuous(breaks = c(0, 2, 4),
                     limits  = c(0, 4.5),
                     labels  = c(0, 2, 4)) +
  theme_classic(base_size = 24) +
  theme(axis.title = element_text(size = 80),
        axis.text = element_text(size = 56),
        axis.text.x = element_text(size = 56),
        axis.text.y = element_text(size = 42),
        legend.text = element_text(size = 56),
        legend.title = element_text(size = 64),
        strip.text = element_text(size = 64),
        legend.title.position = "left",
        legend.position = "top",
        legend.key.size = unit(3, "cm")) +
  guides(
  fill = guide_legend(nrow = 1, byRow = TRUE,
                      override.aes = list(size = 30,
                                          pattern="none")),
  pattern = guide_legend(
    override.aes = list(
      pattern = c("stripe", "none"),
      # for legend only: paint the stripe bands black so they are visible
      pattern_fill   = c("white", NA),
      # no border needed on stripe in legend
      pattern_colour = c("white", NA),
      # keep the legend box fill identical to what you used before
      fill   = c("darkgrey", "darkgrey"),
      colour = c(NA, NA)
    )
  ),
  alpha = "none",
  size = "none",
  label = "none"
)


# Save
ggsave(here::here(resValidDir, "differentialAbundance",
                  "scaledOrthoganolDistance_ofCandidates_fromDiaganol.png"),
       plt,
       units="px",
       width=10000,
       height=6000)

# Add to list
newList <- list(plt)
names(newList) <- c("diffAbundanceComparison_orthoganolDistance")
plotList <- append(plotList,
                   newList)

```

##### Plot

```{r, fig.height=24, fig.width=32}

# Adjust colours
condStudyCols <- condCols
names(condStudyCols) <- paste0("Discovery.", names(condStudyCols))
lightCols <- lighten(condCols, 0.65)
names(lightCols) <- paste0("Validation.", names(lightCols))
condStudyCols <- c(condStudyCols, lightCols)

# Add in merged condition/variable
compareFCsTall$variable <- gsub("TrainTest", "Discovery", compareFCsTall$variable)
compareFCsTall["CondData"] <- paste(compareFCsTall$variable, compareFCsTall$Condition, sep=".")

# Get min
compareFCsAggr <- aggregate(as.formula(paste0("value ~ ", 
                                              paste(setdiff(colnames(compareFCsTall),
                                                            c("value", "CondData", "variable")),
                                                    collapse=" + "))),
                            data=compareFCsTall[compareFCsTall$SigInValidation != "ns",],
                            FUN=base::min)

p1 <- ggplot(compareFCsTall[compareFCsTall$Protein != "CRP",],
       aes(x=value,
           y=Protein,
           shape=SignCheck,
           fill=CondData)) +
  geom_vline(xintercept=0,
             linewidth=1,
             color="darkgrey") +
  geom_segment(data=compareFCsCopy[compareFCsCopy$Protein != "CRP",],
               mapping=aes(x=Validation,
                           xend=TrainTest,
                           y=Protein,
                           yend=Protein),
               color="black",
               linewidth=1.5,
               inherit.aes=FALSE) +
  geom_point(color="white",
             fill="white",
             alpha=1,
             size=12) +
  geom_point(stroke=1.5,
             alpha=0.9,
             size=12) +
  scale_shape_manual(values=c("Yes"=21, 
                              "No"=22)) +
  scale_fill_manual(values=condStudyCols) +
  facet_grid(cols=vars(Condition),
             rows=vars(SigInValidation),
             labeller = labeller(SigInValidation=c(No="ns",
                                                   Yes="FDR < 0.05")),
             scale="free",
             space="free_y") +
  scale_x_continuous(limits=c(-2.5, 2.5),
                     breaks=c(-2, -1, 0, 1, 2)) +
  theme_classic(base_size=52)

# Plot
compareFCsTall$variable <- factor(compareFCsTall$variable,
                                  levels=c("Discovery", "Validation"))
p2 <- ggplot(compareFCsTall[compareFCsTall$Protein == "CRP",],
       aes(x=value,
           y=Protein,
           fill=CondData,
           shape=SignCheck)) +
  geom_vline(xintercept=0,
             
             linewidth=1,
             color="darkgrey") +
  geom_segment(data=compareFCsCopy[compareFCsCopy$Protein == "CRP",],
               mapping=aes(x=Validation,
                           xend=TrainTest,
                           y=Protein,
                           yend=Protein),
               color="black",
               linewidth=1.5,
               inherit.aes=FALSE) +
  geom_point(color="white",
             fill="white",
             alpha=1,
             size=12) +
  geom_point(stroke=1.5,
             alpha=0.9,
             size=12) +
  scale_shape_manual(values=c("Yes"=21, 
                              "No"=22)) +
  scale_fill_manual(values=condStudyCols) +
  facet_grid(cols=vars(Condition),
             scale="free_y",
             space="free_y") +
  scale_x_continuous(limits=c(-10.5, 2),
                     breaks=seq(-9, 2, 3),
                     labels=seq(-9, 2, 3)) +
  theme_classic(base_size=52)

# Combine
p1 + p2 + 
  plot_layout(heights=c(1, 0.05),
              guides="collect",
              axis_titles="collect_x") &
  guides(alpha=guide_legend(override.aes=list(shape=21)),
         fill=guide_legend(override.aes=list(shape=21)))

# Add to list
newList <- list(p1,
                p2)
names(newList) <- c("barbellPlot_logFC_discoVsValid",
                    "barbellPlot_logFC_discoVsValid_justCRP")
plotList <- append(plotList,
                   newList)

```


#### Save

```{r}

# Output
dir.create(here::here(resDir, "crossSectional", "validation","differentialAbundance"),
           recursive=TRUE)

# Save
write.table(childLimmaRes_vsCNO_ANOVA_Condition,
            here::here(resDir, "crossSectional", "validation", "differentialAbundance",
                       "limmaResults_ANOVA_wrtCNOvContrasts_validationData.tsv"), sep="\t", quote=FALSE, row.names=FALSE) 

```

### Save plots

```{r}

# Save as RDS
saveRDS(plotList,
        here::here(figDir, 
                   "14_validationDataDiffAbundance_figs.rds"))

```




