---
title: "Biological Contextualisation"
author: "Euan McDonnell"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source parameters, functions and install packages
if (!require(here)) { install.packages("here") }
here::i_am("16_biologicalContextualisation.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

# Metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Source colours
source(here::here("params", "colors.R"))

```

### Read in plot list

```{r}

# Read in validation directory
validResDir <- readRDS(here::here("params", 
                                  paste0("validationAnalysis_outputDirectory_",imputMethod,".rds")))
 
```

### Read in sig protein sets

```{r}

# Read in differential protein sets
allProtSets <- readRDS(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance", 
                       paste0("chosenProteins_featureSelection_pThr",as.character(candidatePThr),"_treatLog2Thr",as.character(log2ThrVal),".rds")))

```

### Read in 

```{r}

# read in 
childLimmaRes_vsCNO_ANOVA_Condition <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                             "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData.tsv"), sep="\t", header=TRUE)
childLimmaRes_vsCNO_ANOVA_Condition_dresden <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                                     "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData.tsv"), sep="\t", header=TRUE)
sigRes <- list(AllSig=childLimmaRes_vsCNO_ANOVA_Condition,
               DresdenOnly=childLimmaRes_vsCNO_ANOVA_Condition_dresden)

```

### Load data

#### Abundance data

```{r}

# Read in cyclic Loess-normalised data
quantTable_allSamples <- read.table(here::here(diaDir, "normalised", 
                                    "metaCrossSectional_NormalyzerDE_byCondition", "CycLoess-normalized.txt"), sep="\t", header=TRUE, row.names=1)

# Chosen imputed data
quantTable_chosenMethod <- read.table(here::here(impDir, paste0("quantTable_imputedWithbPCA.tsv")), sep="\t", header=TRUE, row.names=1)

# Match metadata
trainMetaData <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "biomarkers",
                                       "allSamplesMetadata_takenForSelection.tsv"), sep="\t", header=TRUE)
trainQuantTable <- quantTable_allSamples[,trainMetaData$Sample_ID]

# Get common
quantTable <- quantTable_allSamples[,trainMetaData$Sample_ID]

# Keep only selected metadata
quantTable <- quantTable[,!grepl("QC",colnames(quantTable))]
quantTable_noMissing <- quantTable[!is.na(rowSums(quantTable)),]

```

#### Read in desired proteins

```{r}

# Read in chosen proteins
finalTargets <- read.table(here::here(validResDir,
                                      paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,".tsv")), 
                           sep="\t", 
                           header=TRUE)
rownames(finalTargets) <- finalTargets$Protein

```

#### Get FCs wrt Healthy for target proteins

```{r}

# Get targets
finalTargets_vsHealthy <- merge(finalTargets[,c("Protein",
                                                "SYMBOL",
                                                "GENENAME")], 
                                childLimmaRes_vsCNO_ANOVA_Condition[,c("PG_ID",
                                                                       "CNOvsHealthy")],
                                by.x="Protein",
                                by.y="PG_ID")
colnames(finalTargets_vsHealthy) <- c("Uniprot ID", 
                                      "Gene Symbol",
                                      "Gene Name",
                                      "LogFC (CNO vs Healthy)")
write.table(finalTargets_vsHealthy,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_logFC_CNOvsHealthy.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)

```

## Main

### Functional enrichment analysis

#### Read in pathways

### Functional enrichment analysis

#### Get universe as entrez ID

```{r}

# Get annotations
background <- AnnotationDbi::select(org.Hs.eg.db,
                                   keys=rownames(quantTable_allSamples),
                                   keytype="UNIPROT",
                                   columns=c("ENTREZID",
                                             "SYMBOL"))
backgroundList <- unique(as.character(background$UNIPROT[!is.na(background$UNIPROT)]))
length(backgroundList)

```

#### Read in pathways

```{r}

# Read in reactome pathways
reacPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c2.cp.reactome.v2024.1.Hs.entrez.gmt"))
keggPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c2.cp.kegg_medicus.v2024.1.Hs.entrez.gmt"))
gobpPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c5.go.bp.v2024.1.Hs.entrez.gmt"))

# Add Uniprot IDs
reacPathData <- merge(reacPaths, background, by.x="gene", by.y="ENTREZID")
keggPathData <- merge(keggPaths, background, by.x="gene", by.y="ENTREZID")
gobpPathData <- merge(gobpPaths, background, by.x="gene", by.y="ENTREZID")

# Make list
reacPaths <- lapply(split(reacPathData[,-1]$UNIPROT, reacPathData[,-1]$term),function(x) c(x))
keggPaths <- lapply(split(keggPathData[,-1]$UNIPROT, keggPathData[,-1]$term),function(x) c(x))
gobpPaths <- lapply(split(gobpPathData[,-1]$UNIPROT, gobpPathData[,-1]$term),function(x) c(x))
names(reacPaths) <- mgsub(names(reacPaths), c("REACTOME_", "_"), c("", " "))
names(keggPaths) <- mgsub(names(keggPaths), c("KEGG_", "_"), c("", " "))
names(gobpPaths) <- mgsub(names(gobpPaths), c("GOBP_", "_"), c("", " "))

# Remove caes with none
reacPaths <- reacPaths[sapply(reacPaths, function(term) length(term) > 0)]
keggPaths <- keggPaths[sapply(keggPaths, function(term) length(term) > 0)]
gobpPaths <- gobpPaths[sapply(gobpPaths, function(term) length(term) > 0)]

```

#### Annotate proteins to terms

This is to just save as a table.

```{r}

# Combine
finalTargetsReac <- merge(finalTargets,
                          reacPathData,
                          by.x="ENTREZID",
                          by.y="gene")
finalTargetsKegg <- merge(finalTargets,
                          keggPathData,
                          by.x="ENTREZID",
                          by.y="gene")
finalTargetsGobp <- merge(finalTargets,
                          gobpPathData,
                          by.x="ENTREZID",
                          by.y="gene")

# Check coverage
length(unique(finalTargetsReac$Protein))
length(unique(finalTargetsKegg$Protein))
length(unique(finalTargetsGobp$Protein))

# Save
write.table(finalTargets,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,".tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsReac,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedReactomePathways.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsKegg,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedKeggPathways.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsGobp,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedGoBP.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)

```


