---
title: "Biological Contextualisation"
author: "Euan McDonnell"
format: html
editor: source
---

# Notes

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source parameters, functions and install packages
if (!require(here)) { install.packages("here") }
here::i_am("16_biologicalContextualisation.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

# Metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata_filtered.R"))

# Source colours
source(here::here("params", "colors.R"))

```

### Read in plot list

```{r}

# Read in validation directory
validResDir <- readRDS(here::here("params", 
                                  paste0("validationAnalysis_outputDirectory_",imputMethod,".rds")))
 
```

### Read in sig protein sets

```{r}

# Read in differential protein sets
allProtSets <- readRDS(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance", 
                       paste0("chosenProteins_featureSelection_pThr",as.character(candidatePThr),"_treatLog2Thr",as.character(log2ThrVal),".rds")))

```

### Read in 

```{r}

# read in 
childLimmaRes_vsCNO_ANOVA_Condition <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                             "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData.tsv"), sep="\t", header=TRUE)
childLimmaRes_vsCNO_ANOVA_Condition_dresden <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "differentialAbundance",
                                                                     "limmaResults_ANOVA_wrtCNOvContrasts_crossSectionalData.tsv"), sep="\t", header=TRUE)
sigRes <- list(AllSig=childLimmaRes_vsCNO_ANOVA_Condition,
               DresdenOnly=childLimmaRes_vsCNO_ANOVA_Condition_dresden)

```

### Load data

#### Abundance data

```{r}

# Read in cyclic Loess-normalised data
quantTable_allSamples <- read.table(here::here(diaDir, "normalised", 
                                    "metaCrossSectional_NormalyzerDE_byCondition", "CycLoess-normalized.txt"), sep="\t", header=TRUE, row.names=1)

# Chosen imputed data
quantTable_chosenMethod <- read.table(here::here(impDir, paste0("quantTable_imputedWithbPCA.tsv")), sep="\t", header=TRUE, row.names=1)

# Match metadata
trainMetaData <- read.table(here::here(resCrossDiscoDir, paste0("trainSplitProp",trainSplitProp), "sampleRandom", "biomarkers",
                                       "allSamplesMetadata_takenForSelection.tsv"), sep="\t", header=TRUE)
trainQuantTable <- quantTable_allSamples[,trainMetaData$Sample_ID]

# Get common
quantTable <- quantTable_allSamples[,trainMetaData$Sample_ID]

# Keep only selected metadata
quantTable <- quantTable[,!grepl("QC",colnames(quantTable))]
quantTable_noMissing <- quantTable[!is.na(rowSums(quantTable)),]

```

#### Read in desired proteins

```{r}

# Read in chosen proteins
finalTargets <- read.table(here::here(validResDir,
                                      paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,".tsv")), 
                           sep="\t", 
                           header=TRUE)
rownames(finalTargets) <- finalTargets$Protein

```

#### Get FCs wrt Healthy for target proteins

```{r}

# Get targets
finalTargets_vsHealthy <- merge(finalTargets[,c("Protein",
                                                "SYMBOL",
                                                "GENENAME")], 
                                childLimmaRes_vsCNO_ANOVA_Condition[,c("PG_ID",
                                                                       "CNOvsHealthy")],
                                by.x="Protein",
                                by.y="PG_ID")
colnames(finalTargets_vsHealthy) <- c("Uniprot ID", 
                                      "Gene Symbol",
                                      "Gene Name",
                                      "LogFC (CNO vs Healthy)")
write.table(finalTargets_vsHealthy,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_logFC_CNOvsHealthy.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)

```

## Main

### Functional enrichment analysis

#### Read in pathways

### Functional enrichment analysis

#### Get universe as entrez ID

```{r}

# Get annotations
background <- AnnotationDbi::select(org.Hs.eg.db,
                                   keys=rownames(quantTable_allSamples),
                                   keytype="UNIPROT",
                                   columns=c("ENTREZID",
                                             "SYMBOL"))
backgroundList <- unique(as.character(background$UNIPROT[!is.na(background$UNIPROT)]))
length(backgroundList)

```

#### Read in pathways

```{r}

# Read in reactome pathways
reacPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c2.cp.reactome.v2024.1.Hs.entrez.gmt"))
keggPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c2.cp.kegg_medicus.v2024.1.Hs.entrez.gmt"))
gobpPaths <- clusterProfiler::read.gmt(here::here(baseDataDir, "annotation", "ontology", "c5.go.bp.v2024.1.Hs.entrez.gmt"))

# Add Uniprot IDs
reacPathData <- merge(reacPaths, background, by.x="gene", by.y="ENTREZID")
keggPathData <- merge(keggPaths, background, by.x="gene", by.y="ENTREZID")
gobpPathData <- merge(gobpPaths, background, by.x="gene", by.y="ENTREZID")

# Make list
reacPaths <- lapply(split(reacPathData[,-1]$UNIPROT, reacPathData[,-1]$term),function(x) c(x))
keggPaths <- lapply(split(keggPathData[,-1]$UNIPROT, keggPathData[,-1]$term),function(x) c(x))
gobpPaths <- lapply(split(gobpPathData[,-1]$UNIPROT, gobpPathData[,-1]$term),function(x) c(x))
names(reacPaths) <- mgsub(names(reacPaths), c("REACTOME_", "_"), c("", " "))
names(keggPaths) <- mgsub(names(keggPaths), c("KEGG_", "_"), c("", " "))
names(gobpPaths) <- mgsub(names(gobpPaths), c("GOBP_", "_"), c("", " "))

# Remove caes with none
reacPaths <- reacPaths[sapply(reacPaths, function(term) length(term) > 0)]
keggPaths <- keggPaths[sapply(keggPaths, function(term) length(term) > 0)]
gobpPaths <- gobpPaths[sapply(gobpPaths, function(term) length(term) > 0)]

```

#### Annotate proteins to terms

This is to just save as a table.

```{r}

# Combine
finalTargetsReac <- merge(finalTargets,
                          reacPathData,
                          by.x="ENTREZID",
                          by.y="gene")
finalTargetsKegg <- merge(finalTargets,
                          keggPathData,
                          by.x="ENTREZID",
                          by.y="gene")
finalTargetsGobp <- merge(finalTargets,
                          gobpPathData,
                          by.x="ENTREZID",
                          by.y="gene")

# Check coverage
length(unique(finalTargetsReac$Protein))
length(unique(finalTargetsKegg$Protein))
length(unique(finalTargetsGobp$Protein))

# Save
write.table(finalTargets,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,".tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsReac,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedReactomePathways.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsKegg,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedKeggPathways.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)
write.table(finalTargetsGobp,
            here::here("results", "final", "crossSectional",
                       paste0("finalTargetProteins_withValidationAnnotation_imputedBy",imputMethod,"_allAnnotatedGoBP.tsv")),
            sep="\t",
            quote=FALSE,
            row.names=FALSE)

```

#### Perform tests

```{r}


getSetGenes <- function(pathData,
                        background) {

  # Loop over
  pathData["overlapGeneNames"] <- NA
  pathData["overlapUniprotIDs"] <- NA
  for (i in seq_len(nrow(pathData))) {

    # Get 
    uniProtIDs <- unlist(pathData[i,]$overlapGenes)
    geneIDs <- background[which(background$UNIPROT %in% uniProtIDs),]$SYMBOL
    
    # Add
    pathData[i,]["overlapGeneNames"] <- paste(geneIDs,
                                          collapse=", ")
    pathData[i,]["overlapUniprotIDs"] <- paste(uniProtIDs,
                                               collapse=", ")

  }

  # Return
  return(pathData[,-7])

}


```

##### Differentially abundant proteins

```{r}

# Loop over protein sets
protSets2Test <- c("AllSig", "DresdenOnly")
for (protSet in protSets2Test) {
  
  print(protSet)
  
  # Get prots as entrezid
  testProts <- allProtSets$limma$padj[[protSet]]
  
  # Enrichment test
  reacRes <- as.data.frame(fgsea::fora(reacPaths,
                         genes=testProts,
                         universe=intersect(backgroundList, 
                                            backgroundList),
                         minSize=10))
  print(reacRes[reacRes$padj < 0.05,])
  keggRes <- as.data.frame(fgsea::fora(keggPaths,
                         genes=testProts,
                         universe=intersect(backgroundList, 
                                            backgroundList),
                         minSize=10))
  print(keggRes[keggRes$padj < 0.05,])
  gobpRes <- as.data.frame(fgsea::fora(gobpPaths,
                         genes=testProts,
                         universe=intersect(backgroundList, 
                                            backgroundList),
                         minSize=10))
  print(gobpRes[gobpRes$padj < 0.05,])
  
  if (nrow(reacRes[reacRes$padj < 0.05,]) > 0) {
  
    # Save
    write.table(getSetGenes(reacRes[reacRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.tsv")),
                sep="\t",
                quote=FALSE,
                row.names=FALSE)
    write.xlsx(getSetGenes(reacRes[reacRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.xlsx")))
    
  }
  
  if (nrow(keggRes[keggRes$padj < 0.05,]) > 0) {
    
    write.table(getSetGenes(keggRes[keggRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.tsv")),
                sep="\t",
                quote=FALSE,
                row.names=FALSE)
    write.xlsx(getSetGenes(keggRes[keggRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.xlsx")))
    
  }
  
  if (nrow(gobpRes[gobpRes$padj < 0.05,]) > 0) {
    
    
    write.table(getSetGenes(gobpRes[gobpRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.tsv")),
                sep="\t",
                quote=FALSE,
                row.names=FALSE)
    write.xlsx(getSetGenes(gobpRes[gobpRes$padj < 0.05,],
                            background),
                here::here("results", "final", "crossSectional", 
                           paste0(protSet, "_OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.xlsx")))
    
  }
  
}

```

##### CNO vs all other conditions

The idea here is to see how proteins are dysregulated in CNO relative to other conditions (as the baseline) and enrich for up- or down-regulation.

```{r, fig.height=8. fig.width=8}

# Make design
rownames(trainMetaData) <- trainMetaData$Sample_ID
childCrossDesign_noInt <- model.matrix(~ 0 + Condition + Z_Age + Sex + PLATE, 
                                       data=trainMetaData)

# Loop over conditions
nonCNO <- c("Healthy",
            "Crohn",
            "InfOsteitis",
            "Oncology")
allToReturn <- vector(mode="list",
                      length=length(nonCNO))
for (i in seq_along(nonCNO)) {
  
  # Create contrast
  x <- nonCNO[[i]]
  childCross_contrastM_vsOtherCond <- makeContrasts(CNOvsOtherCond = paste0("ConditionCNO-Condition",x),
                                                    levels=childCrossDesign_noInt)
    
  # Re-run with date
  limmaRes_cnoOtherCond <- runLimma(trainQuantTable[,rownames(childCrossDesign_noInt)],
                                    designM=childCrossDesign_noInt,  
                                    levels2Test=paste0("ConditionCNO-Condition",x),
                                    contrasts2Fit=childCross_contrastM_vsOtherCond)
  limmaRes_cnoOtherCond_sig <- limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05,]
  print(nrow(limmaRes_cnoOtherCond_sig))
  
  # Enrichment test
  reacResUp <- as.data.frame(fgsea::fora(reacPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC > 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  keggResUp <- as.data.frame(fgsea::fora(keggPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC > 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  gobpResUp <- as.data.frame(fgsea::fora(gobpPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC > 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  
  reacResDn <- as.data.frame(fgsea::fora(reacPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC < 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  keggResDn <- as.data.frame(fgsea::fora(keggPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC < 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  gobpResDn <- as.data.frame(fgsea::fora(gobpPaths,
                        genes=limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05 & limmaRes_cnoOtherCond$logFC < 0,]$PG_ID,
                        universe=intersect(backgroundList, 
                                           backgroundList),
                        minSize=10))
  
  reacResUp["Condition"] <- nonCNO[[i]]
  keggResUp["Condition"] <- nonCNO[[i]]
  gobpResUp["Condition"] <- nonCNO[[i]]
  reacResDn["Condition"] <- nonCNO[[i]]
  keggResDn["Condition"] <- nonCNO[[i]]
  gobpResDn["Condition"] <- nonCNO[[i]]
  
  reacResUp["Direction"] <- "Up"
  keggResUp["Direction"] <- "Up"
  gobpResUp["Direction"] <- "Up"
  reacResDn["Direction"] <- "Down"
  keggResDn["Direction"] <- "Down"
  gobpResDn["Direction"] <- "Down"
  
  # Combine
  reacResUp <- getSetGenes(reacResUp,
                           background)
  keggResUp <- getSetGenes(keggResUp,
                           background)
  gobpResUp <- getSetGenes(gobpResUp,
                           background)
  
  reacResDn <- getSetGenes(reacResDn,
                           background)
  keggResDn <- getSetGenes(keggResDn,
                           background)
  gobpResDn <- getSetGenes(gobpResDn,
                           background)
  
  # Add database
  reacResUp["Database"] <- "Reactome"
  keggResUp["Database"] <- "KEGG"
  gobpResUp["Database"] <- "GO BP"
  reacResDn["Database"] <- "Reactome"
  keggResDn["Database"] <- "KEGG"
  gobpResDn["Database"] <- "GO BP"
  
  # Save
  write.table(reacResUp[reacResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  write.table(keggResUp[keggResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentKEGG_sigPadj0.05_logFCUp.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  write.table(gobpResUp[gobpResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentGOBP_sigPadj0.05_logFCUp.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  
  write.table(reacResDn[reacResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCDown.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  write.table(keggResDn[keggResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentKEGG_sigPadj0.05_logFCDown.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  write.table(gobpResDn[gobpResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentGOBP_sigPadj0.05_logFCDown.tsv")),
              sep="\t",
              quote=FALSE,
              row.names=FALSE)
  
  # Save excel
  write.xlsx(reacResUp[reacResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCUp.xlsx")))
  write.xlsx(keggResUp[keggResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentKEGG_sigPadj0.05_logFCUp.xlsx")))
  write.xlsx(gobpResUp[gobpResUp$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentGOBP_sigPadj0.05_logFCUp.xlsx")))
  
  write.xlsx(reacResDn[reacResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentReactome_sigPadj0.05_logFCDown.xlsx")))
  write.xlsx(keggResDn[keggResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentKEGG_sigPadj0.05_logFCDown.xlsx")))
  write.xlsx(gobpResDn[gobpResDn$padj < 0.05,],
              here::here("results", "final", "crossSectional", 
                         paste0("CNOvs",nonCNO[[i]],"OnAllProteins_foraEnrichmentGOBP_sigPadj0.05_logFCDown.xlsx")))
  
  # C
  allRes <- do.call("rbind",
                    list(reacResUp[reacResUp$padj < 0.05,],
                         # keggResUp[keggResUp$padj < 0.05,],
                         # gobpResUp[gobpResUp$padj < 0.05,],
                         reacResDn[reacResDn$padj < 0.05,] #,
                         # keggResDn[keggResDn$padj < 0.05,],
                         # gobpResDn[gobpResDn$padj < 0.05,]
                         ))
  allRes <- data.frame(allRes)
  rownames(allRes) <- NULL
  print(allRes[,c(7,8,1,3,4)])
  
  allToReturn[[i]] <- allRes
  
}

```


```{r, fig.height=7.5, fig.width=8}

# Plot
allToReturn |>
  bind_rows()|>
  filter(Direction == "Up") |>
  mutate(pathway = mgsub(pathway,
                         c("FACTOR IGF",
                           "FACTOR BINDING"),
                         c("FACTOR\nIGF",
                           "\nFACTOR BINDING"))) |>
  ggplot(aes(x=padj,
             y=reorder(pathway,
                       padj))) +
    geom_bar(stat="identity",
             fill="firebrick") +
    labs(x="Adjusted P-Value",
         y="") +
    facet_grid(rows=vars(Condition),
               cols=vars(Direction),
               scale="free_y",
               space="free_y") +
    theme_classic() +
    theme(strip.text.y = element_text(angle = 0))

```

```{r, fig.height=7.5, fig.width=6}

allToReturn |>
  bind_rows()|>
  filter(Direction == "Down") |>
  mutate(pathway = mgsub(pathway,
                         c("FACTOR IGF",
                           "FACTOR BINDING"),
                         c("FACTOR\nIGF",
                           "\nFACTOR BINDING"))) |>
  ggplot(aes(x=padj,
             y=reorder(pathway,
                       padj))) +
    geom_bar(stat="identity",
             fill="mediumblue") +
    labs(x="Adjusted P-Value",
         y="") +
    facet_grid(rows=vars(Condition),
               cols=vars(Direction),
               scale="free_y",
               space="free_y") +
    theme_classic() +
    theme(strip.text.y = element_text(angle = 0))


```

##### All candidates

```{r}

# Enrichment test
reacRes <- as.data.frame(fgsea::fora(reacPaths,
                                  genes=finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                  universe=backgroundList,
                                  minSize=10))
keggRes <- as.data.frame(fgsea::fora(keggPaths,
                                  genes=finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                  universe=backgroundList,
                                  minSize=2))
gobpRes <- as.data.frame(fgsea::fora(gobpPaths,
                                  genes=finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                  universe=backgroundList,
                                  minSize=10))
reacRes[reacRes$padj < 0.05,]
keggRes[keggRes$padj < 0.05,]
gobpRes[gobpRes$padj < 0.05,]

```

##### Up CNO vs other conditions

```{r}

for (i in seq_along(nonCNO)) {
  
  # Model matrix
  childCrossDesign_noInt <- model.matrix(~ 0 + Condition + Z_Age + Sex + PLATE, 
                                         data=trainMetaData[which(trainMetaData$Condition %in% c("CNO", nonCNO[[i]])),])

  # Create contrast
  x <- nonCNO[[i]]
  childCross_contrastM_vsOtherCond <- makeContrasts(CNOvsOtherCond = paste0("ConditionCNO-Condition",x),
                                                    levels=childCrossDesign_noInt)
    
  # Re-run with date
  limmaRes_cnoOtherCond <- runLimma(trainQuantTable[,rownames(childCrossDesign_noInt)],
                                    designM=childCrossDesign_noInt,  
                                    levels2Test=paste0("ConditionCNO-Condition",x),
                                    contrasts2Fit=childCross_contrastM_vsOtherCond)
  limmaRes_cnoOtherCond_sig <- limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$adj.P.Val < 0.05,]

  # Enrichment test
  reacResUp <- as.data.frame(fgsea::fora(reacPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC > 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=10))
  keggResUp <- as.data.frame(fgsea::fora(keggPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC > 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=2))
  gobpResUp <- as.data.frame(fgsea::fora(gobpPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC > 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=10))

  # Combine
  reacResUp <- getSetGenes(reacResUp,
                           background)
  keggResUp <- getSetGenes(keggResUp,
                           background)
  gobpResUp <- getSetGenes(gobpResUp,
                           background)
  
  reacResUp["Condition"] <- nonCNO[[i]]
  keggResUp["Condition"] <- nonCNO[[i]]
  gobpResUp["Condition"] <- nonCNO[[i]]
  
  reacResUp["Direction"] <- "Up"
  keggResUp["Direction"] <- "Up"
  gobpResUp["Direction"] <- "Up"
  
  # C
  allRes <- do.call("rbind",
                    list(reacResUp[reacResUp$padj < 0.05,],
                         keggResUp[keggResUp$padj < 0.05,],
                         gobpResUp[gobpResUp$padj < 0.05,]))
  allRes <- data.frame(allRes)
  rownames(allRes) <- NULL
  print(allRes[,c(7,8,1,3,4)])
  
  if (nrow(allRes[allRes$padj < 0.05,]) > 0) {
  
    # Save
    write.table(allRes[allRes$padj < 0.05,],
                here::here("results", "final", "crossSectional",
                           paste0("multiModalCandidates_foraEnrichment_upRegProts_inCNOvs",nonCNO[[i]],"_sigFDR0.05.tsv")),
                sep="\t",
                quote=FALSE,
                row.names=FALSE)
    write.xlsx(allRes[allRes$padj < 0.05,],
                here::here("results", "final", "crossSectional",
                           paste0("multiModalCandidates_foraEnrichment_upRegProts_inCNOvs",nonCNO[[i]],"_sigFDR0.05.xlsx")))
 
  }
  
}

```

##### Down CNO vs other conditions

```{r}

for (i in seq_along(rev(nonCNO))) {
  
  # Model matrix
  childCrossDesign_noInt <- model.matrix(~ 0 + Condition + Z_Age + Sex + PLATE, 
                                         data=trainMetaData[which(trainMetaData$Condition %in% c("CNO", nonCNO[[i]])),])
  
  # Create contrast
  x <- nonCNO[[i]]
  childCross_contrastM_vsOtherCond <- makeContrasts(CNOvsOtherCond = paste0("ConditionCNO-Condition",x),
                                                    levels=childCrossDesign_noInt)
    
  # Re-run with date
  limmaRes_cnoOtherCond <- runLimma(trainQuantTable[,rownames(childCrossDesign_noInt)],
                                    designM=childCrossDesign_noInt,  
                                    levels2Test=paste0("ConditionCNO-Condition",x),
                                    contrasts2Fit=childCross_contrastM_vsOtherCond)
  limmaRes_cnoOtherCond_sig <- limmaRes_cnoOtherCond # [limmaRes_cnoOtherCond$adj.P.Val < 0.05,]

  # Enrichment test
  reacResDn <- as.data.frame(fgsea::fora(reacPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC < 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=10))
  keggResDn <- as.data.frame(fgsea::fora(keggPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC < 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=2))
  gobpResDn <- as.data.frame(fgsea::fora(gobpPaths,
                                    genes=intersect(finalTargets[finalTargets$Model != "BiModal",]$Protein,
                                                    limmaRes_cnoOtherCond[limmaRes_cnoOtherCond$logFC < 0,]$PG_ID),
                                    universe=backgroundList,
                                    minSize=10))

  # Combine
  reacResDn <- getSetGenes(reacResDn,
                           background)
  keggResDn <- getSetGenes(keggResDn,
                           background)
  gobpResDn <- getSetGenes(gobpResDn,
                           background)
  
  reacResDn["Condition"] <- nonCNO[[i]]
  keggResDn["Condition"] <- nonCNO[[i]]
  gobpResDn["Condition"] <- nonCNO[[i]]
  
  reacResDn["Direction"] <- "Down"
  keggResDn["Direction"] <- "Down"
  gobpResDn["Direction"] <- "Down"
  
  # C
  allRes <- do.call("rbind",
                    list(reacResDn[reacResDn$padj < 0.05,],
                         keggResDn[keggResDn$padj < 0.05,],
                         gobpResDn[gobpResDn$padj < 0.05,]))
  allRes <- data.frame(allRes)
  rownames(allRes) <- NULL
  print(allRes[,c(9,10,1,3,4)])
  
  if (nrow(allRes[allRes$padj < 0.05,]) > 0) {

    # Save
    write.table(allRes[allRes$padj < 0.05,],
                here::here("results", "final", "crossSectional",
                           paste0("multiModalCandidates_foraEnrichment_downRegProts_inCNOvs",nonCNO[[i]],"_sigFDR0.05.tsv")),
                sep="\t",
                quote=FALSE,
                row.names=FALSE)
    write.xlsx(allRes[allRes$padj < 0.05,],
                here::here("results", "final", "crossSectional",
                           paste0("multiModalCandidates_foraEnrichment_downRegProts_inCNOvs",nonCNO[[i]],"_sigFDR0.05.xlsx")))
 
  }

}

```

#### Plot

```{r}

# Get significant proteins
sigProts <- strsplit(reacResDn[reacResDn$padj < 0.05,]$overlapUniprotIDs,
                     split=", ")[[1]]

# Get tall
tallData <- reshape2::melt(as.matrix(quantTable_allSamples[sigProts,trainMetaData$Sample_ID]))
tallData <- merge(trainMetaData,
                  tallData,
                  by.x="Sample_ID",
                  by.y="Var2")

tallData <- merge(tallData,
                  background,
                  by.x="Var1",
                  by.y="UNIPROT")

```

Plot significant term protein abundances

```{r}

# Loop over healthy/oncology
conds <- c("HealthyOncology")
for (cond in conds) {

  # Plot
  p <- ggplot(tallData[which(tallData$Condition %in% c("CNO", "Healthy", "Oncology")),],
         aes(x=interaction(SYMBOL,
                           Condition),
             y=value,
             fill=Condition)) +
    geom_boxplot(outlier.shape=NA,
                 color="black",
                 width=0.5) +
    geom_point(position=position_jitterdodge(),
               alpha=0.65,
               show.legend=FALSE,
               size=8) +
    labs(x="Candidate Proteins Overlapping With\nReactome Term 'Initial Triggering of Complement'",
         y="Log2 Abundance") +
    facet_wrap(~ SYMBOL,
               nrow=1,
               scales="free",
               space="free_x") +
    scale_fill_manual(values=condCols) +
    scale_color_manual(values=condCols) +
    theme_classic(base_size=56) +
    theme(plot.tag = element_text(face = "bold", 
                                            size = 100),
                    plot.title=element_text(size=100),
                    axis.text.y=element_text(size=72),
                    axis.text.x=element_blank(),
                    axis.ticks.x=element_blank(),
                    strip.text.x=element_text(size=84),
                    axis.title=element_text(size=108),
                    legend.text=element_text(size=84),
                    legend.title=element_text(size=92),
                    legend.position="top",
                    legend.direction="vertical",
                    legend.key.size = unit(9, "line")) &
              guides(fill=guide_legend(override.aes=list(size=28),
                                       nrow=1))
  
  # Save
  ggsave(here::here(figDir, 
                    paste0("figZZ_downRegCNOvsHealthyProts_sigReacPathwayProtss_abunancePlotted_justCNOvs",cond,".png")),
         p,
         width=15000,
         height=5000,
         units="px",
         limitsize=FALSE)
  ggsave(here::here(figDir,
                    paste0("figZZ_downRegCNOvsHealthyProts_sigReacPathwayProtss_abunancePlotted_justCNOvs",cond,".pdf")),
         p,
         width=15000,
         height=5000,
         device="pdf",
         units="px",
         limitsize=FALSE)
  
}

```

```{r}

# Plot
p <- ggplot(tallData,
       aes(x=interaction(SYMBOL,
                         Condition),
           y=value,
           fill=Condition)) +
  geom_boxplot(outlier.shape=NA,
               color="black",
               width=0.5) +
  geom_point(position=position_jitterdodge(),
             alpha=0.65,
             show.legend=FALSE,
             size=8) +
  labs(x="Candidate Proteins Overlapping With\nReactome Term 'Initial Triggering of Complement'",
       y="Log2 Abundance") +
  facet_wrap(~ SYMBOL,
             nrow=1,
             scales="free",
             space="free_x") +
  scale_fill_manual(values=condCols) +
  scale_color_manual(values=condCols) +
  theme_classic(base_size=56) +
  theme(plot.tag = element_text(face = "bold", 
                                          size = 100),
                  plot.title=element_text(size=100),
                  axis.text.y=element_text(size=72),
                  axis.text.x=element_blank(),
                  axis.ticks.x=element_blank(),
                  strip.text.x=element_text(size=84),
                  axis.title=element_text(size=108),
                  legend.text=element_text(size=84),
                  legend.title=element_text(size=92),
                  legend.position="top",
                  legend.direction="vertical",
                  legend.key.size = unit(9, "line")) &
            guides(fill=guide_legend(override.aes=list(size=28),
                                     nrow=1))

# Save
ggsave(here::here(figDir, 
                  "figZZ_downRegCNOvsHealthyProts_sigReacPathwayProtss_abunancePlotted_allConditions.png"),
       p,
       width=15000,
       height=6000,
       limitsize = FALSE,
       units="px")
ggsave(here::here(figDir,
                  "figZZ_downRegCNOvsHealthyProts_sigReacPathwayProtss_abunancePlotted_allConditions.pdf"),
       p,
       width=15000,
       height=6000,
       limitsize = FALSE,
       device="pdf",
       units="px")

```

#### Plot VNN1

```{r, fig.height=7, fig.width=20}

# Get tall
tallData <- reshape2::melt(as.matrix(quantTable_allSamples["O95497",
                                                           trainMetaData$Sample_ID]))
tallData <- merge(trainMetaData,
                  tallData,
                  by.x="Sample_ID",
                  by.y="Var2")
tallData <- merge(tallData,
                  background,
                  by.x="Var1",
                  by.y="UNIPROT")

# Plot
p <- ggplot(tallData,
       aes(x=Condition,
           y=value,
           fill=Condition)) +
  geom_boxplot(outlier.shape=NA,
               color="black",
               width=0.5) +
  geom_point(position=position_jitterdodge(),
             alpha=0.65,
             show.legend=FALSE,
             size=3) +
  facet_grid(cols=vars(Sex)) +
  labs(x="Condition",
       y="Log2 VNN1\nAbundance") +
  scale_fill_manual(values=condCols) +
  theme_classic(base_size=32)

p

```

### Networks

#### STRING data

##### Read in data

```{r}

# Read in String data
stringData <- read.table(here::here("data", "base", "annotation", "9606.protein.links.v12.0.txt"), 
                         sep=" ", 
                         header=TRUE)
colnames(stringData) <- c("Prot1", "Prot2", "Score")

# Remove species name
stringData$Prot1 <- gsub("9606[.]", "", stringData$Prot1)
stringData$Prot2 <- gsub("9606[.]", "", stringData$Prot2)

```

##### Convert Uniprot IDs in Network to Entrez ID

```{r}

# Convert IDs
termConversion <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db, 
                                        keys=union(stringData$Prot1,
                                                   stringData$Prot2),
                                        keytype="ENSEMBLPROT",
                                        columns=c("UNIPROT", "ENTREZID", "SYMBOL"))
termConversion <- termConversion[which(termConversion$UNIPROT %in% rownames(quantTable_allSamples) |
                                       termConversion$UNIPROT %in% rownames(quantTable_allSamples)),]

# Replace
stringData <- merge(stringData,
                     termConversion,
                     by.x="Prot1",
                     by.y="ENSEMBLPROT")
stringData <- merge(stringData,
                     termConversion,
                     by.x="Prot2",
                     by.y="ENSEMBLPROT")

# De-duplicate
stringData_dedup <- stringData[!duplicated(stringData[,c("ENTREZID.x", "ENTREZID.y")]),]

```

##### Filter

```{r}

# View scores
hist(stringData_dedup$Score)

# Filter interactions
stringData_dedup <- stringData_dedup[stringData_dedup$Score > 800,]

```

#### Make graph

```{r}

# Make igraph
stringGraph <- igraph::graph_from_edgelist(as.matrix(stringData_dedup[,c("Prot1", "Prot2")]),
                                           directed=FALSE)

```

### OmnipathR

#### Load interaction databases

```{r}

# Get others
omniPathPPI <- post_translational(organism="hsa")

```

#### Make network

```{r}

# Make igraph
g <- graph_from_edgelist(as.matrix(omniPathPPI[,c(1,2)]),
                         directed=TRUE)

```

#### Centrality

##### Undirected

```{r}

# Calculate hubs and betweenness
allNetHub <- hits_scores(g)
allNetBetw <- betweenness(g)

# Compare to out proteins
candidProtHub <- allNetHub$hub[intersect(names(allNetHub$hub), 
                                         finalTargets$Protein)]
mean(allNetHub$hub[intersect(allDetectedProts, names(allNetHub$hub))])
mean(candidProtHub)
sort(candidProtHub)

# Betweenness
candidProtBetw <- allNetBetw[intersect(names(allNetBetw), 
                                       finalTargets$Protein)]
mean(allNetBetw[intersect(allDetectedProts, names(allNetHub$hub))])
mean(candidProtBetw)
sort(candidProtBetw)

# Add to 
finalTargets["Hubness"] <- NA
finalTargets["Betweenness"] <- NA
finalTargets[intersect(names(candidProtHub), finalTargets$Protein),]["Hubness"] <- candidProtHub
finalTargets[intersect(names(candidProtBetw), finalTargets$Protein),]["Betweenness"] <- candidProtBetw

# Correlate
cor(finalTargets$Betweenness, 
    finalTargets$MeanVarImp,
    use="pairwise.complete.obs")
cor(finalTargets$Hubness, 
    finalTargets$MeanVarImp,
    use="pairwise.complete.obs")

```

#### Network enrichment test

NEAT method: https://link.springer.com/article/10.1186/s12859-016-1203-6#Sec2

##### Convert Uniprot IDs in Network to Entrez ID

```{r}

# Convert IDs
termConversion <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db, 
                                        keys=union(omniPathPPI$source,
                                                   omniPathPPI$target),
                                        keytype="UNIPROT",
                                        columns=c("UNIPROT", "ENTREZID", "SYMBOL"))

# Replace
omniPathPPI <- merge(omniPathPPI,
                     termConversion,
                     by.x="source",
                     by.y="UNIPROT")
omniPathPPI <- merge(omniPathPPI,
                     termConversion,
                     by.x="target",
                     by.y="UNIPROT")

# De-duplicate
omniPathPPI_dedup <- omniPathPPI[!duplicated(omniPathPPI[,c("ENTREZID.x", "ENTREZID.y")]),]

```

##### Get network defined as only query proteins and those within 1 edge of them

Keep only proteins in local neighbourhood of all detected proteins. Terms containing only these proteins are the only ones that could possibly be tested.

```{r}

# Restrict network to only those connected to desired terms
omniPathPPI_1stDegreeAdj <- getNodeDistances(g, 
                                             intersect(rownames(quantTable_allSamples), 
                                                       V(g)$name),
                                             V(g)$name)
omniPathPPI_1stDegreeAdj[omniPathPPI_1stDegreeAdj > 1] <- 0
omniPathPPI_1stDegreeAdj <- omniPathPPI_1stDegreeAdj[rowSums(omniPathPPI_1stDegreeAdj) > 0,
                                                     colSums(omniPathPPI_1stDegreeAdj) > 0]

# Make edgelist
omniPathPPI_1stDegreeEdges <- reshape2::melt(omniPathPPI_1stDegreeAdj)
omniPathPPI_1stDegreeEdges <- omniPathPPI_1stDegreeEdges[omniPathPPI_1stDegreeEdges$value > 0,c("Var1", "Var2")]

# Convert IDs
termConversion <- AnnotationDbi::select(org.Hs.eg.db::org.Hs.eg.db, 
                                        keys=union(omniPathPPI_1stDegreeEdges$Var1,
                                                   omniPathPPI_1stDegreeEdges$Var2),
                                        keytype="UNIPROT",
                                        columns=c("UNIPROT", "ENTREZID", "SYMBOL"))

# Replace
omniPathPPI_1stDegreeEdges <- merge(omniPathPPI_1stDegreeEdges,
                     termConversion,
                     by.x="Var1",
                     by.y="UNIPROT")
omniPathPPI_1stDegreeEdges <- merge(omniPathPPI_1stDegreeEdges,
                     termConversion,
                     by.x="Var2",
                     by.y="UNIPROT")

# De-duplicate
omniPathPPI_1stDegreeEdges_dedup <- omniPathPPI_1stDegreeEdges[!duplicated(omniPathPPI_1stDegreeEdges[,c("ENTREZID.x", "ENTREZID.y")]),]
omniPathPPI_1stDegreeEdges_dedup <- omniPathPPI_1stDegreeEdges_dedup[!is.na(omniPathPPI_1stDegreeEdges_dedup$ENTREZID.x) &
                                                                     !is.na(omniPathPPI_1stDegreeEdges_dedup$ENTREZID.y),]

```

##### Get proteins in the network

```{r}

# Get proteins detected
detectedProts <- rownames(quantTable_allSamples)
detectedProts <- intersect(detectedProts,
                           union(omniPathPPI$source, omniPathPPI$target))

# Subset network to only detected proteins
omniPathPPI_detectedProts <- omniPathPPI[which(omniPathPPI$source %in% detectedProts &
                                               omniPathPPI$target %in% detectedProts),]

# Get proteins
prots2Test <- union(omniPathPPI_detectedProts$ENTREZID.x,
                    omniPathPPI_detectedProts$ENTREZID.y)
prots2Test # THis is the background - needs to be proteins that were detected

# Define protein sets of interest
protSetsOfInterest <- list(WholePanel=finalTargets$ENTREZID, 
                           MultiModal=finalTargets[finalTargets$Model != "BiModal",]$ENTREZID,
                           BiModal=finalTargets[finalTargets$Model != "MultiModal",]$ENTREZID,
                           InValidation=finalTargets[finalTargets$InValidation == "Yes",]$ENTREZID,
                           MultiModalInValidation=finalTargets[finalTargets$Model != "MultiModal" &
                                                               finalTargets$InValidation == "Yes",]$ENTREZID)

```

##### Filter terms 

Need to make sure we keep only terms that have any overlap with the network

```{r}

keepOnlyTermsWithOverlap <- function(protIDs,
                                     termList,
                                     minOvlp=1,
                                     minTermSize=1) {

  # Loop over terms and return
  checkOverlap <- sapply(names(termList), function(termName) {
    
    # Test if any common proteins
    commonProts <- intersect(termList[[termName]], 
                             protIDs)
    
    # Filter
    check <- length(commonProts) >= minOvlp & # A minimum number of proteins must overlap
             length(termList[[termName]]) > minTermSize # Terms have to be of a minimum size
    
    # Return where not
    return(check)
    
  })
  
  # Return
  return(termList[checkOverlap])
  
}

```

##### Run Method

###### Reactome

```{r}

performNeatTest <- function(...) {
  
  # Run NEAT
  neatRes <- neat_defineBackground(...)
  neatRes <- data.frame(neatRes)
  
  # Rename
  colnames(neatRes) <- c("Panel Set", "Term", "Number of Edges", "Expected Number of Edges", "PVal", "AdjPVal", "Conclusion")
  
  # Return
  return(neatRes[order(neatRes$AdjPVal,
                       decreasing=FALSE),])
  
}

# Check significant
neatResReac <- performNeatTest(alist = protSetsOfInterest,
                               blist = keepOnlyTermsWithOverlap(protIDs=union(omniPathPPI_1stDegreeEdges_dedup$ENTREZID.x,
                                                                              omniPathPPI_1stDegreeEdges_dedup$ENTREZID.y),
                                                                termList=reacPaths,
                                                                minOvlp=3,
                                                                minTermSize=10), 
                               network = as.matrix(omniPathPPI_1stDegreeEdges_dedup[,c("ENTREZID.x", "ENTREZID.y")]),
                               nettype = 'undirected', 
                               nodes = prots2Test, 
                               alpha = 0.05, 
                               mtc.type = 'fdr')

# Get significant
neatResReac[neatResReac$AdjPVal < 0.05 & neatResReac$`Panel Set` == "MultiModal",]

```

###### KEGG

```{r}

# Check significant
neatResKegg <- performNeatTest(alist = protSetsOfInterest,
                               blist = keepOnlyTermsWithOverlap(protIDs=union(omniPathPPI_1stDegreeEdges_dedup$ENTREZID.x,
                                                                              omniPathPPI_1stDegreeEdges_dedup$ENTREZID.y),
                                                                termList=keggPaths,
                                                                minOvlp=1,
                                                                minTermSize=1), 
                               network = as.matrix(omniPathPPI_1stDegreeEdges_dedup[,c("ENTREZID.x", "ENTREZID.y")]),
                               nettype = 'undirected', 
                               nodes = prots2Test, 
                               alpha = 0.05, 
                               mtc.type = 'fdr')

# Get significant
neatResKegg[neatResKegg$AdjPVal < 0.05 & neatResKegg$`Panel Set` == "MultiModal",]

```

###### GO BP

```{r}

# Check significant
neatResGobp <- performNeatTest(alist = protSetsOfInterest,
                               blist = keepOnlyTermsWithOverlap(protIDs=union(omniPathPPI_1stDegreeEdges_dedup$ENTREZID.x,
                                                                              omniPathPPI_1stDegreeEdges_dedup$ENTREZID.y),
                                                                termList=gobpPaths,
                                                                minOvlp=1,
                                                                minTermSize=1), 
                               network = as.matrix(omniPathPPI_1stDegreeEdges_dedup[,c("ENTREZID.x", "ENTREZID.y")]),
                               nettype = 'undirected', 
                               nodes = prots2Test, 
                               alpha = 0.05, 
                               mtc.type = 'fdr')

# Get significant
neatResGobp[neatResGobp$AdjPVal < 0.05 & neatResGobp$`Panel Set` == "MultiModal",]

```






