---
title: "Figure Preparation"
author: "Euan McDonnell"
format: html
editor: source
---

# Code

## Pre-amble

### Run pre-amble script

```{r}

# Source metadata
here::i_am("17_figurePreparation.qmd")
source(here::here("utilities", "R", "codeModules", "boilerPlate", "loadParamsLibrariesFunctions.R"))

# Source discovery study metadata
source(here::here("utilities", "R", "codeModules", "data", "readMetadata.R"))

# Source colours
source(here::here("params", "colors.R"))

# Adjust colours
condStudyCols <- condCols
names(condStudyCols) <- paste0("Discovery.", names(condStudyCols))
lightCols <- lighten(condCols, 0.5)
names(lightCols) <- paste0("Validation.", names(lightCols))
condStudyCols <- c(condStudyCols, lightCols)

# Source Eva's ML biomarker functions
# source(here::here("utilities", "ExPeDiTe_publication", "ExpediteFunctions.R"))

```

### Read in plot list

```{r}

# List plots
# allFigs <- list.files(here::here("results", runName, "figs"),
#                       pattern="*.rds")

# Load
# plotList <- vector(mode="list",
#                    length=length(allFigs))
# for (i in seq_along(plotList)) {
#   
#   # Read in RDS
#   plotList[[i]] <- readRDS(here::here("results", runName, "figs", allFigs[[i]]))
#   
# }
# names(plotList) <- allFigs

# # Save as RDS
# plotList <- readRDS(here::here(figDir, 
#                     "19_validationDataPredictionModelling_figs.rds"))
 
```

## Main

### Primary figures

#### Figure 2

Candidates summaries.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`06_differentialAbundance_figs.rds` = readRDS(here::here("results", runName, "figs", "06_differentialAbundance_figs.rds")),
                `10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

plotList$`10_finalModel_figs.rds`$PPCABimProtsCondSitePC12$layers

removeGeom <- function(ggplot2_object, geom_type) {
  # Delete layers that match the requested type.
  layers <- lapply(ggplot2_object$layers, function(x) {
    if (class(x$geom)[1] == geom_type) {
      NULL
    } else {
      x
    }
  })
  # Delete the unwanted layers.
  layers <- layers[!sapply(layers, is.null)]
  ggplot2_object$layers <- layers
  ggplot2_object
}

# Combine
fig2Plt1 <- ((plotList$`10_finalModel_figs.rds`$PPCABimProtsCondSitePC12 + labs(tag="C") +
               ggtitle("Probabilistic PCA on\nBi-Modal Proteins") +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     plot.title=element_text(size=84),
                     axis.text.y=element_text(size=56),
                     axis.text.x=element_text(size=56),
                     axis.title=element_text(size=84)) +
               guides(fill=guide_legend(override.aes=list(size=20),
                                        nrow=3))) +
              (plotList$`10_finalModel_figs.rds`$PPCAGrpProtsCondSitePC12 + labs(tag="D") +
               ggtitle("Probabilistic PCA on\nMulti-Modal Proteins") +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     plot.title=element_text(size=84),
                     axis.text.y=element_text(size=56),
                     axis.text.x=element_text(size=56),
                     axis.title=element_text(size=84))) +
               guides(fill=guide_legend(override.aes=list(size=20),
                                        nrow=3)))
fig2Plt2 <- (plotList$`06_differentialAbundance_figs.rds`$allProtsPPCA_colByCondSitePC12 + 
               labs(tag="A",
                    fill="CondSite") +
               ggtitle("Probabilistic PCA on\nAll Proteins") +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     plot.title=element_text(size=84),
                     axis.text.y=element_text(size=56),
                     axis.text.x=element_text(size=56),
                     axis.title=element_text(size=84)) +
               scale_y_continuous(labels=function(x) sprintf("%.1f", x)) +
               guides(fill=guide_legend(override.aes=list(size=20),
                                        nrow=3)) +
            geom_point(color="white",
                       size=12) +
            geom_point(mapping=aes(fill=CondSite),
                       shape=21,
                       alpha=0.85,
                       size=12) +
               scale_fill_manual(breaks=factor(c("CNO_Dresden", 
                                                 "Crohn_Dresden",
                                                 "Oncology_Dresden", 
                                                 "Healthy_Liverpool",
                                                 "Healthy_Vanderbilt",
                                                 "InfOsteitis_Vanderbilt"),
                                 levels=c("CNO_Dresden", 
                                          "Crohn_Dresden",
                                          "Oncology_Dresden", 
                                          "Healthy_Liverpool",
                                          "Healthy_Vanderbilt",
                                          "InfOsteitis_Vanderbilt")),
                                 labels=c("CNO (Dresden)", 
                                          "Crohn (Dresden)",
                                          "Oncology (Dresden)", 
                                          "Healthy (Liverpool)",
                                          "Healthy (Vanderbilt)",
                                          "InfOsteitis (Vanderbilt)"),
                                 values=condSiteCols) +
                theme_classic(base_size=36) +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     legend.position="none",
                     legend.title=element_text(size=88),
                     axis.text.y=element_text(size=56),
                     axis.text.x=element_text(size=56),
                     legend.text=element_text(size=64),
                     axis.title=element_text(size=96),
                     plot.title=element_text(size=96))) +
  (ggplot() + theme_void() + labs(tag="B") + theme(plot.tag = element_text(face = "bold", size = 100))) +
             plot_layout(widths=c(1, 1))
# figS9Plt3 <- (ggplot() + theme_void() + labs(tag="D") + theme(plot.tag = element_text(face = "bold", size = 100)))

# Combine
fig2Plt <- ((fig2Plt2) /
           (fig2Plt1 + 
            plot_layout(guides="collect") &
            geom_point(color="white",
                       size=12) &
            geom_point(mapping=aes(fill=CondSite),
                       shape=21,
                       alpha=0.85,
                       size=12) &
               scale_fill_manual(breaks=factor(c("CNO_Dresden", 
                                                 "Crohn_Dresden",
                                                 "Oncology_Dresden", 
                                                 "Healthy_Liverpool",
                                                 "Healthy_Vanderbilt",
                                                 "InfOsteitis_Vanderbilt"),
                                 levels=c("CNO_Dresden", 
                                          "Crohn_Dresden",
                                          "Oncology_Dresden", 
                                          "Healthy_Liverpool",
                                          "Healthy_Vanderbilt",
                                          "InfOsteitis_Vanderbilt")),
                                 labels=c("CNO (Dresden)", 
                                          "Crohn (Dresden)",
                                          "Oncology (Dresden)", 
                                          "Healthy (Liverpool)",
                                          "Healthy (Vanderbilt)",
                                          "InfOsteitis (Vanderbilt)"),
                                 values=condSiteCols) &
               labs(fill="Condition\nand Site") &
               theme(legend.position="bottom", 
                     legend.direction="horizontal",
                     legend.title.position="left",
                     legend.title=element_text(size=96),
                     legend.text=element_text(size=76),
                     axis.title=element_text(size=88),
                     plot.title=element_text(size=96)))) &
               guides(fill=guide_legend(override.aes=list(size=32),
                                        nrow=3))

# Visualise
fig2Plt

# Save
ggsave(here::here(figDir,
                  "fig2_finalCandidatePPCA.png"),
       fig2Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "fig2_finalCandidatePPCA.pdf"),
       fig2Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# Remove to clear space
rm(fig2Plt)
rm(fig2Plt1)
rm(fig2Plt2)
rm(plotList)

```

#### Figure 3

Candidates summaries.

Non-imputed protein abundances so some NAs for some protein abundances in some samples.

```{r, fig.height=44, fig.width=32}

# Read in 
plotList = list(`10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

# Change size
# plotList$`10_finalModel_figs.rds`$selectProtsBim_abundanceBoxplots$layers[[2]]$aes_params$size <- 3.5
# plotList$`10_finalModel_figs.rds`$selectProtsGrp_abundanceBoxplots$layers[[2]]$aes_params$size <- 3.5
# plotList$`10_finalModel_figs.rds`$selectProtsBim_abundanceBoxplots$layers[[2]]$position$width <- 0
# plotList$`10_finalModel_figs.rds`$selectProtsGrp_abundanceBoxplots$layers[[2]]$position$width <- 0
# plotList$`10_finalModel_figs.rds`$selectProtsBim_abundanceBoxplots$layers[[2]]$position <- position_jitterdodge()
# plotList$`10_finalModel_figs.rds`$selectProtsGrp_abundanceBoxplots$layers[[2]]$position <- position_jitterdodge()

# CHange outlier shape
plotList$`10_finalModel_figs.rds`$selectProtsBim_abundanceBoxplots$layers[[1]]$geom_params$outlier.shape <- NA
plotList$`10_finalModel_figs.rds`$selectProtsGrp_abundanceBoxplots$layers[[1]]$geom_params$outlier.shape <- NA

# Combine
fig3Plt <- ((plotList$`10_finalModel_figs.rds`$selectProtsBim_abundanceBoxplots + 
               labs(tag="A") + 
               ggtitle("Bi-Modal Selected Proteins")) /
            (plotList$`10_finalModel_figs.rds`$selectProtsGrp_abundanceBoxplots + 
               labs(tag="B") + 
               ggtitle("Multi-Modal Selected Proteins")) +
                 plot_layout(heights=c(1, 2, 0.1),
                             guides="collect") &
                 geom_jitter(size=3.5,
                             shape=21,
                             show.legend=FALSE,
                             mapping=aes(fill=CondSite),
                             position=position_jitterdodge(),
                             alpha=0.65,
                             seed=1) &
                  guides(fill=guide_legend(nrow=3))) +
              plot_spacer() &
                 labs(fill="Condition\nand Site",
                      y=bquote(Log[2]( Abundance))) &
               scale_fill_manual(breaks=factor(c("CNO_Dresden", 
                                                 "Crohn_Dresden",
                                                 "Healthy_Liverpool",
                                                 "Healthy_Vanderbilt",
                                                 "InfOsteitis_Vanderbilt",
                                                 "Oncology_Dresden"),
                                 levels=c("CNO_Dresden", 
                                          "Crohn_Dresden",
                                          "Healthy_Liverpool",
                                          "Healthy_Vanderbilt",
                                          "InfOsteitis_Vanderbilt",
                                          "Oncology_Dresden")),
                                 labels=c("CNO (Dresden)", 
                                          "Crohn (Dresden)",
                                          "Healthy (Liverpool)",
                                          "Healthy (Vanderbilt)",
                                          "InfOsteitis (Vanderbilt)",
                                          "Oncology (Dresden)"),
                                 values=condSiteCols) &
                 theme(plot.tag = element_text(face = "bold",
                                               size = 100),
                       strip.background =element_rect(fill="white"),
                       axis.title=element_text(size=88),
                       plot.title=element_text(size=96),
                       axis.text.y=element_text(size=44),
                       # axis.title.x=element_text(size=88),
                       # axis.title.y=element_text(size=88),
                       legend.position="bottom",
                       legend.direction="horizontal",
                       legend.title.position="left",
                       legend.text=element_text(size=76),
                       legend.title=element_text(size=92),
                       strip.text=element_text(size=52),
                       legend.key.size = unit(9, "line")) &
                 guides(fill=guide_legend(byrow=TRUE,
                                          ncol=2))

# View
fig3Plt

# Save
ggsave(here::here(figDir,
                  "fig3_finalCandidateSummaries.png"),
       fig3Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "fig3_finalCandidateSummaries.pdf"),
       fig3Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# # Remove to clear space
rm(fig3Plt)

```

#### Figure 4

Model metric performances on the discovery cohort.

Oncology missing in 1 split, so 4 NAs for Sensitivity/F1/Bal Acc/AUROC

```{r, fig.height=40, fig.width=32}

# Change size and alpha of points
plotList$`10_finalModel_figs.rds`$biModalProtsMetrics$layers[[2]]$aes_params$size <- 10
plotList$`10_finalModel_figs.rds`$biModalProtsMetrics$layers[[2]]$aes_params$alpha <- 0.675
plotList$`10_finalModel_figs.rds`$biModalMultiGrpProtsMetrics$layers[[2]]$aes_params$size <- 10
plotList$`10_finalModel_figs.rds`$biModalMultiGrpProtsMetrics$layers[[2]]$aes_params$alpha <- 0.675
plotList$`10_finalModel_figs.rds`$multiModelProtsMetrics$layers[[2]]$aes_params$size <- 8
plotList$`10_finalModel_figs.rds`$multiModelProtsMetrics$layers[[2]]$aes_params$alpha <- 0.675

# --- create a common theme to reduce grob-width differences ---
common_theme <- theme(
  strip.text = element_text(size = 52),
  axis.text.x  = element_text(size = 64),
  plot.margin = margin(5, 5, 5, 5)
)

# --- build each sub-plot separately for clarity ---
pA <- plotList$`10_finalModel_figs.rds`$biModalProtsMetrics +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "A", title = "Bi-Modal") +
  common_theme +
  theme(axis.text.x = element_text(size=72),
        axis.text.y = element_text(size = 64),
        axis.title.y = element_text(size = 100,
                                        angle=90),
        legend.position="none") +
  # hide fill legend here so only one plot supplies it to be collected
  guides(color = "none")

pB <- plotList$`10_finalModel_figs.rds`$biModalMultiGrpProtsMetrics +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "B", 
       title = "Bi-Modal\n(Whole Panel)") +
  common_theme +
  # remove redundant left axis elements to avoid double-reserving left space
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 72),
        legend.position="none")

# This plot will supply the legend (so don't hide its legend)
pC <- plotList$`10_finalModel_figs.rds`$multiModelProtsMetrics +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "C") +
  common_theme +
  # remove left y-axis items so they don't take extra space
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(-0.275, -0.1),
        legend.direction="horizontal",
        axis.text.x = element_text(size = 72, 
                                   angle = 90, 
                                   hjust = 1, 
                                   vjust = 0.5))

# bottom row plot (held-out data)
pD <- plotList$`10_finalModel_figs.rds`$multiModelProtsMetrics_heldOutData +
  geom_text(mapping = aes(y = value - 0.05, label = round(value, 2)),
            color = "white", hjust = 1, angle = 90, size = 24) +
  facet_wrap(~ variable, nrow = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "D") +
  common_theme +
  theme(legend.direction = "horizontal",
        legend.position = "bottom",
        legend.key.size = unit(6, "line"),
        legend.title = element_text(size = 88),
        legend.text = element_text(size = 80),
        axis.text.x = element_blank())

# --- Combine with patchwork ---
# Important notes applied:
# - only pC provides the legend (show.legend = TRUE); guides = "collect" will therefore
#   collect a single, consistent legend and place it where we request
# - widths are set to favour the right-hand multi-panel plot, but because we've
#   removed duplicated left-axis grob widths and standardized theme settings,
#   columns will be much more proportional
fig4Plt <- (
  (pA + pB + pC) +
    plot_layout(widths = c(1, 1, 1.5)) &
    labs(color="Condition") &
    theme(
      legend.title = element_text(size = 88),
      legend.text = element_text(size = 80),
      axis.text.x = element_blank()) &
    guides(color=guide_legend(override.aes=list(size=28),
                              nrow=2,
                              byrow=TRUE))
) /
  plot_spacer() /
  (pD + 
  theme(axis.title.y = element_text(size = 100,
                                    angle=90),
        axis.text.y  = element_text(size = 64))) +
  plot_layout(heights = c(1, 0.1, 0.75), axes = "collect_y", axis_titles = "collect_y") &
  labs(fill = "Condition") &
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.05, 1.05)) &
  theme(
    plot.tag = element_text(face = "bold", size = 100),
    plot.title = element_text(size = 88),
    strip.text = element_text(size = 84),
    axis.title.x = element_blank(),
  ) &
  guides(fill=guide_legend(nrow=2,
                           byrow = TRUE))

fig4Plt

# Save
ggsave(here::here(figDir, 
                  "fig4_discoveryModelPerformances.png"),
       fig4Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "fig4_discoveryModelPerformances.pdf"),
       fig4Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# Remove to clear space
rm(pA)
rm(pB)
rm(pC)
rm(pD)
rm(fig4Plt)
# rm(plotList)

```

#### Figure 5

Plot protein abundances by validation and discovery.

Could measure variance between validation and discovery, per group.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`15_validationDataPredictionModelling_figs_bPCA.rds` = readRDS(here::here("results", runName, "figs", "15_validationDataPredictionModelling_figs_bPCA.rds")))

# Adjust colours
condStudyCols <- condCols
names(condStudyCols) <- paste0("Discovery.", names(condStudyCols))
lightCols <- lighten(condCols, 0.5)
names(lightCols) <- paste0("Validation.", names(lightCols))
condStudyCols <- c(condStudyCols, lightCols)

# Order
condStudyCols <- condStudyCols[c("Discovery.CNO", 
                                 "Validation.CNO",
                                 "Discovery.Healthy",
                                 "Discovery.Crohn", 
                                 "Discovery.Oncology", 
                                 "Discovery.InfOsteitis",
                                 "Validation.Healthy",
                                 "Validation.Crohn",
                                 "Validation.Oncology",
                                 "Validation.InfOsteitis")]

# Combine
fig5Plt <- ( (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$biModalProts_abundanceBoxplots_validVsDisco_bPCA + labs(tag="A") + ggtitle("Bi-Modal Selected Proteins") +
               labs(y="Scaled Abundance") +
               theme(strip.text=element_text(size=56),
                     legend.position="none")) /
             (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$multiGroupProts_abundanceBoxplots_validVsDisco_bPCA + labs(tag="B") + ggtitle("Multi-Modal Selected Proteins") +
               labs(y="Scaled Abundance") +
               theme(strip.text=element_text(size=56),
                     legend.position="bottom") +
               guides(fill=guide_legend(override.aes=list(size=32),
                                        nrow=5)))) +
  plot_layout(heights=c(1, 1.5),
              nrow=2) &
  labs(fill="Condition\nby Cohort") &
  scale_y_continuous(labels = scales::label_number(accuracy = 1)) & 
  geom_point(mapping=aes(fill=interaction(Data_Group, 
                                          Condition)),
             position=position_jitterdodge(),
             show.legend=FALSE,
             alpha=0.35,
             size=4) &
  scale_fill_manual(breaks=factor(c("Discovery.CNO", 
                                    "Discovery.Crohn", 
                                    "Discovery.Healthy", 
                                    "Discovery.InfOsteitis",
                                    "Discovery.Oncology",
                                    "Validation.CNO",
                                    "Validation.Crohn",
                                    "Validation.Healthy",
                                    "Validation.InfOsteitis",
                                    "Validation.Oncology"),
                           levels=c("Discovery.CNO", 
                                    "Discovery.Crohn", 
                                    "Discovery.Healthy",
                                    "Discovery.InfOsteitis",
                                    "Discovery.Oncology", 
                                    "Validation.CNO",
                                    "Validation.Crohn",
                                    "Validation.Healthy",
                                    "Validation.InfOsteitis",
                                    "Validation.Oncology")),
                           labels=c("CNO (Discovery)",
                                    "Crohn (Discovery)",
                                    "Healthy (Discovery)",
                                    "InfOsteitis (Discovery)",
                                    "Oncology (Discovery)",
                                    "CNO (Validation)",
                                    "Crohn (Validation)",
                                    "Healthy (Validation)",
                                    "InfOsteitis (Validation)",
                                    "Oncology (Validation)"),
                      values=condStudyCols) &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        strip.text=element_text(size=64),
        legend.text=element_text(size=72),
        legend.title=element_text(size=84),
        legend.title.position="left",
        plot.title=element_text(size=96),
        axis.text.y=element_text(size=44),
        axis.title.y=element_text(size=88),
        axis.title.x=element_text(size=88))

fig5Plt

# Save
ggsave(here::here(figDir, 
                  "fig5_candidateAbundance_discoveryAndValidation.png"),
       fig5Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "fig5_candidateAbundance_discoveryAndValidation.pdf"),
       fig5Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# Remove to clear space
rm(fig5Plt)
rm(plotList)

```

#### Figure 6

Comparison between discovery and validation differential abundance analysis.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`14_validationDataDiffAbundance_figs.rds` = readRDS(here::here("results", runName, "figs", "14_validationDataDiffAbundance_figs.rds")))

# Adjust asterisks
plotList$`14_validationDataDiffAbundance_figs.rds`$diffAbundanceComparison_orthoganolDistance$layers[[2]]$aes_params$size <- 30
plotList$`14_validationDataDiffAbundance_figs.rds`$diffAbundanceComparison_orthoganolDistance$layers[[2]]$aes_params$hjust <- -0.45

fig6Plt2 <- (plotList$`14_validationDataDiffAbundance_figs.rds`$diffAbundanceComparison_orthoganolDistance +
              geom_vline(xintercept=0,
                         color="darkgrey") +
              labs(y="Candidate Protein",
                   x="Scaled Distance",
                   pattern="Consistent LogFC") +
                   scale_x_continuous(breaks = c(0, 2, 4),
                                      limits  = c(0, 4.75),
                                      labels  = c(0, 2, 4)) +
              theme(axis.text.x=element_text(size=56),
                    legend.key=element_rect(colour="white",
                                            size=5)))

# Combine
fig6Plt <- fig6Plt2 +
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        legend.position = "bottom",
        legend.direction="horizontal",
        strip.background.x=element_blank(),
        strip.text.y=element_text(size=92),
        strip.text.x=element_text(size=80),
        axis.title.y=element_text(size=108),
        axis.title.x=element_text(size=108),
        axis.text.x=element_text(size=88),
        axis.text.y=element_text(size=68),
        legend.text=element_text(size=88),
        legend.title=element_text(size=100),
        legend.title.position="top",
        panel.spacing.y = unit(2, "lines"),
        legend.key.size = unit(6, "line")) +
               scale_fill_manual(values=condCols,
                                 breaks=c(factor(c("CNO", 
                                                "Crohn", 
                                                "Healthy", 
                                                "InfOsteitis",
                                                "Oncology"),
                                               levels=c("CNO", 
                                                        "Crohn", 
                                                        "Healthy", 
                                                        "InfOsteitis",
                                                        "Oncology")))) +
        guides(fill=guide_legend(override.aes=list(size=30,
                                                   pattern="none"),
                                 nrow=2,
                                 byrow=TRUE),
               pattern=guide_legend(override.aes=list(size=30,
                                                      shape=c(22, 21)),
                                    nrow=2),
               alpha="none")
fig6Plt

# Save
ggsave(here::here(figDir, 
                  "fig6_differentialAbundance_discoveryVsValidation.png"),
       fig6Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "fig6_differentialAbundance_discoveryVsValidation.pdf"),
       fig6Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

# Remove to clear space
rm(fig6Plt1)
rm(fig6Plt2)
rm(fig6Plt)
rm(plotList)
gc()

```

#### Figure 7

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`15_validationDataPredictionModelling_figs_bPCA.rds` = readRDS(here::here("results", runName, "figs", "15_validationDataPredictionModelling_figs_bPCA.rds")))

# Adjust size
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_validData_bPCA$layers[[2]]$aes_params <- c(hjust=-100,
                                                                                                                  plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_validData_bPCA$layers[[2]]$aes_params)
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_validData_bPCA$layers[[2]]$aes_params$colour <- "white"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_validData_bPCA$layers[[2]]$aes_params$size <- NA
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_multiGroup_bPCA$layers[[2]]$aes_params$size <- 12
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_multiGroup_bPCA$layers <- plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_multiGroup_bPCA$layers[-3]

# Plot
fig7Plt <- ((plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_validData_bPCA + labs(tag="A") + 
               scale_y_continuous(breaks=c(0, 0.25, 0.5, 0.75, 1),
                                  limits=c(-0.1, 1)) +
               geom_hline(yintercept=0,
                          color="black") +
               geom_text(mapping=aes(label=ifelse(round(value, 2) <= 0.3, NA, round(value, 2)),
                                     x=variable,
                                     y=value,
                                     ),
                         position=position_dodge(width=0.9),
                         hjust=1.075,
                         color="white",
                         angle=90,
                         size=20) +
               geom_text(mapping=aes(label=ifelse(round(value, 2) > 0.3, NA, round(value, 2)),
                                     x=variable,
                                     y=value,
                                     ),
                         position=position_dodge(width=0.9),
                         hjust=-0.1,
                         color="black",
                         angle=90,
                         size=20) +
               scale_color_manual(values=c("X"="white",
                                           "Y"="black")) +
               scale_fill_manual(values=condCols,
                                 breaks=c(factor(c("CNO", 
                                                "Crohn", 
                                                "Healthy", 
                                                "InfOsteitis",
                                                "Oncology"),
                                               levels=c("CNO", 
                                                        "Crohn", 
                                                        "Healthy", 
                                                        "InfOsteitis",
                                                        "Oncology")))) +
               facet_wrap(~ variable,
                          ncol=3,
                          scale="free_x") +
               geom_hline(yintercept=0,
                          color="black",
                          linewidth=2) +
               scale_y_continuous(limits=c(-0.1, 1.05)) +
               theme(strip.text=element_text(size=48),
                     legend.title.position="left",
                     legend.key.size = unit(7.5, "line")) +
               guides(fill=guide_legend(override.aes=list(size=32),
                                                          nrow=2,
                                                          byrow=TRUE))) /
            (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_multiGroup_bPCA + labs(tag="B",
                                                        x="Mis-Predicted") +
               theme(strip.text=element_text(size=64),
                     axis.text.x=element_text(size=72),
                     legend.title.position="top",
                     legend.box="horizontal", 
                     legend.margin=margin()) +
               guides(fill=guide_legend(override.aes=list(size=36, 
                                                          shape=21),
                                                          nrow=2,
                                        byrow=TRUE),
                      shape=guide_legend(override.aes=list(size=36),
                                                           nrow=2),
                      color="none")) +
             plot_layout(heights=c(1.75, 1))) &
             labs(fill="Condition") &
             theme(plot.tag = element_text(face = "bold", 
                                           size = 100),
                   strip.text=element_text(size=84),
                   plot.title=element_text(size=64),
                   axis.text.y=element_text(size=64),
                   axis.title.y=element_text(size=100),
                   axis.title.x=element_text(size=100),
                   legend.position="bottom",
                   legend.title=element_text(size=88),
                   legend.text=element_text(size=80))

fig7Plt

# Save
ggsave(here::here(figDir, 
                  "fig7_multiModalModelPerformance_onValidation.png"),
       fig7Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "fig7_multiModalModelPerformance_onValidation.pdf"),
       fig7Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(fig7Plt)
# rm(plotList)

```

#### Figure 8

```{r}

library(ggplot2)
library(ggforce)
library(sf)

# --- your sets ---
set1 <- c("GSN", "LRG1", "CRISP3", "PGK1", "CAT",
          "ICAM1", "CD14",
          "CRP", "FCN2", "FCN3",
          "F2", "C5", "C4A", "C3", "ALDOA")

set2 <- c("MST1", "TNC", "SERPINA10",
          "ICAM1", "CD14",
          "ITIH3",
          "F2", "C5", "C4A", "C3", "ALDOA")

set3 <- c("SERPINA5",
          "CRP", "FCN2", "FCN3",
          "ITIH3",
          "F2", "C5", "C4A", "C3", "ALDOA")

n1_only <- setdiff(set1, union(set2, set3))
n2_only <- setdiff(set2, union(set1, set3))
n3_only <- setdiff(set3, union(set1, set2))
n12     <- setdiff(intersect(set1, set2), set3)
n23     <- setdiff(intersect(set2, set3), set1)
n13     <- setdiff(intersect(set1, set3), set2)
n123    <- intersect(intersect(set1, set2), set3)

fmt <- function(x) paste(sort(x), collapse = "\n")

# Increase the spread slightly to shrink the center (triple) overlap
# faster than the pairwise overlaps
circle_params <- data.frame(
  x0 = c(-1.15, 1.15, 0),     # was -0.55, 0.55 — push circles apart a bit
  y0 = c(0.9, 0.9, -0.5),     # was 0.35, 0.35, -0.7
  r  = 2.2,
  category = c("Immune", "Signalling", "Coagulation")
)

make_circle_sf <- function(x0, y0, r, n = 200) {
  st_buffer(st_point(c(x0, y0)), dist = r, nQuadSegs = n / 4)
}

circ1 <- make_circle_sf(circle_params$x0[1], circle_params$y0[1], circle_params$r[1])
circ2 <- make_circle_sf(circle_params$x0[2], circle_params$y0[2], circle_params$r[2])
circ3 <- make_circle_sf(circle_params$x0[3], circle_params$y0[3], circle_params$r[3])

region_geom <- list(
  n1_only = st_difference(circ1, st_union(circ2, circ3)),
  n2_only = st_difference(circ2, st_union(circ1, circ3)),
  n3_only = st_difference(circ3, st_union(circ1, circ2)),
  n12     = st_difference(st_intersection(circ1, circ2), circ3),
  n23     = st_difference(st_intersection(circ2, circ3), circ1),
  n13     = st_difference(st_intersection(circ1, circ3), circ2),
  n123    = st_intersection(st_intersection(circ1, circ2), circ3)
)

# --- safe centroid extraction: returns c(NA, NA) if geometry is empty ---
get_centroid_xy <- function(geom) {
  if (is.null(geom) || length(geom) == 0 || st_is_empty(geom)) {
    return(c(x = NA_real_, y = NA_real_))
  }
  ctr <- st_centroid(geom)
  xy  <- st_coordinates(ctr)
  c(x = xy[1, 1], y = xy[1, 2])
}

coords_list <- lapply(region_geom, get_centroid_xy)
coords <- do.call(rbind, coords_list)

# check for any empty regions before proceeding
if (any(is.na(coords))) {
  empty_regions <- names(region_geom)[apply(coords, 1, function(r) any(is.na(r)))]
  warning("These regions came back empty (circles may not overlap enough / too much): ",
          paste(empty_regions, collapse = ", "))
}

labels <- data.frame(
  x = coords[, "x.X"],
  y = coords[, "y.Y"],
  label = c(fmt(n1_only), fmt(n2_only), fmt(n3_only),
            fmt(n12), fmt(n23), fmt(n13), fmt(n123))
)

cat_labels <- data.frame(
  x = c(-1.6, 1.6, 0),
  y = c(2.5, 2.5, -3.1),
  label = c("Immune", "Signalling", "Coagulation")
)

p1 <- ggplot() +
  geom_circle(data = circle_params,
              aes(x0 = x0, 
                  y0 = y0, 
                  r = r, 
                  fill = category),
              alpha = 0.4, 
              color = NA) +
  geom_circle(data = circle_params,
              aes(x0 = x0, 
                  y0 = y0, 
                  r = r),
              fill = NA, 
              color = "white", 
              linewidth = 4) +
  geom_text(data = labels, 
            aes(x = x, 
                y = y, 
                label = label),
            size = 28, 
            lineheight = 0.85, 
            na.rm = TRUE) +
  geom_text(data = cat_labels, 
            aes(x = ifelse(label == "Coagulation",
                           x,
                           x * 1.65), 
                y = ifelse(label == "Coagulation",
                           y,
                           y * 1.3), 
                label = label),
            size = 40, 
            fontface = "bold") +
  scale_fill_manual(values = c("red", "blue", "green")) +
  coord_fixed(
    clip = "off"
  ) +
  theme_void() +
  theme(legend.position = "none") +
  theme(
    plot.title = element_text(
      size = 128,
      face = "bold",
      hjust = 0,
      margin = margin(b = 20, l = -200)   # negative left margin pulls title further left
    ),
    plot.background  = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )

# --- save ---
ggsave(
  filename = here::here(resDir, "figs",
                         "fig8_selectedProteins_functionalCategories_vennDiagram.png"),
  plot = p1 + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 120)),
  height = 10000, width = 10000, units = "px", dpi = 300, limitsize=FALSE,
  bg = "white"
)

# --- save ---
ggsave(
  filename = here::here(resDir, "figs",
                         "fig8_selectedProteins_functionalCategories_vennDiagram.pdf"),
  plot = p1 + theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 120)),
  height = 10000, width = 10000, units = "px", dpi = 300, limitsize=FALSE, device="pdf",
  bg = "white"
)

```

### Supplementary

#### Figure S1

```{r, fig.height=40, fig.width=32}

# Read in
plotList = list(`06_differentialAbundance_figs.rds` = readRDS(here::here("results", runName, "figs", "06_differentialAbundance_figs.rds")),
                `11_validationDataProcessQC_figs.rds` = readRDS(here::here("results", runName, "figs", "11_validationDataProcessQC_figs.rds")))

# Combine
figS1Plt1 <- (plotList$`06_differentialAbundance_figs.rds`$conditionGroupNumbers_trainData + labs(tag="A", title="Discovery") +
                geom_text(aes(label=ifelse(value > 0,
                                           value,
                                           NA),
                              x=Site,
                              y=value),
                          color="white",
                          position = position_stack(vjust = .5),
                          size=20,
                          show.legend=FALSE) +
                scale_color_manual(values=c("black",
                                            "white"))) +
             (plotList$`11_validationDataProcessQC_figs.rds`$sampleConditionGroupNumbers_validData + labs(tag="B", title="Validation") +
                geom_text(aes(label=ifelse(value > 0,
                                           value,
                                           NA),
                              vjust=ifelse(value == 1,
                                           -0.65,
                                           0),
                              x=Site,
                              y=value,
                              color=ifelse(value == 1, "black", "white")),
                          position = position_stack(vjust = .5),
                          size=20,
                          show.legend=FALSE) +
                scale_color_manual(values=c("black",
                                            "white"))) +
             (plotList$`06_differentialAbundance_figs.rds`$conditionGroupNumbers_trainData_justSamplesWithDate + labs(tag="C", title="Discovery\nwith Date") +
                geom_text(aes(label=ifelse(value > 0,
                                           value,
                                           NA),
                              vjust=ifelse(value == 4,
                                           -1,
                                           0),
                              x=Site,
                              y=value,
                              color=ifelse(value == 4, "black", "white")),
                          position = position_stack(vjust = .5),
                          size=20,
                          show.legend=FALSE) +
                scale_color_manual(values=c("black",
                                            "white"))) +
              plot_layout(guides="collect",
                          axes="collect") &
              theme(plot.tag = element_text(face = "bold",
                                           size = 100),
                   plot.title=element_text(size=80),
                   axis.text.x=element_text(size=52,
                                            angle=90,
                                            hjust=1,
                                            vjust=0.5),
                   axis.text.y=element_text(size=64),
                   axis.title.y=element_text(size=72),
                   axis.title.x=element_text(size=72),
                   strip.text=element_text(size=64),
                   legend.text=element_text(size=64),
                   legend.title=element_text(size=72),
                   legend.position="bottom",
                   legend.key.size = unit(6.5, "line"),
                   legend.title.position="left",
                   legend.direction="horizontal") &
  guides(fill=guide_legend(override.aes=list(size=20,
                                             nrow=2)))
figS1Plt2 <- plotList$`06_differentialAbundance_figs.rds`$sampleDateDistribution + labs(tag="D") + 
               theme(plot.tag = element_text(face = "bold",
                                             size = 100),
                     plot.title=element_text(size=80),
                     axis.text.x=element_text(size=56),
                     axis.text.y=element_text(size=64),
                     axis.title.y=element_text(size=72),
                     axis.title.x=element_text(size=72),
                     strip.text=element_text(size=64),
                     legend.text=element_text(size=64),
                     legend.title=element_text(size=72),
                     legend.position="bottom",
                     legend.title.position="left",
                     legend.direction="horizontal") &
  guides(fill=guide_legend(override.aes=list(size=28)))

# Plot
figS1Plt1[[1]] <- figS1Plt1[[1]] + theme(legend.position="none")
figS1Plt1[[3]] <- figS1Plt1[[3]] + theme(legend.position="none")
figS1Plt <- ( figS1Plt1 + theme(axis.text.x=element_text(size=52,
                                                         angle=90,
                                                         hjust=1,
                                                         vjust=0.5)) +
                          guides(fill=guide_legend(override.aes=list(size=20,
                                                                     nrow=2))) ) / 
              figS1Plt2
figS1Plt

# Save
ggsave(here::here(figDir, 
                  "figS1_conditionSiteGroupDistributions.png"),
       figS1Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS1_conditionSiteGroupDistributions.pdf"),
       figS1Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(figS1Plt1)
rm(figS1Plt2)
rm(figS1Plt3)
rm(figS1Plt)
rm(plotList)

```

#### Figure S2

```{r, fig.height=20, fig.width=32}

# Read in
plotList = list(`02_dataProcessQC_figs.rds` = readRDS(here::here("results", runName, "figs", "02_dataProcessQC_figs.rds")),
                `03_dataNormalisation_figs.rds` = readRDS(here::here("results", runName, "figs", "03_dataNormalisation_figs.rds")))

# Change text size
plotList$`02_dataProcessQC_figs.rds`$NProtsAllProts$layers[[3]]$aes_params$size <- 20
plotList$`02_dataProcessQC_figs.rds`$NProtsSerumQCProts$layers[[3]]$aes_params$size <- 20
plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_raw$layers[[1]]$aes_params$linewidth <- 0.5
plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_raw$layers[[1]]$aes_params$size <- 2
plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm$layers[[1]]$aes_params$linewidth <- 0.5
plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm$layers[[1]]$aes_params$size <- 2

# Combine
figS2Plt1 <- ((plotList$`02_dataProcessQC_figs.rds`$NProtsAllProts + labs(tag="A") + 
                ggtitle("All Proteins") + ylab("Number of\nRetained Proteins") +
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%"))) +
             (plotList$`02_dataProcessQC_figs.rds`$NProtsSerumQCProts + labs(tag="B")) + 
               ggtitle("Serum QC Proteins") + ylab("Number of\nRetained Proteins") +
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%"))) +
             plot_layout(axes="collect",
                         axis_titles="collect") &
             labs(x="Allowed Percentage\nof Missing Values") &
             theme(plot.tag = element_text(face = "bold",
                                           size = 100),
                   plot.title=element_text(size=84),
                   axis.text.x=element_text(size=72,
                                            angle=90,
                                            vjust=0.5,
                                            hjust=1),
                   axis.text.y=element_text(size=64),
                   axis.title.y=element_text(size=84),
                   axis.title.x=element_text(size=72),
                   strip.text=element_text(size=64),
                   legend.text=element_text(size=48),
                   legend.title=element_text(size=64),
                   legend.position="none")
figS2Plt2 <- ((plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_raw + labs(tag="C",
                                                                                                         x="Discovery Samples",
                                                                                                         y=bquote(atop(Log[2], (Abundance)))) +
                ggtitle("Un-Normalised")) +
             (plotList$`03_dataNormalisation_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm + labs(tag="D",
                                                                                                                     x="Discovery Samples",
                                                                                                                     y=bquote(atop(Log[2], (Abundance)))) +
                ggtitle("Cyclic LOESS-\nNormalised")) &
               theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank()) &
                scale_fill_manual(values=condCols,
                                  breaks=c(factor(c("CNO", 
                                                 "Crohn", 
                                                 "Healthy", 
                                                 "InfOsteitis",
                                                 "Oncology"),
                                                levels=c("CNO", 
                                                         "Crohn", 
                                                         "Healthy", 
                                                         "InfOsteitis",
                                                         "Oncology"))))) +
                plot_layout(guides="collect") &
                theme(plot.tag = element_text(face = "bold", 
                                              size = 100),
                      plot.title=element_text(size=72),
                      axis.text.y=element_text(size=64),
                      axis.title.y=element_text(size=84),
                      axis.title.x=element_text(size=72),
                      legend.text=element_text(size=72),
                      legend.title=element_text(size=84),
                      legend.title.position="left",
                      legend.position="bottom",
                      legend.direction="horizontal",
                      legend.key.size = unit(9, "line")) &
                guides(fill=guide_legend(nrow=2,
                                         byrow=TRUE))


# Plot
figS2Plt <- figS2Plt1 / figS2Plt2
figS2Plt

# Save
ggsave(here::here(figDir, 
                  "figS2_discoveryData_missingValueQC.png"),
       figS2Plt,
       width=10000,
       height=11000,
       units="px")
ggsave(here::here(figDir,
                  "figS2_discoveryData_missingValueQC.pdf"),
       figS2Plt,
       width=10000,
       height=11000,
       device="pdf",
       units="px")

rm(plotList)
rm(figS2Plt)

```

#### Figure S3

```{r, fig.height=28, fig.width=32}

# Read in 
plotList = list(`02_dataProcessQC_figs.rds` = readRDS(here::here("results", runName, "figs", "02_dataProcessQC_figs.rds")))

# Combine
figS3Plt1 <- ((plotList$`02_dataProcessQC_figs.rds`$UnFiltProtAbundanceSerumQC + labs(tag="A",
                                                                                      x="Serum QC") + ggtitle("All Proteins") +
                 theme(axis.text.y=element_text(size=88))) +
                
              (plotList$`02_dataProcessQC_figs.rds`$UnFiltProtAbundanceInstrQC + labs(tag="B",
                                                                                      x="Instrument QC") + ggtitle("All Proteins") +
                 theme(axis.text.y=element_text(size=88)))) +
  
              plot_layout(axis_titles="collect_y")

figS3Plt2 <- ((plotList$`02_dataProcessQC_figs.rds`$SerumQCFiltProtAbundanceSerumQC + labs(tag="C",
                                                                                            x="Serum QC") + ggtitle("Serum QC Proteins") +
                  theme(axis.text.y=element_blank())) +
              (plotList$`02_dataProcessQC_figs.rds`$SerumQCFiltProtAbundanceInstrQC + labs(tag="D",
                                                                                                        x="Instrument QC",
                                                                                                        y="Log2 Protein\nAbundance") + ggtitle("Serum QC Proteins") +
                              theme(axis.text.y=element_blank()))) +
              
              plot_layout(axis_titles="collect_y")

# Plot
figS3Plt <- figS3Plt1 / figS3Plt2 & 
  theme(plot.tag = element_text(face = "bold",
                                size = 100),
                    plot.title=element_text(size=96),
                    strip.text=element_text(size=88),
                    axis.text.x=element_text(size=64,
                                             angle=90,
                                             hjust=1,
                                             vjust=0.5),
                    axis.title=element_text(size=100))

# Plot
figS3Plt

# Save
ggsave(here::here(figDir, 
                  "figS3_discoveryData_qcSampleDistributions.png"),
       figS3Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "figS3_discoveryData_qcSampleDistributions.pdf"),
       figS3Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

rm(plotList)
rm(figS3Plt)
rm(figS3Plt1)
rm(figS3Plt2)

```


#### Figure S4

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`06_differentialAbundance_figs.rds` = readRDS(here::here("results", runName, "figs", "06_differentialAbundance_figs.rds")))

# Combine
figS4Plt1 <- ((plotList$`06_differentialAbundance_figs.rds`$sigDresdenProtsPPCA_colByCondSitePC12 + labs(tag="B") +
               ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins") +
               scale_fill_manual(breaks=factor(c("CNO_Dresden", 
                                                 "Crohn_Dresden",
                                                 "Oncology_Dresden", 
                                                 "Healthy_Liverpool",
                                                 "Healthy_Vanderbilt",
                                                 "InfOsteitis_Vanderbilt"),
                                 levels=c("CNO_Dresden", 
                                          "Crohn_Dresden",
                                          "Oncology_Dresden", 
                                          "Healthy_Liverpool",
                                          "Healthy_Vanderbilt",
                                          "InfOsteitis_Vanderbilt")),
                                 labels=c("CNO (Dresden)", 
                                          "Crohn (Dresden)",
                                          "Oncology (Dresden)", 
                                          "Healthy (Liverpool)",
                                          "Healthy (Vanderbilt)",
                                          "InfOsteitis (Vanderbilt)"),
                                 values=condSiteCols) +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     plot.title=element_text(size=64),
                     axis.text.y=element_text(size=56),
                     axis.title=element_text(size=72),
                     legend.text=element_text(size=60),
                     legend.title=element_text(size=68),
                     legend.position="bottom",
                     legend.direction="horizontal",
                     legend.title.position="top") +
               labs(fill="Condition\nAnd Site") +
               guides(fill=guide_legend(override.aes=list(size=20),
                                        nrow=3))) +
              (plotList$`06_differentialAbundance_figs.rds`$sigDresdenProtsPPCA_colByPlate + labs(tag="C") +
               ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins") +
               scale_fill_manual(labels=c("Plate 1",
                                          "Plate 2",
                                          "Plate 3"),
                                 values=plateCols) +
               theme(plot.tag = element_text(face = "bold", 
                                size = 100),
                     plot.title=element_text(size=64),
                     axis.text.y=element_text(size=56),
                     axis.title=element_text(size=72),
                     legend.text=element_text(size=72),
                     legend.title=element_text(size=84),
                     legend.position="bottom",
                     legend.direction="horizontal",
                     legend.title.position="top")) +
               guides(fill=guide_legend(override.aes=list(size=20),
                                        nrow=3)))
figS4Plt2 <- plot_spacer() +
            (ggplot() + theme_void() + labs(tag="A") + theme(plot.tag = element_text(face = "bold", size = 100))) +
             plot_spacer() +
             plot_layout(widths=c(0.5, 1, 0.5))
# figS4Plt3 <- (ggplot() + theme_void() + labs(tag="D") + theme(plot.tag = element_text(face = "bold", size = 100)))

# Combine
figS4Plt <- figS4Plt2 /
            figS4Plt1 +
            plot_layout(heights=c(1.5, 1))

# Visualise
figS4Plt

# Save
ggsave(here::here(figDir,
                  "figS4_ppcaOnDiffAbundSigProteins.png"),
       figS4Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS4_ppcaOnDiffAbundSigProteins.pdf"),
       figS4Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(plotList)
rm(figS4Plt)
rm(figS4Plt1)
rm(figS4Plt2)

```


#### Figure S5

Candidates summaries.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`04_eda_figs.rds` = readRDS(here::here("results", runName, "figs", "04_eda_figs.rds")))

plotList$`04_eda_figs.rds`$normAbundancePPCA_colByPlate$layers[[1]]$aes_params$size <- 16

# Plot PPCA
figS5Plt2 <- (plotList$`04_eda_figs.rds`$normAbundancePPCA_colByPlate +
              ggtitle("Probabilistic PCA On All Proteins")) +
              geom_point(size=16,
                         color="white") +
              geom_point(size=16,
                         mapping=aes(fill=PLATE),
                         alpha=0.85,
                         shape=21) +
            theme_classic() +
            theme(plot.tag = element_text(face = "bold", 
                                          size = 100),
                  plot.title=element_text(size=128),
                  axis.text=element_text(size=84),
                  axis.title=element_text(size=128),
                  legend.text=element_text(size=108),
                  legend.title=element_text(size=128),
                  legend.position="right",
                  legend.direction="vertical") &
            guides(fill=guide_legend(override.aes=list(size=28)))
 
figS5Plt <- figS5Plt2
figS5Plt

# Save
ggsave(here::here(figDir, 
                  "figS5_probPCAOnNormalisedData.png"),
       figS5Plt,
       width=10000,
       height=7500,
       units="px")
ggsave(here::here(figDir,
                  "figS5_probPCAOnNormalisedData.pdf"),
       figS5Plt,
       width=10000,
       height=7500,
       device="pdf",
       units="px")

rm(figS5Plt)
rm(figS5Plt1)
rm(figS5Plt2)
rm(plotList)

```

#### Figure S6

```{r, fig.height=32, fig.width=32}

# Read in 
plotList = list(`04_eda_figs.rds` = readRDS(here::here("results", runName, "figs", "06_differentialAbundance_figs.rds")))

# Plot PPCA
fig6Plt <- (((plotList$`04_eda_figs.rds`$sigDresdenProtsPPCA_colByCondSitePC12 + labs(tag="A") + ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins")) +
             (plotList$`04_eda_figs.rds`$sigDresdenProtsPPCA_colByCondSitePC23 + labs(tag="B") + ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins"))) /
            ((plotList$`04_eda_figs.rds`$sigDresdenProtsPPCA_colByCondSitePC34 + labs(tag="C") + ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins")) +
             (plotList$`04_eda_figs.rds`$sigDresdenProtsPPCA_colByCondSitePC45 + labs(tag="D") + ggtitle("PPCA on Dresden-Only\nDifferentially Abundant Proteins")))) +
            plot_layout(guides="collect") &
            labs(fill="Condition\nAnd Site") &
               scale_fill_manual(breaks=factor(c("CNO_Dresden", 
                                                 "Crohn_Dresden",
                                                 "Oncology_Dresden", 
                                                 "Healthy_Liverpool",
                                                 "Healthy_Vanderbilt",
                                                 "InfOsteitis_Vanderbilt"),
                                 levels=c("CNO_Dresden", 
                                          "Crohn_Dresden",
                                          "Oncology_Dresden", 
                                          "Healthy_Liverpool",
                                          "Healthy_Vanderbilt",
                                          "InfOsteitis_Vanderbilt")),
                                 labels=c("CNO (Dresden)", 
                                          "Crohn (Dresden)",
                                          "Oncology (Dresden)", 
                                          "Healthy (Liverpool)",
                                          "Healthy (Vanderbilt)",
                                          "InfOsteitis (Vanderbilt)"),
                                 values=condSiteCols) &
            theme(plot.tag = element_text(face = "bold", 
                                          size = 100),
                  plot.title=element_text(size=64),
                  axis.text=element_text(size=64),
                  axis.title=element_text(size=86),
                  legend.text=element_text(size=56),
                  legend.title=element_text(size=64),
                  legend.position="bottom",
                  legend.title.position="left",
                  legend.direction="horizontal") &
            guides(fill=guide_legend(override.aes=list(size=20)))

fig6Plt

# Save
ggsave(here::here(figDir, 
                  "figS6_ppcaByCondSite.png"),
       fig6Plt,
       width=10000,
       height=11500,
       units="px")
ggsave(here::here(figDir, 
                  "figS6_ppcaByCondSite.pdf"),
       fig6Plt,
       width=10000,
       height=11500,
       units="px")
# 
# # rm(fig6Plt)
# rm(plotList)

```

#### Figure S8

#### Figure S9

#### Figure S10

Permutation importance of discovery data.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

# Combine
figS10Plt <- (plotList$`10_finalModel_figs.rds`$varImportancePerGroup_discovery_biModal + labs(tag="A") +
                scale_x_continuous(breaks=c(0, 0.03, 0.06)) +
                ggtitle("Bi-Modal")) /
             (plotList$`10_finalModel_figs.rds`$varImportancePerGroup_discovery_biModalWithMultiGroupProteins + labs(tag="B") +
                scale_x_continuous(breaks=c(0, 0.02, 0.04)) +
                ggtitle("Bi-Modal with All Panel Proteins")) +
  plot_layout(axis_titles = "collect_x",
              guides = "collect",
              heights = c(0.45,1.1)) &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=100),
        axis.text.y=element_text(size=56),
        axis.text.x=element_text(size=48),
        axis.title.y=element_text(size=84),
        axis.title.x=element_text(size=84),
        strip.text=element_text(size=72),
        legend.title=element_text(size=72),
        legend.text=element_text(size=64),
        legend.justification='left',
        legend.position="bottom") &
  labs(y="Selected Protein",
       x="Permutation Importance") &
  scale_color_manual(values=condCols) &
  guides(color=guide_legend(override.aes=list(size=24),
                            nrow=1,
                            byrow=TRUE))

figS10Plt

# Save
ggsave(here::here(figDir, 
                  "figS10_discoveryModel_permutationImportance_biModal.png"),
       figS10Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS10_discoveryModel_permutationImportance_biModal.pdf"),
       figS10Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(plotList)
rm(figS10Plt)

```

#### Figure S11

Permutation importance of discovery data.

```{r, fig.height=24, fig.width=32}

# Read in 
plotList = list(`10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

# Combine
figS11Plt <- (plotList$`10_finalModel_figs.rds`$varImportancePerGroup_discovery_multiGroup +
                scale_x_continuous(breaks=c(0, 0.03, 0.06)) +
                ggtitle("Multi-Modal")) +
  plot_layout(axis_titles = "collect_x") &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=100),
        axis.text.y=element_text(size=56),
        axis.text.x=element_text(size=48),
        axis.title.y=element_text(size=84),
        axis.title.x=element_text(size=84),
        strip.text=element_text(size=72),
        legend.title=element_text(size=72),
        legend.text=element_text(size=64),
        legend.justification='left',
        legend.position="bottom") &
  labs(y="Selected Protein",
       x="Permutation Importance") &
  scale_color_manual(values=condCols) &
  guides(color=guide_legend(override.aes=list(size=24),
                            nrow=1,
                            byrow=TRUE))

figS11Plt

# Save
ggsave(here::here(figDir, 
                  "figS11_discoveryModel_permutationImportance_multiGroup.png"),
       figS11Plt,
       width=10000,
       height=8500,
       units="px")
ggsave(here::here(figDir,
                  "figS11_discoveryModel_permutationImportance_multiGroup.pdf"),
       figS11Plt,
       width=10000,
       height=8500,
       device="pdf",
       units="px")

rm(figS11Plt)
rm(plotList)

```

#### Figure S12

Create standalone legend for evaluating randomisation

```{r, fig.height=7, fig.width=32}

# Modify iris
data(iris)
iris$Species <- as.character(iris$Species)
iris <- iris[iris$Species != "virginica",]
iris[iris$Species == "setosa",]["Species"] <- "Observed"
iris[iris$Species == "versicolor",]["Species"] <- "Randomised"

# Get dummy plot
iris$Species <- factor(iris$Species)
dummyPlt <- ggplot(iris,
                   aes(y=Sepal.Length,
                       x=Species,
                       color=Species)) + 
                   geom_vline(aes(xintercept=Species,
                                  color=Species),
                              linetype=1,
                              linewidth=3.5,
                              key_glyph = "path") +
                   scale_color_manual(values=c("Observed"="mediumblue",
                                               "Randomised"="salmon")) +
                   geom_bar(stat="identity",
                            color="black") +
                   labs(color="Median MCC") +
                   theme_minimal() +
                   theme(legend.direction="vertical",
                         legend.title.position="left",
                         legend.key.width=unit(20, "line"),
                         legend.title=element_text(size=88),
                         legend.text=element_text(size=80)) +
  guides(color=guide_legend(override.aes=list(linewidth=4)))

# Using the cowplot package
legend <- cowplot::get_legend(dummyPlt)
dummyPlt

```

Plot shuffled label and decoy protein abundances for 10-fold training data. On discovery data.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

# Combine
figS12Plt <- (plotList$`10_finalModel_figs.rds`$shuffledLabelsMCC_biModalProts + labs(tag="A")) /
           # plot_spacer() /
            (plotList$`10_finalModel_figs.rds`$shuffledLabelsMCC_multiModalProts + theme(legend.position="none") + labs(tag="B")) /
           # plot_spacer() /
           (plotList$`10_finalModel_figs.rds`$decoyProtsMCC_biModalProts + labs(tag="C")) /
           # plot_spacer() /
            (plotList$`10_finalModel_figs.rds`$decoyProtsMCC_multiModalProts + labs(tag="D")) /
           (legend) +
  plot_layout(heights = c(1,1,1,1,0.5)) &
  labs(y="MCC") &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=64),
        legend.title=element_blank(),
        axis.text.y=element_text(size=42),
        axis.title.y=element_text(size=72),
        axis.title.x=element_text(size=72),
        strip.text.x=element_text(size=72),
        strip.text.y=element_text(size=52))

figS12Plt

# Save
ggsave(here::here(figDir, 
                  "figS12_discoveryModel_10FoldSplit_randomisationPerformance.png"),
       figS12Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS12_discoveryModel_10FoldSplit_randomisationPerformance.pdf"),
       figS12Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(figS12Plt)
rm(plotList)

```

#### Figure S13

Plot shuffled label and decoy protein abundances for held-out test data. On discovery data

```{r, fig.height=24, fig.width=32}

# Read in 
plotList = list(`10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))


plotList$`10_finalModel_figs.rds`$shuffledLabelsMCC_heldOutTestData$layers[[3]]$aes_params$width <- 0.3

# Combine
figS13Plt <- (plotList$`10_finalModel_figs.rds`$shuffledLabelsMCC_heldOutTestData + labs(tag="A") +
                theme(legend.position = "none")) /
             (plotList$`10_finalModel_figs.rds`$decoyProtsMCC_heldOutTestData + labs(tag="B") +
                labs(fill="Condition") +
                theme(legend.position="bottom",
                      legend.direction="horizontal",
                      legend.text=element_text(size=68),
                      legend.key.size = unit(6.5, "line"),
                      legend.title=element_text(size=76))) +
  plot_layout(heights = c(1,1)) &
  labs(y="MCC",
       x="Predictions") &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=72),
        axis.text.y=element_text(size=72),
        axis.title.y=element_text(size=100),
        axis.title.x=element_text(size=100),
        strip.text.x=element_text(size=72),
        strip.text.y=element_text(size=84))
figS13Plt

# Save
ggsave(here::here(figDir, 
                  "figS13_discoveryModel_heldOutData_randomisationPerformance.png"),
       figS13Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "figS13_discoveryModel_heldOutData_randomisationPerformance.pdf"),
       figS13Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

rm(figS13Plt)
rm(plotList)

```

#### Figure S14

Missing protein/sample enumaration, filtration and abundance distribution evaluation.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`11_validationDataProcessQC_figs.rds` = readRDS(here::here("results", runName, "figs", "11_validationDataProcessQC_figs.rds")),
                `12_validationDataNormEDA_figs.rds` = readRDS(here::here("results", runName, "figs", "12_validationDataNormEDA_figs.rds")))

# Change text size
plotList$`11_validationDataProcessQC_figs.rds`$NProtsAllProts_validData$layers[[3]]$aes_params$size <- 20
plotList$`11_validationDataProcessQC_figs.rds`$NProtsSerumQCProts_validData$layers[[3]]$aes_params$size <- 20
plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_raw_validData$layers[[1]]$aes_params$linewidth <- 0.5
plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_raw_validData$layers[[1]]$aes_params$size <- 2
plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm_validData$layers[[1]]$aes_params$linewidth <- 0.5
plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm_validData$layers[[1]]$aes_params$size <- 2

# Combine
figS14Plt <- ((plotList$`11_validationDataProcessQC_figs.rds`$UnFiltProtAbundanceSerumQC_validData + labs(tag="A",
                                                                                                          x="Serum QC",
                                                                                                          y="Log2 Protein\nAbundance") +
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%")) +
                 ggtitle("All Proteins") +
                 theme(axis.text.x=element_text(size=56),
                       axis.title.y=element_text(size=72),
                       axis.text.y=element_text(size=56),
                       axis.title.x=element_text(size=72))) +
              (plotList$`11_validationDataProcessQC_figs.rds`$SerumQCFiltProtAbundanceSerumQC_validData + labs(tag="B",
                                                                                                          x="Serum QC",
                                                                                                          y="Log2 Protein\nAbundance") +
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%")) +
                 ggtitle("Filtered for Serum QC Proteins") +
                 theme(axis.text.x=element_text(size=56),
                       axis.title.y=element_blank(),
                       axis.text.y=element_blank(),
                       axis.title.x=element_text(size=72))) + plot_layout(axis_titles="collect")) /
             ((plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_raw_validData + labs(tag="C",
                                                                                                     
                                                                                                                                              x="Validation Samples",
                                                                                                                           y=expression(atop(Log[2],
                                                                                                                                             "(Abundance)"))) +
                 ggtitle("Un-Normalised")) +
                 theme(axis.title.y=element_text(size=72),
                       axis.text.y=element_text(size=56)) +
              (plotList$`12_validationDataNormEDA_figs.rds`$sampleAbundDistribution_colByCondition_cyclicLoessNorm_validData + labs(tag="D",
                                                                                                                                    x="Validation Samples",
                                                                                                                                    y=expression(atop(Log[2],
                                                                                                                                                      "(Abundance)")))) /
                 ggtitle("Cyclic LOESS-Normalised") +
                 theme(axis.title.y=element_blank(),
                       axis.text.y=element_blank())) +
             ((plotList$`11_validationDataProcessQC_figs.rds`$NProtsAllProts_validData + labs(tag="E",
                                                                                    y="Number of\nProteins") +  
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%")) +
                 scale_y_continuous(limits=c(0, 1050)) +
                 ggtitle("Unfiltered") +
                 labs(x="Threshold of\nMissing Values") +
                 theme(axis.text.x=element_text(size=52),
                       axis.title.x=element_text(size=64),
                       axis.text.y=element_text(size=56))) +
              (plotList$`11_validationDataProcessQC_figs.rds`$NProtsSerumQCProts_validData + labs(tag="F",
                                                                                    y="Number of\nProteins") + 
               scale_x_discrete(labels=c("UnFilt", "25%", "20%", "15%", "10%", "5%")) +
                 scale_y_continuous(limits=c(0, 1050)) +
                 ggtitle("Filtered for Serum QC Proteins") +
                 labs(x="Threshold of\nMissing Values") +
                 theme(axis.text.x=element_text(size=52),
                       axis.title.x=element_text(size=64),
                       axis.title.y=element_blank(),
                       axis.text.y=element_blank()))) +
  plot_layout(guides="collect",
              axis_titles = "collect_y") &
                scale_fill_manual(values=condCols,
                                  breaks=c(factor(c("CNO", 
                                                 "Crohn", 
                                                 "Healthy", 
                                                 "InfOsteitis",
                                                 "Oncology"),
                                                levels=c("CNO", 
                                                         "Crohn", 
                                                         "Healthy", 
                                                         "InfOsteitis",
                                                         "Oncology")))) &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=64),
        axis.title=element_text(size=84),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.text=element_text(size=68),
        legend.title=element_text(size=72),
        legend.key.size=unit(4.5, "cm"),
        legend.title.position="left",
        legend.position="bottom",
        legend.direction="horizontal") &
  guides(fill=guide_legend(nrow=2,
                           byrow=TRUE))

# Plot
figS14Plt

# Save
ggsave(here::here(figDir, 
                  "figS14_validationData_sampleQC.png"),
       figS14Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS14_validationData_sampleQC.pdf"),
       figS14Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# rm(plotList)
rm(figS14Plt)

```

#### Figure S15

Comparative differential abundance analysis comparing Discovery to Validation.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`14_validationDataDiffAbundance_figs.rds` = readRDS(here::here("results", runName, "figs", "14_validationDataDiffAbundance_figs.rds")))

# Edit point sizes
plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid$layers[[4]]$aes_params$size <- 18
plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid_justCRP$layers[[4]]$aes_params$size <- 18
plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid$layers[[1]]$aes_params$linewidth <- 2
plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid_justCRP$layers[[1]]$aes_params$linewidth <- 2

# Plot
figS15Plt <- ((plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid + 
                labs(x="",
                     y="Candidate Protein") +
                facet_grid(cols=vars(Condition),
                           rows=vars(SigInValidation),
                           labeller = labeller(SigInValidation=c(ns="ns",
                                                                 `FDR < 0.05`="Sig")),
                           scale="free",
                           space="free_y") +
                scale_x_continuous(limits=c(-2.75, 2.75),
                                   breaks=c(-2, 0, 2),
                                   labels=c(-2, 0, 2)) +
                theme(legend.position="none",
                      legend.title.position="top",
                      legend.justification = "left",
                      panel.spacing.y = unit(2, "lines"),
                      strip.text.x=element_text(size=80),
                      strip.text.y=element_text(size=92),
                      axis.title.x=element_blank()) +
                 guides(fill=guide_legend(byrow=TRUE))) /
             (plot_spacer() + labs(y="") + theme(legend.position="none")) /
             (plotList$`14_validationDataDiffAbundance_figs.rds`$barbellPlot_logFC_discoVsValid_justCRP + 
                labs(y="",
                     x="Log FC",
                     shape="Consistent\nLogFC",
                     fill="Condition\nBy Cohort") +
                facet_grid(cols=vars(Condition),
                           rows=vars(SigInValidation),
                           labeller = labeller(SigInValidation=c(ns=" ",
                                                                 `FDR < 0.05`="Sig")),
                           scale="free",
                           space="free_y") +
                theme(legend.position = "bottom",
                      legend.direction="horizontal",
                      legend.text=element_text(size=88),
                      legend.title=element_text(size=100),
                      legend.title.position="top",
                      panel.spacing.y = unit(2, "lines"),
                      strip.text.y=element_text(size=92),
                      strip.text.x=element_blank(),
                      legend.key.size = unit(6, "line")) +
                guides(fill=guide_legend(override.aes=list(size=36,
                                                           shape=21,
                                                           pattern="none"),
                                         ncol=2,
                                         byrow=TRUE),
                       shape=guide_legend(override.aes=list(size=36,
                                                            ncol=1,
                                                            shape=c(22, 21)),
                                          ncol=1),
                       alpha="none")) + 
             plot_layout(axis_titles="collect") &
             theme(axis.text=element_text(size=56)) & 
  scale_fill_manual(breaks=factor(c("Discovery.Crohn", 
                                    "Validation.Crohn",
                                    "Discovery.Healthy",
                                    "Validation.Healthy",
                                    "Discovery.InfOsteitis",
                                    "Validation.InfOsteitis", 
                                    "Discovery.Oncology",
                                    "Validation.Oncology"),
                                  levels=c("Discovery.Crohn", 
                                           "Validation.Crohn",
                                           "Discovery.Healthy" ,
                                           "Validation.Healthy",
                                           "Discovery.InfOsteitis",
                                           "Validation.InfOsteitis",
                                           "Discovery.Oncology",
                                           "Validation.Oncology")),
                      labels=c("Crohn (D)",
                               "Crohn (V)",
                               "Healthy (D)",
                               "Healthy (V)",
                               "InfOsteitis (D)",
                               "InfOsteitis (V)",
                               "Oncology (D)",
                               "Oncology (V)"),
                      values=condStudyCols)) &
                      plot_layout(heights=c(1, -0.075, 0.105)) &
                      theme(plot.tag = element_text(face = "bold", 
                                                    size = 100),
                            strip.background.x=element_blank(),
                            axis.title.y=element_text(size=108),
                            axis.title.x=element_text(size=108),
                            axis.text.x=element_text(size=80),
                            axis.text.y=element_text(size=70)) 

# Save
ggsave(here::here(figDir, 
                  "figS15_differentialAbundance_discoveryVsValidation_barbellLogFC.png"),
       figS15Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS15_differentialAbundance_discoveryVsValidation_barbellLogFC.pdf"),
       figS15Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

# Remove to clear space
rm(figS15Plt)
rm(plotList)
gc()

```

#### Figure S16

Showing the variance of CRP

```{r, fig.height=32, fig.width=32}

# Read in 
plotList = list(`15_validationDataPredictionModelling_figs_bPCA.rds` = readRDS(here::here("results", runName, "figs", "15_validationDataPredictionModelling_figs_bPCA.rds")),
                `10_finalModel_figs.rds` = readRDS(here::here("results", runName, "figs", "10_finalModel_figs.rds")))

# Update point sizes
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_byCRPAbundance_inDiscoData_bPCA@layers$geom_point$aes_params["size"] <- 12
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_byCRPAbundance_bPCA@layers$geom_point$aes_params["size"] <- 12

# Increase widths of boxplots
plotList$`10_finalModel_figs.rds`$CNO_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["width"] <- 25
plotList$`10_finalModel_figs.rds`$Healthy_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["width"] <- 25
plotList$`10_finalModel_figs.rds`$Crohn_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["width"] <- 25
plotList$`10_finalModel_figs.rds`$InfOsteitis_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["width"] <- 25
plotList$`10_finalModel_figs.rds`$Oncology_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["width"] <- 25

plotList$`10_finalModel_figs.rds`$CNO_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`10_finalModel_figs.rds`$Healthy_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`10_finalModel_figs.rds`$Crohn_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`10_finalModel_figs.rds`$InfOsteitis_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`10_finalModel_figs.rds`$Oncology_relAbundancePlot_crpAbundance_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"

plotList$`10_finalModel_figs.rds`$CNO_relAbundancePlot_crpAbundance_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`10_finalModel_figs.rds`$Healthy_relAbundancePlot_crpAbundance_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`10_finalModel_figs.rds`$Crohn_relAbundancePlot_crpAbundance_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`10_finalModel_figs.rds`$InfOsteitis_relAbundancePlot_crpAbundance_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`10_finalModel_figs.rds`$Oncology_relAbundancePlot_crpAbundance_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"

plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$CNO_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Healthy_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Crohn_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$InfOsteitis_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Oncology_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_boxplot$aes_params["fill"] <- "white"

plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$CNO_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Healthy_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Crohn_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$InfOsteitis_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Oncology_relAbundancePlot_crpAbundance_vaildData_bPCA@layers$geom_beeswarm$aes_params["colour"] <- "black"

# Plot ranked CRP abundances
p1 <- (( plotList$`10_finalModel_figs.rds`$CNO_relAbundancePlot_crpAbundance_bPCA + labs(x="",
                                                                                         title="CNO",
                                                                                        y="Log2\nAbundance") + theme(axis.text.y=element_text(size=64)) |
        plotList$`10_finalModel_figs.rds`$Healthy_relAbundancePlot_crpAbundance_bPCA + labs(x="",
                                                                                            title="Healthy",
                                                                                        y="") +
                                                                                        theme(axis.text.y=element_blank()) |
        plotList$`10_finalModel_figs.rds`$Crohn_relAbundancePlot_crpAbundance_bPCA + labs(y="",
                                                                                          title="Crohn",
                                                                                          x="Rank of Protein") +
                                                                                     theme(axis.text.y=element_blank()) |
        plotList$`10_finalModel_figs.rds`$InfOsteitis_relAbundancePlot_crpAbundance_bPCA + labs(x="",
                                                                                                title="Inf Osteitis",
                                                                                                y="") +
                                                                                           theme(axis.text.y=element_blank()) |
        plotList$`10_finalModel_figs.rds`$Oncology_relAbundancePlot_crpAbundance_bPCA + labs(x="",
                                                                                             title="Oncology",
                                                                                             y="") +
                                                                                        theme(axis.text.y=element_blank())) &
        scale_y_continuous(limits=c(0, 17),
                           breaks=c(5, 10, 15))) &
  theme(plot.title=element_text(size=64))
p2 <- (( plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$CNO_relAbundancePlot_crpAbundance_vaildData_bPCA + labs(x="",
                                                                                                                              y="Log2\nAbundance",
                                                                                                                              title="CNO") + theme(axis.text.y=element_text(size=64)) |
        plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Healthy_relAbundancePlot_crpAbundance_vaildData_bPCA + labs(x="",
                                                                                                                                  y="",
                                                                                                                                  title="Healthy") +
                                                                                                                             theme(axis.text.y=element_blank()) |
        plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Crohn_relAbundancePlot_crpAbundance_vaildData_bPCA + labs(y="",
                                                                                                                                x="Rank of Protein",
                                                                                                                                title="Crohn") +
                                                                                                                           theme(axis.text.y=element_blank()) |
        plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$InfOsteitis_relAbundancePlot_crpAbundance_vaildData_bPCA + labs(x="",
                                                                                                                                      y="",
                                                                                                                                      title="Inf Osteoitis") +
                                                                                                                           theme(axis.text.y=element_blank()) |
        plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$Oncology_relAbundancePlot_crpAbundance_vaildData_bPCA + labs(x="",
                                                                                                                                   y="",
                                                                                                                                   title="Oncology") +
                                                                                                                           theme(axis.text.y=element_blank())) &
        scale_y_continuous(limits=c(0, 17),
                           breaks=c(5, 10, 15))) &
  theme(plot.title=element_text(size=64))

# Mis-prediction
p3 <- ((plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_byCRPAbundance_inDiscoData_bPCA + labs(title="Discovery") + guides(fill="none")) | 
        plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$misPredict_byCRPAbundance_bPCA + labs(y="",
                                                                                                            title="Validation")) + 
        guides(fill=guide_legend(override.aes=list(size=24,
                                 shape=21),
                                 nrow=1),
        ) +
                scale_fill_manual(values=condCols,
                                  breaks=c(factor(c("CNO", 
                                                 "Crohn", 
                                                 "Healthy"),
                                                levels=c("CNO", 
                                                         "Crohn", 
                                                         "Healthy")))) +
  plot_layout(guides="collect",
              ncol=2) &
  labs(x="CNO Samples",
       fill="Prediction") &
  theme(legend.position="bottom",
        legend.title.position="top",
        legend.title=element_text(size=72),
        legend.text=element_text(size=64),
        axis.text.y=element_text(size=64),
        plot.title=element_text(size=96)) &
  guides(shape=guide_legend(override.aes=list(size=24),
                            nrow=1))
p3

# Define design
design <- "AAAAA
           BBBBB
           KKKKK"

# Combine
figS16Plt <- wrap_plots(list(wrap_elements(p1),
                             wrap_elements(p2),
                             wrap_elements(p3)),
                        design = design) +
             plot_layout(heights=c(1, 1, 1.65)) +
             plot_annotation(tag_levels = "A") &
             theme(axis.title.x=element_text(size=84),
                   axis.text.x=element_text(size=64),
                   plot.tag=element_text(size=100,
                                         face = "bold"),
                   plot.tag.position = c(0.025, 0.94) )
figS16Plt

# Save
ggsave(here::here(figDir, 
                  "figS16_abundanceVariability_misPrediction_CRP.png"),
       figS16Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "figS16_abundanceVariability_misPrediction_CRP.pdf"),
       figS16Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

rm(figS16Plt)
rm(plotList)

```

#### Figure S17

Plot shuffled label and decoy protein abundances for discovery data, using proteins detected in validation

```{r, fig.height=24, fig.width=32}

# Read in 
plotList = list(`15_validationDataPredictionModelling_figs_bPCA.rds` = readRDS(here::here("results", runName, "figs", "15_validationDataPredictionModelling_figs_bPCA.rds")))

# Combine
figS17Plt <- (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$shuffledLabelsMCC_multiGroup_validData_bPCA + labs(tag="A") +
                theme(legend.position = "none")) /
             (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$decoyProtsMCC_multiGroup_validData_bPCA + labs(tag="B") +
                labs(fill="Condition") +
                theme(legend.position="bottom",
                      legend.direction="horizontal",
                      legend.text=element_text(size=72),
                      legend.key.size = unit(7, "line"),
                      legend.title=element_text(size=80)) +
                guides(fill=guide_legend(nrow=2,
                                         byrow=TRUE)) +
                scale_fill_manual(values=condCols,
                                  breaks=c(factor(c("CNO",
                                                 "Crohn",
                                                 "Healthy",
                                                 "InfOsteitis",
                                                 "Oncology"),
                                                levels=c("CNO",
                                                         "Crohn",
                                                         "Healthy",
                                                         "InfOsteitis",
                                                         "Oncology"))))) +
                plot_layout(heights = c(1,1)) &
                labs(y="MCC",
                     x="Predictions") &
                theme(plot.tag = element_text(face = "bold", 
                                              size = 100),
                      plot.title=element_text(size=72),
                      axis.text.y=element_text(size=72),
                      axis.title.y=element_text(size=100),
                      axis.title.x=element_text(size=100),
                      strip.text.x=element_text(size=72),
                      strip.text.y=element_text(size=80))

# Save
ggsave(here::here(figDir, 
                  "figS17_validationData_randomisationPerformance.png"),
       figS17Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "figS17_validationData_randomisationPerformance.pdf"),
       figS17Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

rm(figS17Plt)
rm(plotList)

```

#### Figure S18

Model metric performances on the discovery cohort.

```{r, fig.height=40, fig.width=32}

# Read in 
plotList = list(`15_validationDataPredictionModelling_figs_bPCA.rds` = readRDS(here::here("results", runName, "figs", "15_validationDataPredictionModelling_figs_bPCA.rds")))

# Change size and alpha of points
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModal_bPCA$layers[[2]]$aes_params$size <- 10
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModal_bPCA$layers[[2]]$aes_params$alpha <- 0.675
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModalWithMultiGroupProteins_bPCA$layers[[2]]$aes_params$size <- 10
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModalWithMultiGroupProteins_bPCA$layers[[2]]$aes_params$alpha <- 0.675
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_multiGroup_bPCA$layers[[2]]$aes_params$size <- 8
plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_multiGroup_bPCA$layers[[2]]$aes_params$alpha <- 0.675


# --- create a common theme to reduce grob-width differences ---
common_theme <- theme(
  strip.text = element_text(size = 52),
  axis.text.x  = element_text(size = 64),
  plot.margin = margin(5, 5, 5, 5)
)

# --- build each sub-plot separately for clarity ---
pA <- plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModal_bPCA +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "A", title = "Bi-Modal") +
  common_theme +
  theme(axis.text.x = element_text(size=72),
        axis.text.y = element_text(size = 64),
        axis.title.y = element_text(size = 100,
                                        angle=90),
        legend.position="none") +
  # hide fill legend here so only one plot supplies it to be collected
  guides(color = "none")

pB <- plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_biModalWithMultiGroupProteins_bPCA +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "B", 
       title = "Bi-Modal\n(Whole Panel)") +
  common_theme +
  # remove redundant left axis elements to avoid double-reserving left space
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 72),
        legend.position="none")

# This plot will supply the legend (so don't hide its legend)
pC <- plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$modelMetrics_discoveryWithOnlyValidProts_multiGroup_bPCA +
  facet_wrap(~ variable, ncol = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "C") +
  common_theme +
  # remove left y-axis items so they don't take extra space
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = c(-0.275, -0.1),
        legend.direction="horizontal",
        axis.text.x = element_text(size = 72, 
                                   angle = 90, 
                                   hjust = 1, 
                                   vjust = 0.5))

# bottom row plot (held-out data)
pD <- plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$multiModelProtsMetrics_heldOutData_validData_bPCA +
  geom_text(mapping = aes(y = value - 0.05, label = round(value, 2)),
            color = "white", hjust = 1, angle = 90, size = 24) +
  facet_wrap(~ variable, nrow = 2,
             labeller=labeller(
               variable=c(
                          `Bal Acc` = "B Acc",
                          AUCROC = "AUC",
                          Specificity = "Spec",
                          Sensitivity = "Sens"
               )
             )) +
  labs(tag = "D") +
  common_theme +
  theme(legend.direction = "horizontal",
        legend.position = "bottom",
        legend.key.size = unit(6, "line"),
        legend.title = element_text(size = 88),
        legend.text = element_text(size = 80),
        axis.text.x = element_blank())

# --- Combine with patchwork ---
# Important notes applied:
# - only pC provides the legend (show.legend = TRUE); guides = "collect" will therefore
#   collect a single, consistent legend and place it where we request
# - widths are set to favour the right-hand multi-panel plot, but because we've
#   removed duplicated left-axis grob widths and standardized theme settings,
#   columns will be much more proportional
figS18Plt <- (
  (pA + pB + pC) +
    plot_layout(widths = c(1, 1, 1.5)) &
    labs(color="Condition") &
    theme(
      legend.title = element_text(size = 88),
      legend.text = element_text(size = 80),
      axis.text.x = element_blank()) &
    guides(color=guide_legend(override.aes=list(size=28),
                              nrow=2,
                              byrow=TRUE))
) /
  plot_spacer() /
  (pD + 
  theme(axis.title.y = element_text(size = 100,
                                    angle=90),
        axis.text.y  = element_text(size = 64))) +
  plot_layout(heights = c(1, 0.1, 0.75), axes = "collect_y", axis_titles = "collect_y") &
  labs(fill = "Condition") &
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.05, 1.05)) & 
  theme(
    plot.tag = element_text(face = "bold", size = 100),
    plot.title = element_text(size = 88),
    strip.text = element_text(size = 84),
    axis.title.x = element_blank(),
  ) &
               scale_fill_manual(values=condCols,
                                 breaks=c(factor(c("CNO", 
                                                "Crohn", 
                                                "Healthy", 
                                                "InfOsteitis",
                                                "Oncology"),
                                               levels=c("CNO", 
                                                        "Crohn", 
                                                        "Healthy", 
                                                        "InfOsteitis",
                                                        "Oncology")))) &
               scale_color_manual(values=condCols,
                                  breaks=c(factor(c("CNO", 
                                                 "Crohn", 
                                                 "Healthy", 
                                                 "InfOsteitis",
                                                 "Oncology"),
                                                levels=c("CNO", 
                                                         "Crohn", 
                                                         "Healthy", 
                                                         "InfOsteitis",
                                                         "Oncology")))) &
  guides(fill=guide_legend(nrow=2,
                           byrow=TRUE))

figS18Plt

# Save
ggsave(here::here(figDir, 
                  "figS18_discoveryModelPerformances_filtByProteinsDetectedInValidation.png"),
       figS18Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS18_discoveryModelPerformances_filtByProteinsDetectedInValidation.pdf"),
       figS18Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(figS18Plt)

```

#### Figure S19

Plot shuffled label and decoy protein abundances for 10-fold training data. On discovery data, using the subset of candidate proteins also detected in validation.

```{r, fig.height=40, fig.width=32}

# Combine
figS19Plt <- (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$shuffledLabelsMCC_discoveryWithValidProts_biModalProts + labs(tag="A",
                                                                                                                                          y="MCC")) /
           # plot_spacer() /
            (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$shuffledLabelsMCC_discoveryWithValidProts_multiModalProts + theme(legend.position="none") + labs(tag="B",
                                                                                                                                                                            y="MCC",)) /
           # plot_spacer() /
           (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$decoyProtsMCC_biModalProts_discoveryWithValidProts + labs(tag="C",
                                                                                                                                    y="MCC",
                                                                                                                                    x="Random Decoy Protein Sets")) / # discoveryWithValidProts
           # plot_spacer() /
            (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$decoyProtsMCC_multiModalProts_discoveryWithValidProts + labs(tag="D",
                                                                                                                                        y="MCC",
                                                                                                                                        x="Random Decoy Protein Sets")) / # discoveryWithValidProts
           (legend) +
  plot_layout(heights = c(1,1,1,1,0.5)) &
  labs(y="MCC") &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=72),
        axis.text.y=element_text(size=42),
        axis.title.y=element_text(size=72),
        axis.title.x=element_text(size=72),
        strip.text.x=element_text(size=72),
        strip.text.y=element_text(size=52))

figS19Plt

# Save
ggsave(here::here(figDir, 
                  "figS19_discoveryData_randomisationPerformance_filtByProteinsDetectedInValidation.png"),
       figS19Plt,
       width=10000,
       height=12500,
       units="px")
ggsave(here::here(figDir,
                  "figS19_discoveryData_randomisationPerformance_filtByProteinsDetectedInValidation.pdf"),
       figS19Plt,
       width=10000,
       height=12500,
       device="pdf",
       units="px")

rm(figS19Plt)

```

#### Figure S20

Plot shuffled label and decoy protein abundances for held-out discovery test data, using the subset of candidate proteins also detected in validation.

```{r, fig.height=24, fig.width=32}

# Combine
figS20Plt <- (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$shuffledLabelsMCC_heldOutTestData_discoveryWithValidProts_bPCA + labs(tag="A") +
                theme(legend.position="none")) +
             (plotList$`15_validationDataPredictionModelling_figs_bPCA.rds`$decoyProtsMCC_heldOutTestData_discoveryWithValidProts_bPCA + labs(tag="B") +
                labs(fill="Condition") +
                theme(legend.position="bottom",
                      legend.direction="horizontal",
                      legend.text=element_text(size=68),
                      legend.key.size = unit(6.5, "line"),
                      legend.title=element_text(size=76))) +
  plot_layout(heights = c(1,1)) &
  labs(y="MCC",
       x="Predictions") &
  theme(plot.tag = element_text(face = "bold", 
                                size = 100),
        plot.title=element_text(size=72),
        axis.text.y=element_text(size=72),
        axis.title.y=element_text(size=100),
        axis.title.x=element_text(size=100),
        strip.text.x=element_text(size=72),
        strip.text.y=element_text(size=84))

figS20Plt

# Save
ggsave(here::here(figDir, 
                  "figS20_discoveryData_heldOutData_randomisationPerformance_filtByProteinsDetectedInValidation.png"),
       figS20Plt,
       width=10000,
       height=10000,
       units="px")
ggsave(here::here(figDir,
                  "figS20_discoveryData_heldOutData_randomisationPerformance_filtByProteinsDetectedInValidation.pdf"),
       figS20Plt,
       width=10000,
       height=10000,
       device="pdf",
       units="px")

rm(figS20Plt)
rm(plotList)

```


